---
title: "ScenarioAnalysis_Retrospective"
author: "Laura Geronimo"
date: "2024-09-16"
output: html_document
---

In this worksheet we analyse and compare the total costs to society for S0 (no Sandy), using best estimates.


Recall the assumptions for S0 applied in prior worksheets:

Assumptions:

Federal policy costs:
-Federal Cost Share for Municipal Response and Recovery (NA - no Sandy)
-Federal cost share for response and recovery direct to household (NA - no Sandy)
-Federal IA / NFIP claims (post 2012 events only)
-NFIP policies (Assume NFIP policies same as 2012, with 3%  growth rate on par with change in assessed value)
-Federal cost share for beach nourishment (2021 event only)

Municipal policy costs:
-Municipal cost share for federal response and recovery (NA - no Sandy)
-Municipal cost share for beach nourishment (2022 event only)

Residual Risk:
- Assume 10x as high as First Street estimates, because we assume that homes were not elevated to BFE post-Sandy. Three climate assumptions (low, mid, high)

Protection costs:
- No initial beach nourishment associated with Sandy
- Maintenance costs: Test three assumptions (low, mid, high)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)
library(ggplot2)
library(scales)
library(knitr)
library(formattable)

options(scipen=999)


##Importing Data ####
####Retrospective ####
#S0
S0_FedCostOB_NPV <- read.csv('./data/BaselineData/S0_OB_Inputs/S0_FedCostOB_NPV.csv')
S0_FedRevOB_NPV <- read.csv('./data/BaselineData/S0_OB_Inputs/S0_FedRevOB_NPV.csv')
S0_MuniCostOB_NPV <- read.csv('./data/BaselineData/S0_OB_Inputs/S0_MuniCostOB_NPV.csv')
S0_MuniRevOB_NPV <- read.csv('./data/BaselineData/S0_OB_Inputs/S0_MuniRevOB_NPV.csv')


####Prospective ####
##### Residual Risk (assume 10x, due to no elevations post Sandy####
S0_NPV_ResRisk <- read.csv('./data/BaselineData/S0_OB_Inputs/S0_NPV_ResRisk.csv')

##### Protection Costs (test three assumptions about protection costs) ####
S0_NPV_ProtectCost <- read.csv('./data/BaselineData/BeachNourishment/NPV_ProtectionCost.csv')


```


#S0: Analysis: 2% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost for S0 
#Estimating total Response and recovery cost for households (NA)
#Estimating total costs to repair roads (NA)

#Estimating total costs for beach nourishment (2021 event only: Muni + Fed)
S0_BeachRestore_dis2pct <-  S0_FedCostOB_NPV$dis2pct[S0_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S0_MuniCostOB_NPV$dis2pct[S0_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs
S0_NFIP_PolicyCost_dis2pct <- S0_FedRevOB_NPV$dis2pct[S0_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (post-Sandy only)
S0_NFIP_Claims_dis2pct <- S0_FedCostOB_NPV$dis2pct[S0_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]


# Create a vector of the object names
object_names <- c("S0_BeachRestore_dis2pct",
                  "S0_NFIP_PolicyCost_dis2pct",
                  "S0_NFIP_Claims_dis2pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S0_PolicyCost_dis2pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S0_PolicyCost_dis2pct

# Total Policy cost
S0_TotalPolicyCost_dis2pct <- sum(S0_PolicyCost_dis2pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S0_NPV_ProtectCost_V1_dis2pct <- S0_NPV_ProtectCost$dis2pct[S0_NPV_ProtectCost$Scenario=="V1"]
S0_NPV_ProtectCost_V2_dis2pct <- S0_NPV_ProtectCost$dis2pct[S0_NPV_ProtectCost$Scenario=="V2"]
S0_NPV_ProtectCost_V3_dis2pct <- S0_NPV_ProtectCost$dis2pct[S0_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S0_NPV_ResRisk_low_dis2pct <- S0_NPV_ResRisk$dis2pct[S0_NPV_ResRisk$Scenario=="Low"]
S0_NPV_ResRisk_mid_dis2pct <- S0_NPV_ResRisk$dis2pct[S0_NPV_ResRisk$Scenario=="Medium"]
S0_NPV_ResRisk_high_dis2pct <- S0_NPV_ResRisk$dis2pct[S0_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S0_VLow: Low future costs (low protection costs, low residual risk)
S0_VMid: Mid future costs (mid protection costs, mid residual risk)
S0_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S0_VLow_NPV_dis2pct <- S0_TotalPolicyCost_dis2pct + S0_NPV_ProtectCost_V1_dis2pct + 
  S0_NPV_ResRisk_low_dis2pct

S0_VMid_NPV_dis2pct <- S0_TotalPolicyCost_dis2pct + S0_NPV_ProtectCost_V2_dis2pct + 
  S0_NPV_ResRisk_mid_dis2pct

S0_VHigh_NPV_dis2pct <- S0_TotalPolicyCost_dis2pct + S0_NPV_ProtectCost_V3_dis2pct + 
  S0_NPV_ResRisk_high_dis2pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S0 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S0_TotalPolicyCost_dis2pct, 
           S0_NPV_ResRisk_low_dis2pct, 
           S0_NPV_ProtectCost_V1_dis2pct, 
           S0_TotalPolicyCost_dis2pct, 
           S0_NPV_ResRisk_mid_dis2pct, 
           S0_NPV_ProtectCost_V2_dis2pct,
           S0_TotalPolicyCost_dis2pct, 
           S0_NPV_ResRisk_high_dis2pct, 
           S0_NPV_ProtectCost_V3_dis2pct)
)

# Convert 'CostType' to a factor and reorder the levels
S0$Assumption <- factor(S0$Assumption, levels = c("Low", "Mid", "High"))
S0$CostType <- factor(S0$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S0, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S0: Cost Analysis by Assumptions \n(2% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 4000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)") 


```


#S0: Analysis: 3% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost for S0 
#Estimating total Response and recovery cost for households (NA)
#Estimating total costs to repair roads (NA)

#Estimating total costs for beach nourishment (2021 event only: Muni + Fed)
S0_BeachRestore_dis3pct <-  S0_FedCostOB_NPV$dis3pct[S0_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S0_MuniCostOB_NPV$dis3pct[S0_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs
S0_NFIP_PolicyCost_dis3pct <- S0_FedRevOB_NPV$dis3pct[S0_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (post-Sandy only)
S0_NFIP_Claims_dis3pct <- S0_FedCostOB_NPV$dis3pct[S0_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]


# Create a vector of the object names
object_names <- c("S0_BeachRestore_dis3pct",
                  "S0_NFIP_PolicyCost_dis3pct",
                  "S0_NFIP_Claims_dis3pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S0_PolicyCost_dis3pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S0_PolicyCost_dis3pct

# Total Policy cost
S0_TotalPolicyCost_dis3pct <- sum(S0_PolicyCost_dis3pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S0_NPV_ProtectCost_V1_dis3pct <- S0_NPV_ProtectCost$dis3pct[S0_NPV_ProtectCost$Scenario=="V1"]
S0_NPV_ProtectCost_V2_dis3pct <- S0_NPV_ProtectCost$dis3pct[S0_NPV_ProtectCost$Scenario=="V2"]
S0_NPV_ProtectCost_V3_dis3pct <- S0_NPV_ProtectCost$dis3pct[S0_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S0_NPV_ResRisk_low_dis3pct <- S0_NPV_ResRisk$dis3pct[S0_NPV_ResRisk$Scenario=="Low"]
S0_NPV_ResRisk_mid_dis3pct <- S0_NPV_ResRisk$dis3pct[S0_NPV_ResRisk$Scenario=="Medium"]
S0_NPV_ResRisk_high_dis3pct <- S0_NPV_ResRisk$dis3pct[S0_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S0_VLow: Low future costs (low protection costs, low residual risk)
S0_VMid: Mid future costs (mid protection costs, mid residual risk)
S0_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S0_VLow_NPV_dis3pct <- S0_TotalPolicyCost_dis3pct + S0_NPV_ProtectCost_V1_dis3pct + 
  S0_NPV_ResRisk_low_dis3pct

S0_VMid_NPV_dis3pct <- S0_TotalPolicyCost_dis3pct + S0_NPV_ProtectCost_V2_dis3pct + 
  S0_NPV_ResRisk_mid_dis3pct

S0_VHigh_NPV_dis3pct <- S0_TotalPolicyCost_dis3pct + S0_NPV_ProtectCost_V3_dis3pct + 
  S0_NPV_ResRisk_high_dis3pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S0 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S0_TotalPolicyCost_dis3pct, 
           S0_NPV_ResRisk_low_dis3pct, 
           S0_NPV_ProtectCost_V1_dis3pct, 
           S0_TotalPolicyCost_dis3pct, 
           S0_NPV_ResRisk_mid_dis3pct, 
           S0_NPV_ProtectCost_V2_dis3pct,
           S0_TotalPolicyCost_dis3pct, 
           S0_NPV_ResRisk_high_dis3pct, 
           S0_NPV_ProtectCost_V3_dis3pct)
)

# Convert 'CostType' to a factor and reorder the levels
S0$Assumption <- factor(S0$Assumption, levels = c("Low", "Mid", "High"))
S0$CostType <- factor(S0$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S0, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S0: Cost Analysis by Assumptions \n(3% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 4000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```


#S0: Analysis: 7% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost for S0 
#Estimating total Response and recovery cost for households (NA)
#Estimating total costs to repair roads (NA)

#Estimating total costs for beach nourishment (2021 event only: Muni + Fed)
S0_BeachRestore_dis7pct <-  S0_FedCostOB_NPV$dis7pct[S0_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S0_MuniCostOB_NPV$dis7pct[S0_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs
S0_NFIP_PolicyCost_dis7pct <- S0_FedRevOB_NPV$dis7pct[S0_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (post-Sandy only)
S0_NFIP_Claims_dis7pct <- S0_FedCostOB_NPV$dis7pct[S0_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]


# Create a vector of the object names
object_names <- c("S0_BeachRestore_dis7pct",
                  "S0_NFIP_PolicyCost_dis7pct",
                  "S0_NFIP_Claims_dis7pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S0_PolicyCost_dis7pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S0_PolicyCost_dis7pct

# Total Policy cost
S0_TotalPolicyCost_dis7pct <- sum(S0_PolicyCost_dis7pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S0_NPV_ProtectCost_V1_dis7pct <- S0_NPV_ProtectCost$dis7pct[S0_NPV_ProtectCost$Scenario=="V1"]
S0_NPV_ProtectCost_V2_dis7pct <- S0_NPV_ProtectCost$dis7pct[S0_NPV_ProtectCost$Scenario=="V2"]
S0_NPV_ProtectCost_V3_dis7pct <- S0_NPV_ProtectCost$dis7pct[S0_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S0_NPV_ResRisk_low_dis7pct <- S0_NPV_ResRisk$dis7pct[S0_NPV_ResRisk$Scenario=="Low"]
S0_NPV_ResRisk_mid_dis7pct <- S0_NPV_ResRisk$dis7pct[S0_NPV_ResRisk$Scenario=="Medium"]
S0_NPV_ResRisk_high_dis7pct <- S0_NPV_ResRisk$dis7pct[S0_NPV_ResRisk$Scenario=="High"]

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S0 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S0_TotalPolicyCost_dis7pct, 
           S0_NPV_ResRisk_low_dis7pct, 
           S0_NPV_ProtectCost_V1_dis7pct, 
           S0_TotalPolicyCost_dis7pct, 
           S0_NPV_ResRisk_mid_dis7pct, 
           S0_NPV_ProtectCost_V2_dis7pct,
           S0_TotalPolicyCost_dis7pct, 
           S0_NPV_ResRisk_high_dis7pct, 
           S0_NPV_ProtectCost_V3_dis7pct)
)

# Convert 'CostType' to a factor and reorder the levels
S0$Assumption <- factor(S0$Assumption, levels = c("Low", "Mid", "High"))
S0$CostType <- factor(S0$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S0, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S0: Cost Analysis by Assumptions \n(7% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 4000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```


#Generating comparison table
```{r include=FALSE}
# Define values for PolicyCost, ResidualRisk, and ProtectionCost for each assumption
S0_PolicyCost <- list(
  Low = c(S0_TotalPolicyCost_dis2pct, S0_TotalPolicyCost_dis3pct, S0_TotalPolicyCost_dis7pct),
  Mid = c(S0_TotalPolicyCost_dis2pct, S0_TotalPolicyCost_dis3pct, S0_TotalPolicyCost_dis7pct),
  High = c(S0_TotalPolicyCost_dis2pct, S0_TotalPolicyCost_dis3pct, S0_TotalPolicyCost_dis7pct)
)

S0_ResidualRisk <- list(
  Low = c(S0_NPV_ResRisk_low_dis2pct, S0_NPV_ResRisk_low_dis3pct, S0_NPV_ResRisk_low_dis7pct),
  Mid = c(S0_NPV_ResRisk_mid_dis2pct, S0_NPV_ResRisk_mid_dis3pct, S0_NPV_ResRisk_mid_dis7pct),
  High = c(S0_NPV_ResRisk_high_dis2pct, S0_NPV_ResRisk_high_dis3pct, S0_NPV_ResRisk_high_dis7pct)
)

S0_ProtectCost <- list(
  Low = c(S0_NPV_ProtectCost_V1_dis2pct, S0_NPV_ProtectCost_V1_dis3pct, S0_NPV_ProtectCost_V1_dis7pct),
  Mid = c(S0_NPV_ProtectCost_V2_dis2pct, S0_NPV_ProtectCost_V2_dis3pct, S0_NPV_ProtectCost_V2_dis7pct),
  High = c(S0_NPV_ProtectCost_V3_dis2pct, S0_NPV_ProtectCost_V3_dis3pct, S0_NPV_ProtectCost_V3_dis7pct)
)

# Combine into a single dataframe
S0_SummaryTable <- data.frame(
  Scenario = rep(c("S0"), times =9),
  Assumption = rep(c("Low", "Mid", "High"), each = 3),
  DiscountRate = rep(c("2%", "3%", "7%"), times = 3),
  PolicyCost = unlist(S0_PolicyCost),
  ResidualRisk = unlist(S0_ResidualRisk),
  ProtectionCost = unlist(S0_ProtectCost)
)

# Calculate the total cost for each row
S0_SummaryTable$TotalCost <- rowSums(S0_SummaryTable[, c("PolicyCost", "ResidualRisk", "ProtectionCost")])

# Display the combined table
rownames(S0_SummaryTable) <- NULL
print(S0_SummaryTable)

# Write out table
path <- './data/SummaryAndAnalysis'
write.csv(S0_SummaryTable, file.path(path, "S0_SummaryTable.csv"), row.names = FALSE)
```

