---
title: "Scenario 1 Analysis"
author: "Laura Geronimo"
date: "2024-09-16"
output: html_document
---

In this worksheet we analyse and compare the total costs to society for scenario 1 (Rebuild), using best estimates. We select retrospective policy cost and prospective protection and residual risk costs based on the scenario assumptions and discount rates. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)
library(ggplot2)
library(scales)
library(knitr)
#library(formattable)

options(scipen=999)


##Importing Data ####
####Retrospective ####

#S1
S1_FedCostOB_NPV <- read.csv('./data/BaselineData/S1_OB_Inputs/S1_FedCostOB_NPV.csv')
S1_FedRevOB_NPV <- read.csv('./data/BaselineData/S1_OB_Inputs/S1_FedRevOB_NPV.csv')
S1_MuniCostOB_NPV <- read.csv('./data/BaselineData/S1_OB_Inputs/S1_MuniCostOB_NPV.csv')
S1_MuniRevOB_NPV <- read.csv('./data/BaselineData/S1_OB_Inputs/S1_MuniRevOB_NPV.csv')


#hidden costs to households
Est_Dmg_Claims_Sandy_byScenario <- read.csv('./data/BaselineData/OB_SandyDmgsAndClaimsEstByScenario/Est_Dmg_Claims_Sandy_byScenario.csv')

####Prospective ####
##### Residual Risk ####
S1_NPV_ResRisk <- read.csv('./data/BaselineData/S1_OB_Inputs/S1_NPV_ResRisk.csv')


##### Protection Costs ####
S1_NPV_ProtectCost <- read.csv('./data/BaselineData/BeachNourishment/NPV_ProtectionCost.csv')


```


#S1: Analysis: 2% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S1_ResponseRecoveryMuni_dis2pct <-  S1_FedCostOB_NPV$dis2pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S1_MuniCostOB_NPV$dis2pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery"]

#Estimating total Response and recovery cost for households (Fed share)
S1_ResponseRecoveryHH_dis2pct <- S1_FedCostOB_NPV$dis2pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed)
S1_RepairRoadsBridges_dis2pct <-  S1_FedCostOB_NPV$dis2pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share)
S1_BeachRestore_dis2pct <-  S1_FedCostOB_NPV$dis2pct[S1_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S1_MuniCostOB_NPV$dis2pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs 
S1_NFIP_PolicyCost_dis2pct <- S1_FedRevOB_NPV$dis2pct[S1_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims

S1_NFIP_Claims_dis2pct <- S1_FedCostOB_NPV$dis2pct[S1_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims)
S1_HiddenHHCosts_dis2pct <- Est_Dmg_Claims_Sandy_byScenario$S1[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]

# Create a vector of the object names
object_names <- c("S1_ResponseRecoveryMuni_dis2pct", 
                  "S1_ResponseRecoveryHH_dis2pct", 
                  "S1_RepairRoadsBridges_dis2pct",
                  "S1_BeachRestore_dis2pct",
                  "S1_NFIP_PolicyCost_dis2pct",
                  "S1_NFIP_Claims_dis2pct",
                  "S1_HiddenHHCosts_dis2pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S1_PolicyCost_dis2pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S1_PolicyCost_dis2pct

# Total Policy cost
S1_TotalPolicyCost_dis2pct <- sum(S1_PolicyCost_dis2pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S1_NPV_ProtectCost_V1_dis2pct <- S1_NPV_ProtectCost$dis2pct[S1_NPV_ProtectCost$Scenario=="V1"]
S1_NPV_ProtectCost_V2_dis2pct <- S1_NPV_ProtectCost$dis2pct[S1_NPV_ProtectCost$Scenario=="V2"]
S1_NPV_ProtectCost_V3_dis2pct <- S1_NPV_ProtectCost$dis2pct[S1_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S1_NPV_ResRisk_low_dis2pct <- S1_NPV_ResRisk$dis2pct[S1_NPV_ResRisk$Scenario=="Low"]
S1_NPV_ResRisk_mid_dis2pct <- S1_NPV_ResRisk$dis2pct[S1_NPV_ResRisk$Scenario=="Medium"]
S1_NPV_ResRisk_high_dis2pct <- S1_NPV_ResRisk$dis2pct[S1_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S1_VLow: Low future costs (low protection costs, low residual risk)
S1_VMid: Mid future costs (mid protection costs, mid residual risk)
S1_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S1_VLow_NPV_dis2pct <- S1_TotalPolicyCost_dis2pct + S1_NPV_ProtectCost_V1_dis2pct + 
  S1_NPV_ResRisk_low_dis2pct

S1_VMid_NPV_dis2pct <- S1_TotalPolicyCost_dis2pct + S1_NPV_ProtectCost_V2_dis2pct + 
  S1_NPV_ResRisk_mid_dis2pct

S1_VHigh_NPV_dis2pct <- S1_TotalPolicyCost_dis2pct + S1_NPV_ProtectCost_V3_dis2pct + 
  S1_NPV_ResRisk_high_dis2pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S1 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S1_TotalPolicyCost_dis2pct, 
           S1_NPV_ResRisk_low_dis2pct, 
           S1_NPV_ProtectCost_V1_dis2pct, 
           S1_TotalPolicyCost_dis2pct, 
           S1_NPV_ResRisk_mid_dis2pct, 
           S1_NPV_ProtectCost_V2_dis2pct,
           S1_TotalPolicyCost_dis2pct, 
           S1_NPV_ResRisk_high_dis2pct, 
           S1_NPV_ProtectCost_V3_dis2pct)
)

# Convert 'CostType' to a factor and reorder the levels
S1$Assumption <- factor(S1$Assumption, levels = c("Low", "Mid", "High"))
S1$CostType <- factor(S1$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S1, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S1: Cost Analysis by Assumptions \n(2% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```


#S1: Analysis: 3% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S1_ResponseRecoveryMuni_dis3pct <-  S1_FedCostOB_NPV$dis3pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S1_MuniCostOB_NPV$dis3pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery"]

#Estimating total Response and recovery cost for households (Fed share)
S1_ResponseRecoveryHH_dis3pct <- S1_FedCostOB_NPV$dis3pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed)
S1_RepairRoadsBridges_dis3pct <-  S1_FedCostOB_NPV$dis3pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share)
S1_BeachRestore_dis3pct <-  S1_FedCostOB_NPV$dis3pct[S1_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S1_MuniCostOB_NPV$dis3pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs 
S1_NFIP_PolicyCost_dis3pct <- S1_FedRevOB_NPV$dis3pct[S1_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims

S1_NFIP_Claims_dis3pct <- S1_FedCostOB_NPV$dis3pct[S1_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims)
S1_HiddenHHCosts_dis3pct <- Est_Dmg_Claims_Sandy_byScenario$S1[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]

# Create a vector of the object names
object_names <- c("S1_ResponseRecoveryMuni_dis3pct", 
                  "S1_ResponseRecoveryHH_dis3pct", 
                  "S1_RepairRoadsBridges_dis3pct",
                  "S1_BeachRestore_dis3pct",
                  "S1_NFIP_PolicyCost_dis3pct",
                  "S1_NFIP_Claims_dis3pct",
                  "S1_HiddenHHCosts_dis3pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S1_PolicyCost_dis3pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S1_PolicyCost_dis3pct

# Total Policy cost
S1_TotalPolicyCost_dis3pct <- sum(S1_PolicyCost_dis3pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S1_NPV_ProtectCost_V1_dis3pct <- S1_NPV_ProtectCost$dis3pct[S1_NPV_ProtectCost$Scenario=="V1"]
S1_NPV_ProtectCost_V2_dis3pct <- S1_NPV_ProtectCost$dis3pct[S1_NPV_ProtectCost$Scenario=="V2"]
S1_NPV_ProtectCost_V3_dis3pct <- S1_NPV_ProtectCost$dis3pct[S1_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S1_NPV_ResRisk_low_dis3pct <- S1_NPV_ResRisk$dis3pct[S1_NPV_ResRisk$Scenario=="Low"]
S1_NPV_ResRisk_mid_dis3pct <- S1_NPV_ResRisk$dis3pct[S1_NPV_ResRisk$Scenario=="Medium"]
S1_NPV_ResRisk_high_dis3pct <- S1_NPV_ResRisk$dis3pct[S1_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S1_VLow: Low future costs (low protection costs, low residual risk)
S1_VMid: Mid future costs (mid protection costs, mid residual risk)
S1_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S1_VLow_NPV_dis3pct <- S1_TotalPolicyCost_dis3pct + S1_NPV_ProtectCost_V1_dis3pct + 
  S1_NPV_ResRisk_low_dis3pct

S1_VMid_NPV_dis3pct <- S1_TotalPolicyCost_dis3pct + S1_NPV_ProtectCost_V2_dis3pct + 
  S1_NPV_ResRisk_mid_dis3pct

S1_VHigh_NPV_dis3pct <- S1_TotalPolicyCost_dis3pct + S1_NPV_ProtectCost_V3_dis3pct + 
  S1_NPV_ResRisk_high_dis3pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S1 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S1_TotalPolicyCost_dis3pct, 
           S1_NPV_ResRisk_low_dis3pct, 
           S1_NPV_ProtectCost_V1_dis3pct, 
           S1_TotalPolicyCost_dis3pct, 
           S1_NPV_ResRisk_mid_dis3pct, 
           S1_NPV_ProtectCost_V2_dis3pct,
           S1_TotalPolicyCost_dis3pct, 
           S1_NPV_ResRisk_high_dis3pct, 
           S1_NPV_ProtectCost_V3_dis3pct)
)

# Convert 'CostType' to a factor and reorder the levels
S1$Assumption <- factor(S1$Assumption, levels = c("Low", "Mid", "High"))
S1$CostType <- factor(S1$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S1, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S1: Cost Analysis by Assumptions \n(3% Discount Rate)", x = "Assumption", y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", "Residual Risk" = "red", "Protection Cost" = "black")) +
  theme_minimal()+
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```


#S1: Analysis: 7% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S1_ResponseRecoveryMuni_dis7pct <-  S1_FedCostOB_NPV$dis7pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S1_MuniCostOB_NPV$dis7pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery"]

#Estimating total Response and recovery cost for households (Fed share)
S1_ResponseRecoveryHH_dis7pct <- S1_FedCostOB_NPV$dis7pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed)
S1_RepairRoadsBridges_dis7pct <-  S1_FedCostOB_NPV$dis7pct[S1_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share)
S1_BeachRestore_dis7pct <-  S1_FedCostOB_NPV$dis7pct[S1_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S1_MuniCostOB_NPV$dis7pct[S1_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs 
S1_NFIP_PolicyCost_dis7pct <- S1_FedRevOB_NPV$dis7pct[S1_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims

S1_NFIP_Claims_dis7pct <- S1_FedCostOB_NPV$dis7pct[S1_FedCostOB_NPV$variable=="OB_FedShare_NFIP"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims)
S1_HiddenHHCosts_dis7pct <- Est_Dmg_Claims_Sandy_byScenario$S1[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]

# Create a vector of the object names
object_names <- c("S1_ResponseRecoveryMuni_dis7pct", 
                  "S1_ResponseRecoveryHH_dis7pct", 
                  "S1_RepairRoadsBridges_dis7pct",
                  "S1_BeachRestore_dis7pct",
                  "S1_NFIP_PolicyCost_dis7pct",
                  "S1_NFIP_Claims_dis7pct",
                  "S1_HiddenHHCosts_dis7pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S1_PolicyCost_dis7pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S1_PolicyCost_dis7pct

# Total Policy cost
S1_TotalPolicyCost_dis7pct <- sum(S1_PolicyCost_dis7pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S1_NPV_ProtectCost_V1_dis7pct <- S1_NPV_ProtectCost$dis7pct[S1_NPV_ProtectCost$Scenario=="V1"]
S1_NPV_ProtectCost_V2_dis7pct <- S1_NPV_ProtectCost$dis7pct[S1_NPV_ProtectCost$Scenario=="V2"]
S1_NPV_ProtectCost_V3_dis7pct <- S1_NPV_ProtectCost$dis7pct[S1_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S1_NPV_ResRisk_low_dis7pct <- S1_NPV_ResRisk$dis7pct[S1_NPV_ResRisk$Scenario=="Low"]
S1_NPV_ResRisk_mid_dis7pct <- S1_NPV_ResRisk$dis7pct[S1_NPV_ResRisk$Scenario=="Medium"]
S1_NPV_ResRisk_high_dis7pct <- S1_NPV_ResRisk$dis7pct[S1_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S1_VLow: Low future costs (low protection costs, low residual risk)
S1_VMid: Mid future costs (mid protection costs, mid residual risk)
S1_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S1_VLow_NPV_dis7pct <- S1_TotalPolicyCost_dis7pct + S1_NPV_ProtectCost_V1_dis7pct + 
  S1_NPV_ResRisk_low_dis7pct

S1_VMid_NPV_dis7pct <- S1_TotalPolicyCost_dis7pct + S1_NPV_ProtectCost_V2_dis7pct + 
  S1_NPV_ResRisk_mid_dis7pct

S1_VHigh_NPV_dis7pct <- S1_TotalPolicyCost_dis7pct + S1_NPV_ProtectCost_V3_dis7pct + 
  S1_NPV_ResRisk_high_dis7pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S1 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S1_TotalPolicyCost_dis7pct, 
           S1_NPV_ResRisk_low_dis7pct, 
           S1_NPV_ProtectCost_V1_dis7pct, 
           S1_TotalPolicyCost_dis7pct, 
           S1_NPV_ResRisk_mid_dis7pct, 
           S1_NPV_ProtectCost_V2_dis7pct,
           S1_TotalPolicyCost_dis7pct, 
           S1_NPV_ResRisk_high_dis7pct, 
           S1_NPV_ProtectCost_V3_dis7pct)
)

# Convert 'CostType' to a factor and reorder the levels
S1$Assumption <- factor(S1$Assumption, levels = c("Low", "Mid", "High"))
S1$CostType <- factor(S1$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S1, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S1: Cost Analysis by Assumptions \n(7% Discount Rate)", x = "Assumption", y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", "Residual Risk" = "red", "Protection Cost" = "black")) +
  theme_minimal()+
    scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1000000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```

#Generating comparison table
```{r include=FALSE}
# Define values for PolicyCost, ResidualRisk, and ProtectionCost for each assumption
S1_PolicyCost <- list(
  Low = c(S1_TotalPolicyCost_dis2pct, S1_TotalPolicyCost_dis3pct, S1_TotalPolicyCost_dis7pct),
  Mid = c(S1_TotalPolicyCost_dis2pct, S1_TotalPolicyCost_dis3pct, S1_TotalPolicyCost_dis7pct),
  High = c(S1_TotalPolicyCost_dis2pct, S1_TotalPolicyCost_dis3pct, S1_TotalPolicyCost_dis7pct)
)

S1_ResidualRisk <- list(
  Low = c(S1_NPV_ResRisk_low_dis2pct, S1_NPV_ResRisk_low_dis3pct, S1_NPV_ResRisk_low_dis7pct),
  Mid = c(S1_NPV_ResRisk_mid_dis2pct, S1_NPV_ResRisk_mid_dis3pct, S1_NPV_ResRisk_mid_dis7pct),
  High = c(S1_NPV_ResRisk_high_dis2pct, S1_NPV_ResRisk_high_dis3pct, S1_NPV_ResRisk_high_dis7pct)
)

S1_ProtectCost <- list(
  Low = c(S1_NPV_ProtectCost_V1_dis2pct, S1_NPV_ProtectCost_V1_dis3pct, S1_NPV_ProtectCost_V1_dis7pct),
  Mid = c(S1_NPV_ProtectCost_V2_dis2pct, S1_NPV_ProtectCost_V2_dis3pct, S1_NPV_ProtectCost_V2_dis7pct),
  High = c(S1_NPV_ProtectCost_V3_dis2pct, S1_NPV_ProtectCost_V3_dis3pct, S1_NPV_ProtectCost_V3_dis7pct)
)

# Combine into a single dataframe
S1_SummaryTable <- data.frame(
  Scenario = rep(c("S1"), times =9),
  Assumption = rep(c("Low", "Mid", "High"), each = 3),
  DiscountRate = rep(c("2%", "3%", "7%"), times = 3),
  PolicyCost = unlist(S1_PolicyCost),
  ResidualRisk = unlist(S1_ResidualRisk),
  ProtectionCost = unlist(S1_ProtectCost)
)

# Calculate the total cost for each row
S1_SummaryTable$TotalCost <- rowSums(S1_SummaryTable[, c("PolicyCost", "ResidualRisk", "ProtectionCost")])

# Display the combined table
rownames(S1_SummaryTable) <- NULL
print(S1_SummaryTable)

# Write out table
path <- './data/SummaryAndAnalysis'
write.csv(S1_SummaryTable, file.path(path, "S1_SummaryTable.csv"), row.names = FALSE)
```


#obtaining ratio of federal to municipal expenses
```{r include = FALSE}

#2pct
TotalFedCost_dis2pct <- sum(S1_FedCostOB_NPV$dis2pct)
TotalMuniCost_dis2pct <- sum(S1_MuniCostOB_NPV$dis2pct)
RatioFedMuni_dis2pct <- TotalFedCost_dis2pct / TotalMuniCost_dis2pct 

#3pct
TotalFedCost_dis3pct <- sum(S1_FedCostOB_NPV$dis3pct)
TotalMuniCost_dis3pct <- sum(S1_MuniCostOB_NPV$dis3pct)
RatioFedMuni_dis3pct <- TotalFedCost_dis3pct / TotalMuniCost_dis3pct

#7pct
TotalFedCost_dis7pct <- sum(S1_FedCostOB_NPV$dis7pct)
TotalMuniCost_dis7pct <- sum(S1_MuniCostOB_NPV$dis7pct)
RatioFedMuni_dis7pct <- TotalFedCost_dis7pct / TotalMuniCost_dis7pct
```