---
title: "Scenario 2 Analysis"
author: "Laura Geronimo"
date: "2024-09-16"
output: html_document
---

In this worksheet we analyse and compare the total costs to society Scenario 1 across different assumptions, using best estimates.

Recall the assumptions for S2 applied in prior worksheets:

Assumptions:

Federal policy costs:
-Federal Cost Share for Municipal Response and Recovery (essential only for 2012 and 2013)
    -PA Debris Removal (2012, 2013)
    -PA Protective measures (2012, 2013)
    -Repairs to Route 37 (2012, 2013)
-Federal cost share for response and recovery direct to household (NA: redirected to buyouts)
-Federal IA / NFIP claims (NA: redirected to buyouts)
-Federal cost share for beach nourishment (NA)
-Federal cost share for buyouts = 75% of market price for all homes
-No NFIP premiums after 2012 because all homes have been relocated

Municipal policy costs:
-Municipal cost share for federal response and recovery (essential only for 2012 and 2013):
    -From PA, retaining only debris removal and protective measures for 2012 and 2013
    -NA on HMPG for Ortley Beach
-Municipal cost share for beach nourishment (drops by 100%)NA
-Municipal cost share for buyouts = 25% of market price for all homes

Since all homes are relocated, we assume not residual risk or protection costs


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)
library(ggplot2)
library(scales)
library(knitr)
#library(formattable)

options(scipen=999)


##Importing Data ####
####Retrospective ####

#S2
S2_FedCostOB_NPV <- read.csv('./data/BaselineData/S2_OB_Inputs/S2_FedCostOB_NPV.csv')
S2_FedRevOB_NPV <- read.csv('./data/BaselineData/S2_OB_Inputs/S2_FedRevOB_NPV.csv')
S2_MuniCostOB_NPV <- read.csv('./data/BaselineData/S2_OB_Inputs/S2_MuniCostOB_NPV.csv')
S2_MuniRevOB_NPV <- read.csv('./data/BaselineData/S2_OB_Inputs/S2_MuniRevOB_NPV.csv')


```


#S2: Analysis: 2% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimate of Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 

#Estimate of total Response and recovery cost for municipality (Fed + Muni Share: essential only for 2012 and 2013)
S2_ResponseRecoveryMuni_dis2pct <-
  S2_FedCostOB_NPV$dis2pct[S2_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S2_MuniCostOB_NPV$dis2pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"] 

#Estimate of total Response and recovery cost for households (NA - redirected to buyouts)

#Estimate of total costs to repair roads (Fed)
S2_RepairRoadsBridges_dis2pct <-
  S2_FedCostOB_NPV$dis2pct[S2_FedCostOB_NPV$variable=="OB_FedShare_DOT_Rt37_adj"] 

#Estimate of total costs for beach nourishment (NA)

#Estimate of NFIP policy costs (2012 only)
S2_NFIP_PolicyCost_dis2pct <-
  S2_FedRevOB_NPV$dis2pct[S2_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimate of NFIP claims (NA - redirected to buyout)

#Estimate of 'hidden' costs of rebuilding to households (NA, no rebuilding)

#Estimate of buyout costs (Fed + Muni share)
S2_BuyoutCost_dis2pct <- 
  S2_FedCostOB_NPV$dis2pct[S2_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"] +
  S2_MuniCostOB_NPV$dis2pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S2_ResponseRecoveryMuni_dis2pct", 
                  "S2_RepairRoadsBridges_dis2pct",
                  "S2_NFIP_PolicyCost_dis2pct",
                  "S2_BuyoutCost_dis2pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S2_PolicyCost_dis2pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S2_PolicyCost_dis2pct

# Total Policy cost
S2_TotalPolicyCost_dis2pct <- sum(S2_PolicyCost_dis2pct$Value)
```

###Prospective costs (2022:2063) 
####Protection Costs Assumption - NA
####Residual Risk Assumption  - NA

#S2: Analysis: 3% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimate of Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 

#Estimate of total Response and recovery cost for municipality (Fed + Muni Share - (essential only for 2012 and 2013))
S2_ResponseRecoveryMuni_dis3pct <-
  S2_FedCostOB_NPV$dis3pct[S2_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S2_MuniCostOB_NPV$dis3pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"] 

#Estimate of total Response and recovery cost for households (NA - redirected to buyouts)

#Estimate of total costs to repair roads (Fed)
S2_RepairRoadsBridges_dis3pct <-
  S2_FedCostOB_NPV$dis3pct[S2_FedCostOB_NPV$variable=="OB_FedShare_DOT_Rt37_adj"] 

#Estimate of total costs for beach nourishment (NA)

#Estimate of NFIP policy costs (2012 only)
S2_NFIP_PolicyCost_dis3pct <-
  S2_FedRevOB_NPV$dis3pct[S2_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimate of NFIP claims (NA - redirected to buyout)

#Estimate of 'hidden' costs of rebuilding to households (NA, no rebuilding)

#Estimate of buyout costs (Fed + Muni share)
S2_BuyoutCost_dis3pct <- 
  S2_FedCostOB_NPV$dis3pct[S2_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"] +
  S2_MuniCostOB_NPV$dis3pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S2_ResponseRecoveryMuni_dis3pct", 
                  "S2_RepairRoadsBridges_dis3pct",
                  "S2_NFIP_PolicyCost_dis3pct",
                  "S2_BuyoutCost_dis3pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S2_PolicyCost_dis3pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S2_PolicyCost_dis3pct

# Total Policy cost
S2_TotalPolicyCost_dis3pct <- sum(S2_PolicyCost_dis3pct$Value)
```

###Prospective costs (2022:2063) 
####Protection Costs Assumption - NA
####Residual Risk Assumption  - NA

#S2: Analysis: 7% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimate of Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 

#Estimate of total Response and recovery cost for municipality (Fed + Muni Share - (essential only for 2012 and 2013))
S2_ResponseRecoveryMuni_dis7pct <-
  S2_FedCostOB_NPV$dis7pct[S2_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S2_MuniCostOB_NPV$dis7pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"] 

#Estimate of total Response and recovery cost for households (NA - redirected to buyouts)

#Estimate of total costs to repair roads (Fed)
S2_RepairRoadsBridges_dis7pct <-
  S2_FedCostOB_NPV$dis7pct[S2_FedCostOB_NPV$variable=="OB_FedShare_DOT_Rt37_adj"] 

#Estimate of total costs for beach nourishment (NA)

#Estimate of NFIP policy costs (2012 only)
S2_NFIP_PolicyCost_dis7pct <-
  S2_FedRevOB_NPV$dis7pct[S2_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimate of NFIP claims (NA - redirected to buyout)

#Estimate of 'hidden' costs of rebuilding to households (NA, no rebuilding)

#Estimate of buyout costs (Fed + Muni share)
S2_BuyoutCost_dis7pct <- 
  S2_FedCostOB_NPV$dis7pct[S2_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"] +
  S2_MuniCostOB_NPV$dis7pct[S2_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S2_ResponseRecoveryMuni_dis7pct", 
                  "S2_RepairRoadsBridges_dis7pct",
                  "S2_NFIP_PolicyCost_dis7pct",
                  "S2_BuyoutCost_dis7pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S2_PolicyCost_dis7pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S2_PolicyCost_dis7pct

# Total Policy cost
S2_TotalPolicyCost_dis7pct <- sum(S2_PolicyCost_dis7pct$Value)
```

###Prospective costs (2022:2063) 
####Protection Costs Assumption - NA
####Residual Risk Assumption  - NA


#Generating comparison table
#Generating comparison table
```{r include=FALSE}
# Define values for PolicyCost, ResidualRisk, and ProtectionCost for each assumption
S2_PolicyCost <- list(
  Low = c(S2_TotalPolicyCost_dis2pct, S2_TotalPolicyCost_dis3pct, S2_TotalPolicyCost_dis7pct),
  Mid = c(S2_TotalPolicyCost_dis2pct, S2_TotalPolicyCost_dis3pct, S2_TotalPolicyCost_dis7pct),
  High = c(S2_TotalPolicyCost_dis2pct, S2_TotalPolicyCost_dis3pct, S2_TotalPolicyCost_dis7pct)
)

S2_ResidualRisk <- list(
  Low = c(0,0,0),
  Mid = c(0,0,0),
  High = c(0,0,0)
)

S2_ProtectCost <- list(
  Low = c(0,0,0),
  Mid = c(0,0,0),
  High = c(0,0,0)
)

# Combine into a single dataframe
S2_SummaryTable <- data.frame(
  Scenario = rep(c("S2"), times =9),
  Assumption = rep(c("Low", "Mid", "High"), each = 3),
  DiscountRate = rep(c("2%", "3%", "7%"), times = 3),
  PolicyCost = unlist(S2_PolicyCost),
  ResidualRisk = unlist(S2_ResidualRisk),
  ProtectionCost = unlist(S2_ProtectCost)
)

# Calculate the total cost for each row
S2_SummaryTable$TotalCost <- rowSums(S2_SummaryTable[, c("PolicyCost", "ResidualRisk", "ProtectionCost")])

# Display the combined table
rownames(S2_SummaryTable) <- NULL
print(S2_SummaryTable)

# Write out table
path <- './data/SummaryAndAnalysis'
write.csv(S2_SummaryTable, file.path(path, "S2_SummaryTable.csv"), row.names = FALSE)
```


