---
title: "Scenario 4 Analysis"
author: "Laura Geronimo"
date: "2024-09-16"
output: html_document
---

In this worksheet we analyse and compare the total costs to society Scenario 4 across different assumptions, using best estimates.

Recall the assumptions for S4 applied in prior worksheets:


Federal policy costs:
-Federal Cost Share for Municipal Response and Recovery (assume sticky)
-Federal cost share for response and recovery direct to household (50% less due to proportion redirected to buyouts)
-Federal cost share for roads (assume sticky)
-Federal IA / NFIP claims (50% less due to proportion redirected to buyouts)
-Federal cost share for beach nourishment (Initial repairs double due to widening beach)
-Federal cost share for buyouts = 75% of market price for all homes
-No NFIP premiums after 2012 because all homes have been relocated

Municipal policy costs:
-Municipal cost share for federal response and recovery (assume sticky)
-Municipal cost share for beach nourishment (Initial repairs double due to widening beach)
-Municipal cost share for buyouts = 25% of market price for 921 buyouts

Hidden costs to households (50% less than S1)


Assumptions:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)
library(ggplot2)
library(scales)
library(knitr)
library(formattable)

options(scipen=999)


##Importing Data ####
####Retrospective ####

#S4
S4_FedCostOB_NPV <- read.csv('./data/BaselineData/S4_OB_Inputs/S4_FedCostOB_NPV.csv')
S4_FedRevOB_NPV <- read.csv('./data/BaselineData/S4_OB_Inputs/S4_FedRevOB_NPV.csv')
S4_MuniCostOB_NPV <- read.csv('./data/BaselineData/S4_OB_Inputs/S4_MuniCostOB_NPV.csv')
S4_MuniRevOB_NPV <- read.csv('./data/BaselineData/S4_OB_Inputs/S4_MuniRevOB_NPV.csv')


#hidden costs to households
Est_Dmg_Claims_Sandy_byScenario <- read.csv('./data/BaselineData/OB_SandyDmgsAndClaimsEstByScenario/Est_Dmg_Claims_Sandy_byScenario.csv')

#Estimating buyout costs (Fed + muni share)
S4_BuyoutCosts_dis7pct <- S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"]  + S4_MuniCostOB_NPV$dis7pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

####Prospective ####
##### Residual Risk ####
S4_NPV_ResRisk <- read.csv('./data/BaselineData/S4_OB_Inputs/S4_NPV_ResRisk.csv')

##### Protection Costs ####
S4_NPV_ProtectCost <- read.csv('./data/BaselineData/BeachNourishment/NPV_ProtectionCost.csv')


```


#S4: Analysis: 2% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S4_ResponseRecoveryMuni_dis2pct <-  S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S4_MuniCostOB_NPV$dis2pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"]

#Estimating total Response and recovery cost for households (Fed share: 23% less due to homes relocating)
S4_ResponseRecoveryHH_dis2pct <- S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed: assume sticky)
S4_RepairRoadsBridges_dis2pct <-  S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share: initial fill doubles due to widened beach)
S4_BeachRestore_dis2pct <-  S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S4_MuniCostOB_NPV$dis2pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs (23% less after 2012 due to relocating homes)
S4_NFIP_PolicyCost_dis2pct <- S4_FedRevOB_NPV$dis2pct[S4_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (23% less due to relocating homes)
S4_NFIP_Claims_dis2pct <- S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_NFIP_Claims_adj"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims for remaining households that rebuild)
S4_HiddenHHCosts_dis2pct <- Est_Dmg_Claims_Sandy_byScenario$S4[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]


#Estimating buyout costs (Fed + muni share)
S4_BuyoutCosts_dis2pct <- S4_FedCostOB_NPV$dis2pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"]  + S4_MuniCostOB_NPV$dis2pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S4_ResponseRecoveryMuni_dis2pct", 
                  "S4_ResponseRecoveryHH_dis2pct", 
                  "S4_RepairRoadsBridges_dis2pct",
                  "S4_BeachRestore_dis2pct",
                  "S4_NFIP_PolicyCost_dis2pct",
                  "S4_NFIP_Claims_dis2pct",
                  "S4_HiddenHHCosts_dis2pct",
                  "S4_BuyoutCosts_dis2pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S4_PolicyCost_dis2pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S4_PolicyCost_dis2pct

# Total Policy cost
S4_TotalPolicyCost_dis2pct <- sum(S4_PolicyCost_dis2pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S4_NPV_ProtectCost_V1_dis2pct <- S4_NPV_ProtectCost$dis2pct[S4_NPV_ProtectCost$Scenario=="V1"]
S4_NPV_ProtectCost_V2_dis2pct <- S4_NPV_ProtectCost$dis2pct[S4_NPV_ProtectCost$Scenario=="V2"]
S4_NPV_ProtectCost_V3_dis2pct <- S4_NPV_ProtectCost$dis2pct[S4_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S4_NPV_ResRisk_low_dis2pct <- S4_NPV_ResRisk$dis2pct[S4_NPV_ResRisk$Scenario=="Low"]
S4_NPV_ResRisk_mid_dis2pct <- S4_NPV_ResRisk$dis2pct[S4_NPV_ResRisk$Scenario=="Medium"]
S4_NPV_ResRisk_high_dis2pct <- S4_NPV_ResRisk$dis2pct[S4_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S4_VLow: Low future costs (low protection costs, low residual risk)
S4_VMid: Mid future costs (mid protection costs, mid residual risk)
S4_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S4_VLow_NPV_dis2pct <- S4_TotalPolicyCost_dis2pct + S4_NPV_ProtectCost_V1_dis2pct + 
  S4_NPV_ResRisk_low_dis2pct

S4_VMid_NPV_dis2pct <- S4_TotalPolicyCost_dis2pct + S4_NPV_ProtectCost_V2_dis2pct + 
  S4_NPV_ResRisk_mid_dis2pct

S4_VHigh_NPV_dis2pct <- S4_TotalPolicyCost_dis2pct + S4_NPV_ProtectCost_V3_dis2pct + 
  S4_NPV_ResRisk_high_dis2pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S4 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S4_TotalPolicyCost_dis2pct, 
           S4_NPV_ResRisk_low_dis2pct, 
           S4_NPV_ProtectCost_V1_dis2pct, 
           S4_TotalPolicyCost_dis2pct, 
           S4_NPV_ResRisk_mid_dis2pct, 
           S4_NPV_ProtectCost_V2_dis2pct,
           S4_TotalPolicyCost_dis2pct, 
           S4_NPV_ResRisk_high_dis2pct, 
           S4_NPV_ProtectCost_V3_dis2pct)
)

# Convert 'CostType' to a factor and reorder the levels
S4$Assumption <- factor(S4$Assumption, levels = c("Low", "Mid", "High"))
S4$CostType <- factor(S4$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S4, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S4: Cost Analysis by Assumptions \n(2% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1250000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```



#S4: Analysis: 3% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S4_ResponseRecoveryMuni_dis3pct <-  S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S4_MuniCostOB_NPV$dis3pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"]

#Estimating total Response and recovery cost for households (Fed share: 23% less due to homes relocating)
S4_ResponseRecoveryHH_dis3pct <- S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed: assume sticky)
S4_RepairRoadsBridges_dis3pct <-  S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share: initial fill doubles due to widened beach)
S4_BeachRestore_dis3pct <-  S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S4_MuniCostOB_NPV$dis3pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs (23% less after 2012 due to relocating homes)
S4_NFIP_PolicyCost_dis3pct <- S4_FedRevOB_NPV$dis3pct[S4_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (23% less due to relocating homes)
S4_NFIP_Claims_dis3pct <- S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_NFIP_Claims_adj"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims for remaining households that rebuild)
S4_HiddenHHCosts_dis3pct <- Est_Dmg_Claims_Sandy_byScenario$S4[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]

#Estimating buyout costs (Fed + muni share)
S4_BuyoutCosts_dis3pct <- S4_FedCostOB_NPV$dis3pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"]  + S4_MuniCostOB_NPV$dis3pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S4_ResponseRecoveryMuni_dis3pct", 
                  "S4_ResponseRecoveryHH_dis3pct", 
                  "S4_RepairRoadsBridges_dis3pct",
                  "S4_BeachRestore_dis3pct",
                  "S4_NFIP_PolicyCost_dis3pct",
                  "S4_NFIP_Claims_dis3pct",
                  "S4_HiddenHHCosts_dis3pct",
                  "S4_BuyoutCosts_dis3pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S4_PolicyCost_dis3pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S4_PolicyCost_dis3pct

# Total Policy cost
S4_TotalPolicyCost_dis3pct <- sum(S4_PolicyCost_dis3pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S4_NPV_ProtectCost_V1_dis3pct <- S4_NPV_ProtectCost$dis3pct[S4_NPV_ProtectCost$Scenario=="V1"]
S4_NPV_ProtectCost_V2_dis3pct <- S4_NPV_ProtectCost$dis3pct[S4_NPV_ProtectCost$Scenario=="V2"]
S4_NPV_ProtectCost_V3_dis3pct <- S4_NPV_ProtectCost$dis3pct[S4_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S4_NPV_ResRisk_low_dis3pct <- S4_NPV_ResRisk$dis3pct[S4_NPV_ResRisk$Scenario=="Low"]
S4_NPV_ResRisk_mid_dis3pct <- S4_NPV_ResRisk$dis3pct[S4_NPV_ResRisk$Scenario=="Medium"]
S4_NPV_ResRisk_high_dis3pct <- S4_NPV_ResRisk$dis3pct[S4_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S4_VLow: Low future costs (low protection costs, low residual risk)
S4_VMid: Mid future costs (mid protection costs, mid residual risk)
S4_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S4_VLow_NPV_dis3pct <- S4_TotalPolicyCost_dis3pct + S4_NPV_ProtectCost_V1_dis3pct + 
  S4_NPV_ResRisk_low_dis3pct

S4_VMid_NPV_dis3pct <- S4_TotalPolicyCost_dis3pct + S4_NPV_ProtectCost_V2_dis3pct + 
  S4_NPV_ResRisk_mid_dis3pct

S4_VHigh_NPV_dis3pct <- S4_TotalPolicyCost_dis3pct + S4_NPV_ProtectCost_V3_dis3pct + 
  S4_NPV_ResRisk_high_dis3pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S4 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S4_TotalPolicyCost_dis3pct, 
           S4_NPV_ResRisk_low_dis3pct, 
           S4_NPV_ProtectCost_V1_dis3pct, 
           S4_TotalPolicyCost_dis3pct, 
           S4_NPV_ResRisk_mid_dis3pct, 
           S4_NPV_ProtectCost_V2_dis3pct,
           S4_TotalPolicyCost_dis3pct, 
           S4_NPV_ResRisk_high_dis3pct, 
           S4_NPV_ProtectCost_V3_dis3pct)
)

# Convert 'CostType' to a factor and reorder the levels
S4$Assumption <- factor(S4$Assumption, levels = c("Low", "Mid", "High"))
S4$CostType <- factor(S4$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S4, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S4: Cost Analysis by Assumptions \n(3% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1250000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```




#S4: Analysis: 7% discount rate 
###Total policy costs (Retrospective: 2012-2022)
```{r inclue= FALSE}
## Estimating Retrospective Policy Cost to Rebuild Ortley Beach 2012-2022 
#Estimating total Response and recovery cost for municipality (Fed + Muni Share)
S4_ResponseRecoveryMuni_dis7pct <-  S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToTR_ResponseRecovery"] +
  S4_MuniCostOB_NPV$dis7pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_ResponseRecovery_adj"]

#Estimating total Response and recovery cost for households (Fed share: 23% less due to homes relocating)
S4_ResponseRecoveryHH_dis7pct <- S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToIndiv_ResponseRecovery"]

#Estimating total costs to repair roads (Fed: assume sticky)
S4_RepairRoadsBridges_dis7pct <-  S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_ToRoadsBridge"] 

#Estimating total costs for beach nourishment (Fed + Muni Share: initial fill doubles due to widened beach)
S4_BeachRestore_dis7pct <-  S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BeachNourish_adj"] +
  S4_MuniCostOB_NPV$dis7pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BeachNourish_adj"]

#Estimating NFIP policy costs (23% less after 2012 due to relocating homes)
S4_NFIP_PolicyCost_dis7pct <- S4_FedRevOB_NPV$dis7pct[S4_FedRevOB_NPV$variable=="OB_NFIP_PolicyCost_adj"]

#Estimating NFIP claims (23% less due to relocating homes)
S4_NFIP_Claims_dis7pct <- S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_NFIP_Claims_adj"]

#Estimating 'hidden' costs of rebuilding to households (difference between damages and claims for remaining households that rebuild)
S4_HiddenHHCosts_dis7pct <- Est_Dmg_Claims_Sandy_byScenario$S4[Est_Dmg_Claims_Sandy_byScenario$X=="Est_HiddenHouseholdCosts"]

#Estimating buyout costs (Fed + muni share)
S4_BuyoutCosts_dis7pct <- S4_FedCostOB_NPV$dis7pct[S4_FedCostOB_NPV$variable=="OB_FedShare_BuyoutCost_adj"]  + S4_MuniCostOB_NPV$dis7pct[S4_MuniCostOB_NPV$variable=="OB_LocalShare_BuyoutCost_adj"]

# Create a vector of the object names
object_names <- c("S4_ResponseRecoveryMuni_dis7pct", 
                  "S4_ResponseRecoveryHH_dis7pct", 
                  "S4_RepairRoadsBridges_dis7pct",
                  "S4_BeachRestore_dis7pct",
                  "S4_NFIP_PolicyCost_dis7pct",
                  "S4_NFIP_Claims_dis7pct",
                  "S4_HiddenHHCosts_dis7pct",
                  "S4_BuyoutCosts_dis7pct")


# Use mget to get the values and names
values <- mget(object_names)

# Convert to a dataframe
S4_PolicyCost_dis7pct <- data.frame(Label = names(values), Value = unlist(values))

# Print the dataframe
S4_PolicyCost_dis7pct

# Total Policy cost
S4_TotalPolicyCost_dis7pct <- sum(S4_PolicyCost_dis7pct$Value)
```

###Prospective costs (2022:2063)
####Protection Costs Assumption
```{r include=FALSE}
S4_NPV_ProtectCost_V1_dis7pct <- S4_NPV_ProtectCost$dis7pct[S4_NPV_ProtectCost$Scenario=="V1"]
S4_NPV_ProtectCost_V2_dis7pct <- S4_NPV_ProtectCost$dis7pct[S4_NPV_ProtectCost$Scenario=="V2"]
S4_NPV_ProtectCost_V3_dis7pct <- S4_NPV_ProtectCost$dis7pct[S4_NPV_ProtectCost$Scenario=="V3"]

```

####Residual Risk Assumption 
```{r include = FALSE}
S4_NPV_ResRisk_low_dis7pct <- S4_NPV_ResRisk$dis7pct[S4_NPV_ResRisk$Scenario=="Low"]
S4_NPV_ResRisk_mid_dis7pct <- S4_NPV_ResRisk$dis7pct[S4_NPV_ResRisk$Scenario=="Medium"]
S4_NPV_ResRisk_high_dis7pct <- S4_NPV_ResRisk$dis7pct[S4_NPV_ResRisk$Scenario=="High"]

```

###Summary of NPV by Assumptions :
S4_VLow: Low future costs (low protection costs, low residual risk)
S4_VMid: Mid future costs (mid protection costs, mid residual risk)
S4_VMid: Mid future costs (high protection costs, high residual risk)

```{r include= FALSE}
#Get NPV
S4_VLow_NPV_dis7pct <- S4_TotalPolicyCost_dis7pct + S4_NPV_ProtectCost_V1_dis7pct + 
  S4_NPV_ResRisk_low_dis7pct

S4_VMid_NPV_dis7pct <- S4_TotalPolicyCost_dis7pct + S4_NPV_ProtectCost_V2_dis7pct + 
  S4_NPV_ResRisk_mid_dis7pct

S4_VHigh_NPV_dis7pct <- S4_TotalPolicyCost_dis7pct + S4_NPV_ProtectCost_V3_dis7pct + 
  S4_NPV_ResRisk_high_dis7pct

```

#### Plot 
```{r include = FALSE}

# Create a sample data frame with 3 scenarios and 3 values for each
S4 <- data.frame(
  Assumption= rep(c("Low", "Mid", "High"), each = 3),
  CostType = rep(c("Policy Cost", "Residual Risk", "Protection Cost"), times = 3),
  Cost = c(S4_TotalPolicyCost_dis7pct, 
           S4_NPV_ResRisk_low_dis7pct, 
           S4_NPV_ProtectCost_V1_dis7pct, 
           S4_TotalPolicyCost_dis7pct, 
           S4_NPV_ResRisk_mid_dis7pct, 
           S4_NPV_ProtectCost_V2_dis7pct,
           S4_TotalPolicyCost_dis7pct, 
           S4_NPV_ResRisk_high_dis7pct, 
           S4_NPV_ProtectCost_V3_dis7pct)
)

# Convert 'CostType' to a factor and reorder the levels
S4$Assumption <- factor(S4$Assumption, levels = c("Low", "Mid", "High"))
S4$CostType <- factor(S4$CostType, levels = c("Protection Cost", "Residual Risk", "Policy Cost"))

# Plot the stacked bar chart
ggplot(S4, aes(x = Assumption, y = Cost, fill = CostType)) +
  geom_bar(stat = "identity") +  # Create a stacked bar chart
  labs(title = "S4: Cost Analysis by Assumptions \n(7% Discount Rate)", 
       x = "Assumption", 
       y = "Cost") +
  scale_fill_manual(values = c("Policy Cost" = "blue", 
                               "Residual Risk" = "red", 
                               "Protection Cost" = "black")) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"), limits = c(0, 1250000000)) +
  labs(y = "Cost in Millions of dollars \n(2020USD)")


```

#Generating comparison table
```{r include=FALSE}
# Define values for PolicyCost, ResidualRisk, and ProtectionCost for each assumption
S4_PolicyCost <- list(
  Low = c(S4_TotalPolicyCost_dis2pct, S4_TotalPolicyCost_dis3pct, S4_TotalPolicyCost_dis7pct),
  Mid = c(S4_TotalPolicyCost_dis2pct, S4_TotalPolicyCost_dis3pct, S4_TotalPolicyCost_dis7pct),
  High = c(S4_TotalPolicyCost_dis2pct, S4_TotalPolicyCost_dis3pct, S4_TotalPolicyCost_dis7pct)
)

S4_ResidualRisk <- list(
  Low = c(S4_NPV_ResRisk_low_dis2pct, S4_NPV_ResRisk_low_dis3pct, S4_NPV_ResRisk_low_dis7pct),
  Mid = c(S4_NPV_ResRisk_mid_dis2pct, S4_NPV_ResRisk_mid_dis3pct, S4_NPV_ResRisk_mid_dis7pct),
  High = c(S4_NPV_ResRisk_high_dis2pct, S4_NPV_ResRisk_high_dis3pct, S4_NPV_ResRisk_high_dis7pct)
)

S4_ProtectCost <- list(
  Low = c(S4_NPV_ProtectCost_V1_dis2pct, S4_NPV_ProtectCost_V1_dis3pct, S4_NPV_ProtectCost_V1_dis7pct),
  Mid = c(S4_NPV_ProtectCost_V2_dis2pct, S4_NPV_ProtectCost_V2_dis3pct, S4_NPV_ProtectCost_V2_dis7pct),
  High = c(S4_NPV_ProtectCost_V3_dis2pct, S4_NPV_ProtectCost_V3_dis3pct, S4_NPV_ProtectCost_V3_dis7pct)
)

# Combine into a single dataframe
S4_SummaryTable <- data.frame(
  Scenario = rep(c("S4"), times =9),
  Assumption = rep(c("Low", "Mid", "High"), each = 3),
  DiscountRate = rep(c("2%", "3%", "7%"), times = 3),
  PolicyCost = unlist(S4_PolicyCost),
  ResidualRisk = unlist(S4_ResidualRisk),
  ProtectionCost = unlist(S4_ProtectCost)
)

# Calculate the total cost for each row
S4_SummaryTable$TotalCost <- rowSums(S4_SummaryTable[, c("PolicyCost", "ResidualRisk", "ProtectionCost")])

# Display the combined table
rownames(S4_SummaryTable) <- NULL
print(S4_SummaryTable)

# Write out table
path <- './data/SummaryAndAnalysis'
write.csv(S4_SummaryTable, file.path(path, "S4_SummaryTable.csv"), row.names = FALSE)
```

