---
title: "S1_FedCostsFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---

The purpose of this worksheet is to estimate the NPV for costs to Toms River for Scenario 1. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, and 7% based on justifications provided in the SI.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToFed_fromOB/FedCostOB.csv')
FedCostOB <- FedCostOB[,c(-1)]

```





#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedCostOB), value = TRUE)

# Apply the discount function
FedCostOB_dis2pct <- dis2pct(FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedCostOB_dis2pct <- FedCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis2pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis2pct +
           OB_FedShare_DCA_LGS_UN_adj_dis2pct +
           OB_FedShare_DCA_LPS_adj_dis2pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis2pct +
           OB_FedShare_HMGP_adj_dis2pct +
           OB_FedShare_PA_Debris_adj_dis2pct +
           OB_FedShare_PA_Protect_adj_dis2pct +
           OB_FedShare_PA_PublicBuildings_adj_dis2pct +
           OB_FedShare_PA_PublicUtilities_adj_dis2pct +
           OB_FedShare_PA_Recreational_adj_dis2pct +
           OB_FedShare_CDL_Muni_adj_dis2pct +
           OB_FedShare_CDL_School_adj_dis2pct +
           OB_FedShare_CDL_Services_adj_dis2pct)

#Summing Fed Share for Roads & Bridges
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis2pct= 
           OB_FedShare_DOT_Rt35_adj_dis2pct +
           OB_FedShare_DOT_Rt37_adj_dis2pct +
           OB_FedShare_PA_RoadsBridges_adj_dis2pct +
           OB_FedShare_DOT_LocalAid_adj_dis2pct)


#Summing Fed Share for Response and Recovery Direct to Individuals
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis2pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis2pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis2pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis2pct +
           OB_FedShare_DCA_Resettlement_LMI_dis2pct +
           OB_FedShare_DCA_Resettlement_UN_dis2pct +
           OB_FedShare_DCA_RREM_LMI_dis2pct +
           OB_FedShare_DCA_RREM_UN_dis2pct+
           OB_FedShare_IA_dis2pct)
  
#Summing Fed Share for NFIP Claims Direct to Individuals
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_NFIP_dis2pct= 
           OB_NFIP_BuildingsClaims_adj_dis2pct +
           OB_NFIP_ContentsClaims_adj_dis2pct +
           OB_NFIP_ICCClaims_adj_dis2pct)



#selecting variables
FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis2pct,
         OB_FedShare_ToRoadsBridge_dis2pct,
         OB_FedShare_BeachNourish_adj_dis2pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis2pct,
         OB_FedShare_NFIP_dis2pct)

#pivoting tables longer
S1_FedCostOB_dis2pct_NPV <- FedCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedCostOB), value = TRUE)

# Apply the discount function
FedCostOB_dis3pct <- dis3pct(FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedCostOB_dis3pct <- FedCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis3pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis3pct +
           OB_FedShare_DCA_LGS_UN_adj_dis3pct +
           OB_FedShare_DCA_LPS_adj_dis3pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis3pct +
           OB_FedShare_HMGP_adj_dis3pct +
           OB_FedShare_PA_Debris_adj_dis3pct +
           OB_FedShare_PA_Protect_adj_dis3pct +
           OB_FedShare_PA_PublicBuildings_adj_dis3pct +
           OB_FedShare_PA_PublicUtilities_adj_dis3pct +
           OB_FedShare_PA_Recreational_adj_dis3pct +
           OB_FedShare_CDL_Muni_adj_dis3pct +
           OB_FedShare_CDL_School_adj_dis3pct +
           OB_FedShare_CDL_Services_adj_dis3pct)

#Summing Fed Share for Roads & Bridges
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis3pct= 
           OB_FedShare_DOT_Rt35_adj_dis3pct +
           OB_FedShare_DOT_Rt37_adj_dis3pct +
           OB_FedShare_PA_RoadsBridges_adj_dis3pct +
           OB_FedShare_DOT_LocalAid_adj_dis3pct)


#Summing Fed Share for Response and Recovery Direct to Individuals
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis3pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis3pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis3pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis3pct +
           OB_FedShare_DCA_Resettlement_LMI_dis3pct +
           OB_FedShare_DCA_Resettlement_UN_dis3pct +
           OB_FedShare_DCA_RREM_LMI_dis3pct +
           OB_FedShare_DCA_RREM_UN_dis3pct +
          OB_FedShare_IA_dis3pct)
  
#Summing Fed Share for NFIP Claims Direct to Individuals
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_NFIP_dis3pct= 
           OB_NFIP_BuildingsClaims_adj_dis3pct +
           OB_NFIP_ContentsClaims_adj_dis3pct +
           OB_NFIP_ICCClaims_adj_dis3pct)



#selecting variables
FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis3pct,
         OB_FedShare_ToRoadsBridge_dis3pct,
         OB_FedShare_BeachNourish_adj_dis3pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis3pct,
         OB_FedShare_NFIP_dis3pct)

#pivoting tables longer
S1_FedCostOB_dis3pct_NPV <- FedCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedCostOB), value = TRUE)

# Apply the discount function
FedCostOB_dis7pct <- dis7pct(FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedCostOB_dis7pct <- FedCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis7pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis7pct +
           OB_FedShare_DCA_LGS_UN_adj_dis7pct +
           OB_FedShare_DCA_LPS_adj_dis7pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis7pct +
           OB_FedShare_HMGP_adj_dis7pct +
           OB_FedShare_PA_Debris_adj_dis7pct +
           OB_FedShare_PA_Protect_adj_dis7pct +
           OB_FedShare_PA_PublicBuildings_adj_dis7pct +
           OB_FedShare_PA_PublicUtilities_adj_dis7pct +
           OB_FedShare_PA_Recreational_adj_dis7pct +
           OB_FedShare_CDL_Muni_adj_dis7pct +
           OB_FedShare_CDL_School_adj_dis7pct +
           OB_FedShare_CDL_Services_adj_dis7pct)

#Summing Fed Share for Roads & Bridges
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis7pct= 
           OB_FedShare_DOT_Rt35_adj_dis7pct +
           OB_FedShare_DOT_Rt37_adj_dis7pct +
           OB_FedShare_PA_RoadsBridges_adj_dis7pct +
           OB_FedShare_DOT_LocalAid_adj_dis7pct)


#Summing Fed Share for Response and Recovery Direct to Individuals
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis7pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis7pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis7pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis7pct +
           OB_FedShare_DCA_Resettlement_LMI_dis7pct +
           OB_FedShare_DCA_Resettlement_UN_dis7pct +
           OB_FedShare_DCA_RREM_LMI_dis7pct +
           OB_FedShare_DCA_RREM_UN_dis7pct +
           OB_FedShare_IA_dis7pct)
  
#Summing Fed Share for NFIP Claims Direct to Individuals
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_NFIP_dis7pct= 
           OB_NFIP_BuildingsClaims_adj_dis7pct +
           OB_NFIP_ContentsClaims_adj_dis7pct +
           OB_NFIP_ICCClaims_adj_dis7pct)



#selecting variables
FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis7pct,
         OB_FedShare_ToRoadsBridge_dis7pct,
         OB_FedShare_BeachNourish_adj_dis7pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis7pct,
         OB_FedShare_NFIP_dis7pct)

#pivoting tables longer
S1_FedCostOB_dis7pct_NPV <- FedCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S1_FedCostOB_NPV <- S1_FedCostOB_dis2pct_NPV %>%
  left_join(S1_FedCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S1_FedCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S1_OB_Inputs'

write.csv(S1_FedCostOB_NPV, file.path(path, "S1_FedCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S1_FedCostOB_dis2pct_NPV <- sum(S1_FedCostOB_NPV$dis2pct)
S1_FedCostOB_dis3pct_NPV <- sum(S1_FedCostOB_NPV$dis3pct)
S1_FedCostOB_dis7pct_NPV <- sum(S1_FedCostOB_NPV$dis7pct)

S1_FedCostOB_dis2pct_NPV
S1_FedCostOB_dis3pct_NPV
S1_FedCostOB_dis7pct_NPV
```