---
title: "S4_FedCostsFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---

The purpose of this worksheet is to estimate the NPV for costs to the Federal Gov for Scenario 4. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, and 7% based on justifications provided in the SI.

Assumptions:
-Federal cost share for Municipal Response and recovery (assume sticky)
-Federal cost share for repairs to Roads & Bridges (assume sticky)
-Federal cost share for Response & Recovery direct to Household (50% less proportional to homes bought out- redirected)
-Federal NFIP claims (less amount redirected to buyouts for year 2012, assume the rest are sticky)
-Federal cost share for beach nourishment (assume sticky)

-Federal cost share for buyouts = 75% of market price for 920 bought out homes




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToFed_fromOB/FedCostOB.csv')
FedCostOB <- FedCostOB[,c(-1)]

#for buyout cost estimates
OB_S4_2011_SP <- read.csv('./data/BaselineData/OB_MOD4_Produced/TimeSeries/S4/OB_S4_2011_SalesPrice.csv')
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')

#for S4 NFIP damages and claims estimates
OB_DmgClaims_est <- read.csv('./data/BaselineData/OB_SandyDmgsAndClaimsEstByScenario/Est_Dmg_Claims_Sandy_byScenario.csv')

```

#Adjusting S1 inputs based on S4 assumptions

## Reducing Federal cost share for Response & Recovery direct to Household by 50% (proportional to homes bought out- redirected). 
```{r include=FALSE}
S4_FedCostOB <- FedCostOB

S4_FedCostOB$OB_FedShare_DCA_Housing_LandLordIncentive_LMI <- S4_FedCostOB$OB_FedShare_DCA_Housing_LandLordIncentive_LMI*0.50

S4_FedCostOB$OB_FedShare_DCA_Housing_NEP_LMI <- S4_FedCostOB$OB_FedShare_DCA_Housing_NEP_LMI *0.50

S4_FedCostOB$OB_FedShare_DCA_LMI_HomeownerRebuilding <- S4_FedCostOB$OB_FedShare_DCA_LMI_HomeownerRebuilding*0.50

S4_FedCostOB$OB_FedShare_DCA_Resettlement_LMI <- S4_FedCostOB$OB_FedShare_DCA_Resettlement_LMI*0.50

S4_FedCostOB$OB_FedShare_DCA_Resettlement_UN <- S4_FedCostOB$OB_FedShare_DCA_Resettlement_UN*0.50

S4_FedCostOB$OB_FedShare_DCA_Resettlement_UN <- S4_FedCostOB$OB_FedShare_DCA_Resettlement_UN*0.50

S4_FedCostOB$OB_FedShare_DCA_RREM_LMI <- S4_FedCostOB$OB_FedShare_DCA_RREM_LMI*0.50

S4_FedCostOB$OB_FedShare_DCA_RREM_UN<- S4_FedCostOB$OB_FedShare_DCA_RREM_UN*0.50

S4_FedCostOB$OB_FedShare_IA <- S4_FedCostOB$OB_FedShare_IA*0.50


```


#Federal NFIP claims (less amount redirected to buyouts for year 2012, assume the rest are sticky)
```{r include=FALSE}
#obtaining NFIP claims estimated for S4 cluster
S4_EstNFIPClaims_2012 <- OB_DmgClaims_est$S4[OB_DmgClaims_est$X=="Est_NFIP_Claims"]

#adjusting estimate to 2020 USD
S4_EstNFIPClaims_2012_adj <- S4_EstNFIPClaims_2012 * Adjust$CPI_Multiplier_USD2020[Adjust$Year==2012]

#Creating bucket with original NFIP claims
S4_FedCostOB<- S4_FedCostOB %>%
  mutate(OB_NFIP_Claims_adj= 
           OB_NFIP_BuildingsClaims_adj +
           OB_NFIP_ContentsClaims_adj +
           OB_NFIP_ICCClaims_adj)

#Subtracting 2012 NFIP claims for cluster 3 from total claims
S4_FedCostOB$OB_NFIP_Claims_adj[S4_FedCostOB$Year == 2012] <- 
  S4_FedCostOB$OB_NFIP_Claims_adj[S4_FedCostOB$Year == 2012] - S4_EstNFIPClaims_2012_adj


```

## Federal cost share for buyouts (75 % of market price for 421 homes)
```{r include= FALSE}
#subset estimated sales price to the winzorized estimates
OB_S4_2011_SP <- OB_S4_2011_SP[,c("SalesPrice_LG_W")]

#obtain total sales price value in Ortley Beach
OB_S4_2011_SP_Tot <- sum(OB_S4_2011_SP)


#adjusting by CPI to 2020 USD
y11_CPI_multiplier <- Adjust$CPI_Multiplier_USD2020[Adjust$Year==2011]

OB_S4_2011_SP_Tot_adj <- OB_S4_2011_SP_Tot * y11_CPI_multiplier

#obtaining Fed share of buyout cost (75% cost share)
OB_FedShare_BuyoutCost_adj <- OB_S4_2011_SP_Tot_adj * 0.75

#Adding buyout costs to data frame 
S4_FedCostOB$OB_FedShare_BuyoutCost_adj[S4_FedCostOB$Year==2012] <- OB_FedShare_BuyoutCost_adj
```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S4_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedCostOB), value = TRUE)

# Apply the discount function
S4_FedCostOB_dis2pct <- dis2pct(S4_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedCostOB_dis2pct <- S4_FedCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis2pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis2pct +
           OB_FedShare_DCA_LGS_UN_adj_dis2pct +
           OB_FedShare_DCA_LPS_adj_dis2pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis2pct +
           OB_FedShare_HMGP_adj_dis2pct +
           OB_FedShare_PA_Debris_adj_dis2pct +
           OB_FedShare_PA_Protect_adj_dis2pct +
           OB_FedShare_PA_PublicBuildings_adj_dis2pct +
           OB_FedShare_PA_PublicUtilities_adj_dis2pct +
           OB_FedShare_PA_Recreational_adj_dis2pct +
           OB_FedShare_CDL_Muni_adj_dis2pct +
           OB_FedShare_CDL_School_adj_dis2pct +
           OB_FedShare_CDL_Services_adj_dis2pct)

#Summing Fed Share for Roads & Bridges (Sticky)
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis2pct= 
           OB_FedShare_DOT_Rt35_adj_dis2pct +
           OB_FedShare_DOT_Rt37_adj_dis2pct +
           OB_FedShare_PA_RoadsBridges_adj_dis2pct +
           OB_FedShare_DOT_LocalAid_adj_dis2pct)


#Summing Fed Share for Response and Recovery Direct to Individuals (has been reduced by 23%)
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis2pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis2pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis2pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis2pct +
           OB_FedShare_DCA_Resettlement_LMI_dis2pct +
           OB_FedShare_DCA_Resettlement_UN_dis2pct +
           OB_FedShare_DCA_RREM_LMI_dis2pct +
           OB_FedShare_DCA_RREM_UN_dis2pct +
           OB_FedShare_IA_dis2pct)
  

#selecting variables
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis2pct,
         OB_FedShare_ToRoadsBridge_dis2pct,
         OB_FedShare_BeachNourish_adj_dis2pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis2pct,
         OB_NFIP_Claims_adj_dis2pct,
         OB_FedShare_BuyoutCost_adj_dis2pct)

#pivoting tables longer
S4_FedCostOB_dis2pct_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S4_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedCostOB), value = TRUE)

# Apply the discount function
S4_FedCostOB_dis3pct <- dis3pct(S4_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedCostOB_dis3pct <- S4_FedCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis3pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis3pct +
           OB_FedShare_DCA_LGS_UN_adj_dis3pct +
           OB_FedShare_DCA_LPS_adj_dis3pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis3pct +
           OB_FedShare_HMGP_adj_dis3pct +
           OB_FedShare_PA_Debris_adj_dis3pct +
           OB_FedShare_PA_Protect_adj_dis3pct +
           OB_FedShare_PA_PublicBuildings_adj_dis3pct +
           OB_FedShare_PA_PublicUtilities_adj_dis3pct +
           OB_FedShare_PA_Recreational_adj_dis3pct +
           OB_FedShare_CDL_Muni_adj_dis3pct +
           OB_FedShare_CDL_School_adj_dis3pct +
           OB_FedShare_CDL_Services_adj_dis3pct)

#Summing Fed Share for Roads & Bridges (Sticky)
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis3pct= 
           OB_FedShare_DOT_Rt35_adj_dis3pct +
           OB_FedShare_DOT_Rt37_adj_dis3pct +
           OB_FedShare_PA_RoadsBridges_adj_dis3pct +
           OB_FedShare_DOT_LocalAid_adj_dis3pct)


#Summing Fed Share for Response and Recovery Direct to Individuals (has been reduced by 33%)
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis3pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis3pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis3pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis3pct +
           OB_FedShare_DCA_Resettlement_LMI_dis3pct +
           OB_FedShare_DCA_Resettlement_UN_dis3pct +
           OB_FedShare_DCA_RREM_LMI_dis3pct +
           OB_FedShare_DCA_RREM_UN_dis3pct +
           OB_FedShare_IA_dis3pct)
  

#selecting variables
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis3pct,
         OB_FedShare_ToRoadsBridge_dis3pct,
         OB_FedShare_BeachNourish_adj_dis3pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis3pct,
         OB_NFIP_Claims_adj_dis3pct,
         OB_FedShare_BuyoutCost_adj_dis3pct)

#pivoting tables longer
S4_FedCostOB_dis3pct_NPV <- S4_FedCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S4_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedCostOB), value = TRUE)

# Apply the discount function
S4_FedCostOB_dis7pct <- dis7pct(S4_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedCostOB_dis7pct <- S4_FedCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government / School District
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis7pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis7pct +
           OB_FedShare_DCA_LGS_UN_adj_dis7pct +
           OB_FedShare_DCA_LPS_adj_dis7pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis7pct +
           OB_FedShare_HMGP_adj_dis7pct +
           OB_FedShare_PA_Debris_adj_dis7pct +
           OB_FedShare_PA_Protect_adj_dis7pct +
           OB_FedShare_PA_PublicBuildings_adj_dis7pct +
           OB_FedShare_PA_PublicUtilities_adj_dis7pct +
           OB_FedShare_PA_Recreational_adj_dis7pct +
           OB_FedShare_CDL_Muni_adj_dis7pct +
           OB_FedShare_CDL_School_adj_dis7pct +
           OB_FedShare_CDL_Services_adj_dis7pct)

#Summing Fed Share for Roads & Bridges (Sticky)
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToRoadsBridge_dis7pct= 
           OB_FedShare_DOT_Rt35_adj_dis7pct +
           OB_FedShare_DOT_Rt37_adj_dis7pct +
           OB_FedShare_PA_RoadsBridges_adj_dis7pct +
           OB_FedShare_DOT_LocalAid_adj_dis7pct +
            OB_FedShare_IA_dis7pct)


#Summing Fed Share for Response and Recovery Direct to Individuals (has been reduced by 23%)
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToIndiv_ResponseRecovery_dis7pct= 
           OB_FedShare_DCA_Housing_LandLordIncentive_LMI_dis7pct +
           OB_FedShare_DCA_Housing_NEP_LMI_dis7pct +
           OB_FedShare_DCA_LMI_HomeownerRebuilding_dis7pct +
           OB_FedShare_DCA_Resettlement_LMI_dis7pct +
           OB_FedShare_DCA_Resettlement_UN_dis7pct +
           OB_FedShare_DCA_RREM_LMI_dis7pct +
           OB_FedShare_DCA_RREM_UN_dis7pct)
  

#selecting variables
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis7pct,
         OB_FedShare_ToRoadsBridge_dis7pct,
         OB_FedShare_BeachNourish_adj_dis7pct,
         OB_FedShare_ToIndiv_ResponseRecovery_dis7pct,
         OB_NFIP_Claims_adj_dis7pct,
         OB_FedShare_BuyoutCost_adj_dis7pct)

#pivoting tables longer
S4_FedCostOB_dis7pct_NPV <- S4_FedCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```



#Joining data for export
```{r include = FALSE}

S4_FedCostOB_NPV <- S4_FedCostOB_dis2pct_NPV %>%
  left_join(S4_FedCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S4_FedCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S4_OB_Inputs'

write.csv(S4_FedCostOB_NPV, file.path(path, "S4_FedCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S4_FedCostOB_dis2pct_NPV <- sum(S4_FedCostOB_NPV$dis2pct)
S4_FedCostOB_dis3pct_NPV <- sum(S4_FedCostOB_NPV$dis3pct)
S4_FedCostOB_dis7pct_NPV <- sum(S4_FedCostOB_NPV$dis7pct)

S4_FedCostOB_dis2pct_NPV
S4_FedCostOB_dis3pct_NPV
S4_FedCostOB_dis7pct_NPV
```