---
title: "Municipal Revenues from Ortley Beach"
author: "Laura Geronimo"
date: "2024-08-23"
output: html_document
---
The purpose of this worksheet is to estimate the benefits to Toms River municipality of maintaining properties on Ortley Beach for Scenario 1 (replicating the historical record). Benefits are estimated for the 2012-2022 period. Specifically we estimate:

1) Municipal property tax revenue to Toms River from Ortley Beach properties 
2) School district property tax revenue to the Toms River School District from Ortley Beach properties
3) Federal cost share for municipal recovery and response activities post Superstorm Sandy
4) Federal cost share for repairs to routes 35 and 37 serving Ortley Beach
5) Federal cost share for Beach and dune nourishment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)

##Importing Data ####
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')
TR_Tax <- read.csv('./data/BaselineData/TomsRiver_TaxData.csv')
OB_M4_Improvement <- read_xlsx('./data/BaselineData/OrtleyBeach_MODIV_ImprovementValue.xlsx', sheet=1)

#Data for downscaling
OB_DemoData <- read_xlsx('./data/BaselineData/OrtleyBeach_DemographicData.xlsx', sheet=1)
TR_DemoData <- read_xlsx('./data/BaselineData/TomsRiver_DemographicData.xlsx', sheet=1)

TR_Sandy <- read_xlsx('./data/BaselineData/TomsRiver_SandyImpacts.xlsx', sheet=1)

#Federal cost share for recovery to TR (needs further downscaling to OB)
TR_FedShare_DCA <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/DCA_TR_FederalShare_ByYear.csv')
TR_FedShare_PA <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/PA_TR_FederalShare_ByYear.csv')
TR_FedShare_HMGP <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/HMGP_TR_FederalShare_ByYear.csv')
TR_FedShare_CDL <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/CDL_TR_FederalShare_ByYear.csv')


#Federal cost share for recovery to OB (already downscaled to Ortley Beach (OB))
OB_FedShare_DOT_LocalAid <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/DOT_LocalAid_OB_FedShareEst_ByYear.csv')
OB_FedShare_DOT <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/DOT_OB_FederalShare_ByYear.csv')
OB_BeachNourish <- read.csv('./data/BaselineData/OrtleyBeach_BeachNourishment.csv')




```

#1: Estimating Municipal Property Tax Revenue from Ortley Beach 
```{r include=FALSE}
#selecting multipliers
Adjust <- Adjust %>% select(
  "Year",
  "CPI_Multiplier_USD2020",
  "SLIPD_V3_Multiplier"
)

#selecting needed tax info
TR_Tax <- TR_Tax %>% select(
  "Year",
  "TR_Eq_Ratio",
  "TR_TotalMuniTaxRate",
  "TR_SchoolTaxRate" 
)

#joining multipliers and tax rates to improvement value estimates ####
OB_M4_Improvement <- left_join(OB_M4_Improvement, Adjust, by="Year")
OB_M4_Improvement <- left_join(OB_M4_Improvement, TR_Tax, by="Year")

#Multiplying Improvement value by equalization ratio (Eq_Ratio) to obtain estimate market value####
OB_M4_Improvement$OB_NetEstImprovementValue_EqRatio <- OB_M4_Improvement$OB_NetEstimatedImprovementValue * OB_M4_Improvement$TR_Eq_Ratio

#Multiplying Improvement value by total municipal tax ratio to get Toms River (TR) tax revenues from Ortley Beach (OB) properties####
OB_M4_Improvement$OB_PropTaxRevTo_TR_Muni <- OB_M4_Improvement$OB_NetEstImprovementValue_EqRatio * OB_M4_Improvement$TR_TotalMuniTaxRate

#Multiplying the obtained property tax revenue from Ortley Beach by the State and Local Implicit Price Deflator for GDP (SLIPD) to bring values to 2020 USD (marked by _adj)####
OB_M4_Improvement$OB_PropTaxRevTo_TR_Muni_adj <- OB_M4_Improvement$OB_PropTaxRevTo_TR_Muni * OB_M4_Improvement$SLIPD_V3_Multiplier

#creating standalone table for estimate####
OB_PropTaxRevTo_TR_Muni<- OB_M4_Improvement %>% select(
  "Year",
  "OB_PropTaxRevTo_TR_Muni_adj")
OB_PropTaxRevTo_TR_Muni <- na.omit(OB_PropTaxRevTo_TR_Muni) #dropping 2010 due to NAs

```

#2: Estimating School District Property Tax Revenue from Ortley Beach 
```{r include=FALSE}

#Multiplying Improvement value by total school tax ratio to get Toms River (TR) tax revenues from Ortley Beach (OB) properties####
OB_M4_Improvement$OB_PropTaxRevTo_TR_School <- OB_M4_Improvement$OB_NetEstImprovementValue_EqRatio * OB_M4_Improvement$TR_SchoolTaxRate

#Multiplying the obtained property tax revenue from Ortley Beach by the State and Local Implicit Price Deflator for GDP (SLIPD) to bring values to 2020 USD (marked by _adj)####
OB_M4_Improvement$OB_PropTaxRevTo_TR_School_adj <- OB_M4_Improvement$OB_PropTaxRevTo_TR_School * OB_M4_Improvement$SLIPD_V3_Multiplier

#creating standalone table for estimate ####
OB_PropTaxRevTo_TR_School <- OB_M4_Improvement %>% select(
  "Year",
  "OB_PropTaxRevTo_TR_School_adj")
OB_PropTaxRevTo_TR_School <- na.omit(OB_PropTaxRevTo_TR_School) #dropping 2010 due to NAs

```

#3) Estimating Federal cost share distributed to the Toms River Township for recovery and response activities in Ortley Beach post Superstorm Sandy

##3a) Downscaling HUD / DCA data for to Ortley Beach by project
```{r include=FALSE}

##LGS Support for Governmental Entities LMI####
###Subset the DCA projects to LGS Support for Governmental Entities LMI####
TR_FedShare_DCA_LGS_LMI <- subset(TR_FedShare_DCA, Project=="DCA LGS Support for Governmental Entities LMI")

###Downscaling DCA LGS Support for Governmental Entities LMI (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA LMI funds disbursed per capita #### 
TR_FedShare_DCA_LGS_LMI$LMI_Disbursed_Per_Capita <- TR_FedShare_DCA_LGS_LMI$Federal.Disbursed.Amount.Adj / TR_lmi_pop_estimate

###Estimate DCA LMI funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_LGS_LMI$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_LGS_LMI$LMI_Disbursed_Per_Capita * OB_lmi_pop_estimate 

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_FedShare_DCA_LGS_LMI <- TR_FedShare_DCA_LGS_LMI %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_LGS_LMI) <- c("Year", "OB_FedShare_DCA_LGS_LMI_adj")

##LGS Support for Governmental Entities UN####
###Subset the DCA projects to DCA LGS Support for Governmental Entities UN ####
TR_FedShare_DCA_LGS_UN <- subset(TR_FedShare_DCA, Project=="DCA LGS Support for Governmental Entities UN")

###Downscaling DCA LGS Support for Governmental Entities UN (downscaled by total population) ####
TR_tot_pop_estimate <-  TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-  OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate DCA UN funds disbursed per capita #### 
TR_FedShare_DCA_LGS_UN$UN_Disbursed_Per_Capita <- TR_FedShare_DCA_LGS_UN$Federal.Disbursed.Amount.Adj / 
TR_tot_pop_estimate

###Estimate DCA UN funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_LGS_UN$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_LGS_UN$UN_Disbursed_Per_Capita * OB_tot_pop_estimate  

###Separate downscaled OB estimates for UN projects as standalone table####
OB_FedShare_DCA_LGS_UN <- TR_FedShare_DCA_LGS_UN %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_LGS_UN) <- c("Year", "OB_FedShare_DCA_LGS_UN_adj")


##DCA Local Planning Services####
###Subset the DCA Local Planning Services####
TR_FedShare_DCA_LPS <- subset(TR_FedShare_DCA, Project=="DCA Local Planning Services")

###Downscaling DCA Local Planning Services  (downscaled by total population) ####
TR_tot_pop_estimate <- TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate DCA Local Planning Services disbursed per capita #### 
TR_FedShare_DCA_LPS$LPS_Disbursed_Per_Capita <- TR_FedShare_DCA_LPS$Federal.Disbursed.Amount.Adj / 
TR_tot_pop_estimate

###Estimate DCA Local Planning Services funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_LPS$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_LPS$LPS_Disbursed_Per_Capita * OB_tot_pop_estimate  

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_LPS <- TR_FedShare_DCA_LPS %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_LPS) <- c("Year", "OB_FedShare_DCA_LPS_adj")


##DCA Zoning Code Enforcement Grant UN####
###Subset the DCA Zoning Code Enforcement Grant UN####
TR_FedShare_DCA_ZCEG_UN <- subset(TR_FedShare_DCA, Project=="DCA Zoning Code Enforcement Grant UN")

###Downscaling DCA Zoning Code Enforcement Grant UN  (downscaled by total population) ####
TR_tot_pop_estimate <- TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate DCA Local Planning Services disbursed per capita #### 
TR_FedShare_DCA_ZCEG_UN$ZCEG_Disbursed_Per_Capita <- TR_FedShare_DCA_ZCEG_UN$Federal.Disbursed.Amount.Adj / TR_tot_pop_estimate

###Estimate DCA Local Planning Services funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_ZCEG_UN$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_ZCEG_UN$ZCEG_Disbursed_Per_Capita * OB_tot_pop_estimate  

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_ZCEG_UN <- TR_FedShare_DCA_ZCEG_UN %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_ZCEG_UN) <- c("Year", "OB_FedShare_DCA_ZCEG_UN_adj")
```


##3b) Downscaling HMGP data for to Ortley Beach by project
```{r include=FALSE}

#note: Based on project information, HMGP funds went to public buildings outside of Ortley Beach
head(TR_FedShare_HMGP)

###Downscaling Generators by public buildings####
TR_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_PublicBuildings_SandyImpact"]
OB_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_PublicBuildings_SandyImpact"]

###Estimate HMGP disbursed per Sandy impacted public buildings #### 
TR_FedShare_HMGP$HMGP_Disbursed_Per_PublicBuildings <- TR_FedShare_HMGP$Federal.Disbursed.Amount.Adj / 
TR_PublicBuildings 

###Estimate HMGP disbursed funds disbursed to Ortley Beach #### 
TR_FedShare_HMGP$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_HMGP$HMGP_Disbursed_Per_PublicBuildings * OB_PublicBuildings

###Separate downscaled OB estimates for HMGP projects as standalone table####
OB_FedShare_HMGP<- TR_FedShare_HMGP %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_HMGP) <- c("Year", "OB_FedShare_HMGP_adj")

```

##3c) Downscaling PA data for to Ortley Beach by project

```{r include=FALSE}


##PA Debris Removal by housing units impacted by Sandy####
###Subset PA to Debris Removal####
TR_FedShare_PA_Debris <- subset(TR_FedShare_PA, Project=="A - Debris Removal")

###Downscaling PA Debris Removal (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Debris Removal disbursed per housing unit #### 
TR_FedShare_PA_Debris$PA_Debris_Disbursed_Per_Capita <- TR_FedShare_PA_Debris$Federal.Disbursed.Amount.Adj / TR_HU_SandyImpact 

###Estimate PA Debris Removal funds disbursed to Ortley Beach #### 
TR_FedShare_PA_Debris$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_Debris$PA_Debris_Disbursed_Per_Capita * OB_HU_SandyImpact  

###Separate downscaled OB estimates for Debris Removal projects as standalone table####
OB_FedShare_PA_Debris <- TR_FedShare_PA_Debris %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_PA_Debris) <- c("Year", "OB_FedShare_PA_Debris_adj")


##PA Protective Measures by housing units impacted by Sandy####
###Subset PA to Protective Measures####
TR_FedShare_PA_Protect <- subset(TR_FedShare_PA, Project=="B - Protective Measures")

###Downscaling PA Protective Measures (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Protective Measures disbursed per housing unit #### 
TR_FedShare_PA_Protect$PA_Protect_Disbursed_Per_Capita <- TR_FedShare_PA_Protect$Federal.Disbursed.Amount.Adj / TR_HU_SandyImpact 

###Estimate PA Protective Measures funds disbursed to Ortley Beach #### 
TR_FedShare_PA_Protect$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_Protect$PA_Protect_Disbursed_Per_Capita * OB_HU_SandyImpact  

###Separate downscaled OB estimates for PA Protective Measures as standalone table####
OB_FedShare_PA_Protect <- TR_FedShare_PA_Protect %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_PA_Protect) <- c("Year", "OB_FedShare_PA_Protect_adj")


##PA  Roads & Bridges downscaled by miles of local roadway impacted by Sandy####
###Subset PA to Roads and Bridges####
TR_FedShare_PA_RoadsBridges <- subset(TR_FedShare_PA, Project=="C - Roads & Bridges")

###Downscaling PA Roads and Bridges (downscaled by miles of local roadway impacted by Sandy) ####
TR_Roads_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_LocalRoads_TotalMiles_SandyImpact"]
OB_Roads_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_LocalRoads_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_FedShare_PA_RoadsBridges$PA_Roads_Disbursed_Per_RoadMiles <- TR_FedShare_PA_RoadsBridges$Federal.Disbursed.Amount.Adj / TR_Roads_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_FedShare_PA_RoadsBridges$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_RoadsBridges$PA_Roads_Disbursed_Per_RoadMiles * OB_Roads_SandyImpact  

###Separate downscaled OB estimates for PA Roads and Bridges as standalone table####
OB_FedShare_PA_RoadsBridges <- TR_FedShare_PA_RoadsBridges  %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_PA_RoadsBridges) <- c("Year", "OB_FedShare_PA_RoadsBridges_adj")

##PA  Public Buildings downscaled by public buildings impacted by Sandy####
###Subset PA to Public Buildings####
TR_FedShare_PA_PublicBuildings<- subset(TR_FedShare_PA, Project=="E - Public Buildings")

###Downscaling PA Public Buildings (downscaled by public buildings impacted by Sandy ####
TR_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_PublicBuildings_SandyImpact"]
OB_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_PublicBuildings_SandyImpact"]

###Estimate PA Public Buildings disbursed per road miles impacted#### 
TR_FedShare_PA_PublicBuildings$PA_PublicBuildings_Disbursed_Per_PublicBuildings <- TR_FedShare_PA_PublicBuildings$Federal.Disbursed.Amount.Adj / TR_PublicBuildings

###Estimate PA Public Buildings funds disbursed to Ortley Beach #### 
TR_FedShare_PA_PublicBuildings$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_PublicBuildings$PA_PublicBuildings_Disbursed_Per_PublicBuildings * OB_PublicBuildings 

###Separate downscaled OB estimates for Public Buildings projects as standalone table####
OB_FedShare_PA_PublicBuildings<- TR_FedShare_PA_PublicBuildings %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_PA_PublicBuildings) <- c("Year", "OB_FedShare_PA_PublicBuildings_adj")


##PA  Public Utilities downscaled by housing units impacted by Sandy####
###Subset PA to Public Utilities####
TR_FedShare_PA_PublicUtilities<- subset(TR_FedShare_PA, Project=="F - Public Utilities")

###Downscaling PA Public Utilities (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_FedShare_PA_PublicUtilities$PA_PublicUtilities_Disbursed_Per_PublicUtilities <- TR_FedShare_PA_PublicUtilities$Federal.Disbursed.Amount.Adj / TR_HU_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_FedShare_PA_PublicUtilities$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_PublicUtilities$PA_PublicUtilities_Disbursed_Per_PublicUtilities * OB_HU_SandyImpact 

###Separate downscaled OB estimates for Public Utilities projects as standalone table####
OB_FedShare_PA_PublicUtilities<- TR_FedShare_PA_PublicUtilities %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_PA_PublicUtilities) <- c("Year", "OB_FedShare_PA_PublicUtilities_adj")



##PA  Recreational or Other downscaled by housing units impacted by Sandy####
###Subset PA to Recreational or Other####
TR_FedShare_PA_Recreational<- subset(TR_FedShare_PA, Project=="G - Recreational or Other")

###Downscaling PA Recreational or Other (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_FedShare_PA_Recreational$PA_Recreational_Disbursed_Per_Recreational <- TR_FedShare_PA_Recreational$Federal.Disbursed.Amount.Adj / TR_HU_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_FedShare_PA_Recreational$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_PA_Recreational$PA_Recreational_Disbursed_Per_Recreational * OB_HU_SandyImpact 

###Separate downscaled OB estimates for Recreational or Other projects as standalone table####
OB_FedShare_PA_Recreational<- TR_FedShare_PA_Recreational %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

OB_FedShare_PA_Recreational <- subset(OB_FedShare_PA_Recreational, Disbursed.Year < 2023)

###Renaming columns####
colnames(OB_FedShare_PA_Recreational) <- c("Year", "OB_FedShare_PA_Recreational_adj")

```

##3d) Downscaling FEMA CDL data for to Ortley Beach 

```{r include=FALSE}


##Downscaling CDL Municipality by total population####
###Subset CDL to Municipality####
TR_FedShare_CDL_Muni <- subset(TR_FedShare_CDL, Project=="Municipality")

###Downscaling CDL Municipality (downscaled by total population) ####
TR_tot_pop_estimate <-  TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-  OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate CDL Municipality disbursed per capita #### 
TR_FedShare_CDL_Muni$CDL_Muni_Disbursed_Per_Capita <- TR_FedShare_CDL_Muni$Federal.Disbursed.Amount.Adj / TR_tot_pop_estimate

###Estimate CDL Municipality funds disbursed to Ortley Beach #### 
TR_FedShare_CDL_Muni$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_CDL_Muni$CDL_Muni_Disbursed_Per_Capita * OB_tot_pop_estimate  

###Separate downscaled OB estimates for Debris Removal projects as standalone table####
OB_FedShare_CDL_Muni <- TR_FedShare_CDL_Muni %>% select(
  "As.Of.Date",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_CDL_Muni) <- c("Year", "OB_FedShare_CDL_Muni_adj")

##Downscaling CDL School by total school age children####
###Subset CDL to Schools####
TR_FedShare_CDL_School <- subset(TR_FedShare_CDL, Project=="Schools")

###Downscaling CDL School (downscaled by School age children) ####
TR_TotSchoolAgeChildren_5_17_E <-TR_DemoData$TR_TotSchoolAgeChildren_5_17_E[TR_DemoData$Year== "2012"]
OB_TotSchoolAgeChildren_5_17_E <-  OB_DemoData$OB_TotSchoolAgeChildren_5_17_E[OB_DemoData$Year== "2012"]

###Estimate CDL School disbursed per child #### 
TR_FedShare_CDL_School$CDL_School_Disbursed_Per_Capita <- TR_FedShare_CDL_School$Federal.Disbursed.Amount.Adj / TR_TotSchoolAgeChildren_5_17_E

###Estimate CDL School funds disbursed to Ortley Beach #### 
TR_FedShare_CDL_School$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_CDL_School$CDL_School_Disbursed_Per_Capita * OB_TotSchoolAgeChildren_5_17_E 

###Separate downscaled OB estimates for Debris Removal projects as standalone table####
OB_FedShare_CDL_School <- TR_FedShare_CDL_School %>% select(
  "As.Of.Date",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_CDL_School) <- c("Year", "OB_FedShare_CDL_School_adj")

##Downscaling CDL Services by total population####
###Subset CDL to Services####
TR_FedShare_CDL_Services <- subset(TR_FedShare_CDL, Project=="Services")

###Downscaling CDL Services (downscaled by total population) ####
TR_tot_pop_estimate <-  TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-  OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate CDL Services disbursed per child #### 
TR_FedShare_CDL_Services$CDL_Services_Disbursed_Per_Capita <- TR_FedShare_CDL_Services$Federal.Disbursed.Amount.Adj / TR_tot_pop_estimate

###Estimate CDL Services funds disbursed to Ortley Beach #### 
TR_FedShare_CDL_Services$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_CDL_Services$CDL_Services_Disbursed_Per_Capita * OB_tot_pop_estimate

###Separate downscaled OB estimates for Debris Removal projects as standalone table####
OB_FedShare_CDL_Services <- TR_FedShare_CDL_Services %>% select(
  "As.Of.Date",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_CDL_Services) <- c("Year", "OB_FedShare_CDL_Services_adj")

```

#4) Estimating Federal cost share for repairs to routes 35 and 37 serving Ortley Beach 
Note: this has already downscaled to Ortley beach via text analysis. Results are in OB_FedShare_DOT and OB_FedShare_DOT_LocalAid. 
```{r include=FALSE}
##Splitting to Route 35 & Route 37 ####
OB_FedShare_DOT_Rt35 <- subset(OB_FedShare_DOT, Route==35)
OB_FedShare_DOT_Rt37 <- subset(OB_FedShare_DOT, Route==37)


###Separate downscaled OB estimates for DOT projects as standalone table####
OB_FedShare_DOT_Rt35 <- OB_FedShare_DOT_Rt35 %>% select(
  "Completion.Year",
  "Disbursed.Amount.Adj")

OB_FedShare_DOT_Rt37 <- OB_FedShare_DOT_Rt37 %>% select(
  "Completion.Year",
  "Disbursed.Amount.Adj")

##Aggregating amount by year####
OB_FedShare_DOT_Rt35 <- OB_FedShare_DOT_Rt35 %>%
  group_by(Completion.Year) %>%
  summarise(OB_FedShare_DOT_adj = sum(Disbursed.Amount.Adj, na.rm = TRUE))

OB_FedShare_DOT_Rt37 <- OB_FedShare_DOT_Rt37 %>%
  group_by(Completion.Year) %>%
  summarise(OB_FedShare_DOT_adj = sum(Disbursed.Amount.Adj, na.rm = TRUE))

###Renaming columns####
colnames(OB_FedShare_DOT_Rt35) <- c("Year", "OB_FedShare_DOT_Rt35_adj")

colnames(OB_FedShare_DOT_Rt37) <- c("Year", "OB_FedShare_DOT_Rt37_adj")

##ReducingDOT_LocalAid table####
###Separate downscaled OB estimates DOT_LocalAid projects as standalone table####
OB_FedShare_DOT_LocalAid <- OB_FedShare_DOT_LocalAid %>% select(
  "Year",
  "OB.Est.FedCostShare.Adj")


###Renaming columns####
colnames(OB_FedShare_DOT_LocalAid) <- c("Year", "OB_FedShare_DOT_LocalAid_adj")

```


#5) Estimating Federal cost share for Beach and dune nourishment.
```{r include=FALSE}
#Note: this has already downscaled to Ortley beach via literature review. Reducing to federal contribution

OB_FedShare_BeachNourish <- OB_BeachNourish %>% select(
  "Year",
  "FedStatCostShare_adj"
)

###Renaming columns####
colnames(OB_FedShare_BeachNourish) <- c("Year", "OB_FedShare_BeachNourish_adj")

```


#Knitting tables together to obtain Toms River Municipal Revenues from Ortley Beach by Year
```{r include=FALSE}
library(purrr)
list_of_dfs <- list(OB_PropTaxRevTo_TR_Muni,
                    OB_PropTaxRevTo_TR_School,
                    OB_FedShare_DCA_LGS_LMI,
                    OB_FedShare_DCA_LGS_UN,
                    OB_FedShare_DCA_LPS,
                    OB_FedShare_DCA_ZCEG_UN,
                    OB_FedShare_HMGP,
                    OB_FedShare_PA_Debris,
                    OB_FedShare_PA_Protect,
                    OB_FedShare_PA_RoadsBridges,
                    OB_FedShare_PA_PublicBuildings,
                    OB_FedShare_PA_PublicUtilities,
                    OB_FedShare_PA_Recreational,
                    OB_FedShare_CDL_Muni,
                    OB_FedShare_CDL_School,
                    OB_FedShare_CDL_Services,
                    OB_FedShare_DOT_Rt35,
                    OB_FedShare_DOT_Rt37,
                    OB_FedShare_DOT_LocalAid,
                    OB_FedShare_BeachNourish)

MuniRevOB <- reduce(list_of_dfs, function(x, y) left_join(x, y, by = "Year"))

#dropping 2011
MuniRevOB <- MuniRevOB %>% 
  filter(Year !=2011)

```

#Exporting S1 OB input tables created so far:
```{r include=FALSE}
path <- 'C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/BaselineData/S1_OB_Inputs/Revenues_ToTR_fromOB'

write.csv(MuniRevOB, file.path(path, "MuniRevOB.csv"), row.names = TRUE)


```
