---
title: "S2_Base_MuniRevFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for benefits to Toms River for Scenario 2. We discount the benefits and group them into buckets for analysis.

Assumptions for S2:
-No municipal property tax revenue from OB beyond 2012
-No school property tax revenue from OB beyond 2012
-Limited federal response and recovery (essentials only). Retaining only:
    -PA Debris Removal (2012, 2013)
    -PA Protective measures (2012, 2013)
    -Repairs to Route 37 (2012, 2013)
    
    
-Federal cost share for Route 37 only to enable debris removal (not route 35)
-No federal cost share for Beach and dune nourishment

We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniRevOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Revenues_ToTR_fromOB/MuniRevOB.csv')
MuniRevOB <- MuniRevOB[,c(-1)]

```

# Removing property tax revenue post 2012 (properties bought out)

```{r include = FALSE}
S2_MuniRevOB <- MuniRevOB
S2_MuniRevOB$OB_PropTaxRevTo_TR_Muni_adj[S2_MuniRevOB$Year > 2012] <- NA
S2_MuniRevOB$OB_PropTaxRevTo_TR_School_adj[S2_MuniRevOB$Year > 2012] <- NA

```

# Retaining only essential response and recovery costs
###Limited federal response and recovery (essentials only). Retaining only:
    - PA Debris Removal (2012, 2013)
    - PA Protective measures (2012, 2013)
    - Repairs to Route 37 (2012, 2013)

```{r include = FALSE}
S2_MuniRevOB$OB_FedShare_PA_Debris_adj[S2_MuniRevOB$Year > 2013] <- NA
S2_MuniRevOB$OB_FedShare_PA_Protect_adj[S2_MuniRevOB$Year > 2013] <- NA
S2_MuniRevOB$OB_FedShare_DOT_Rt37_adj[S2_MuniRevOB$Year > 2013] <- NA


#removing other columns
S2_MuniRevOB <- S2_MuniRevOB %>%
  select(Year,
         OB_PropTaxRevTo_TR_Muni_adj,
         OB_PropTaxRevTo_TR_School_adj,
         OB_FedShare_PA_Debris_adj,
         OB_FedShare_PA_Protect_adj,
         OB_FedShare_DOT_Rt37_adj)

```




#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S2_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniRevOB), value = TRUE)

# Apply the discount function
S2_MuniRevOB_dis2pct <- dis2pct(S2_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniRevOB_dis2pct <- S2_MuniRevOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniRevOB_dis2pct_NPV <- S2_MuniRevOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S2_MuniRevOB_dis2pct_NPV<- S2_MuniRevOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis2pct= 
           #OB_FedShare_DCA_LGS_LMI_adj_dis2pct +
           #OB_FedShare_DCA_LGS_UN_adj_dis2pct +
           #OB_FedShare_DCA_LPS_adj_dis2pct +
           #OB_FedShare_DCA_ZCEG_UN_adj_dis2pct +
           #OB_FedShare_HMGP_adj_dis2pct +
           OB_FedShare_PA_Debris_adj_dis2pct +
           OB_FedShare_PA_Protect_adj_dis2pct)
           #OB_FedShare_PA_PublicBuildings_adj_dis2pct +
           #OB_FedShare_PA_PublicUtilities_adj_dis2pct +
           #OB_FedShare_PA_Recreational_adj_dis2pct +
           #OB_FedShare_CDL_Muni_adj_dis2pct +
           #OB_FedShare_CDL_School_adj_dis2pct +
           #OB_FedShare_CDL_Services_adj_dis2pct)
                                                          
#Summing Federal Flows for Transportation
S2_MuniRevOB_dis2pct_NPV<- S2_MuniRevOB_dis2pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis2pct= 
           #OB_FedShare_PA_RoadsBridges_adj_dis2pct +
           #OB_FedShare_DOT_Rt25_adj_dis2pct +
           OB_FedShare_DOT_Rt37_adj_dis2pct)
           #OB_FedShare_DOT_LocalAid_adj_dis2pct)

#selecting variables
S2_MuniRevOB_dis2pct_NPV <- S2_MuniRevOB_dis2pct_NPV %>%
  select(OB_PropTaxRevTo_TR_Muni_adj_dis2pct,
         OB_PropTaxRevTo_TR_School_adj_dis2pct,
         OB_FedShare_ResponseRecovery_dis2pct,
         OB_FedShare_RoadsBridges_dis2pct)
         #OB_FedShare_BeachNourish_adj_dis2pct)

#pivoting tables longer
S2_MuniRevOB_dis2pct_NPV <- S2_MuniRevOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```




#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S2_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniRevOB), value = TRUE)

# Apply the discount function
S2_MuniRevOB_dis3pct <- dis3pct(S2_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniRevOB_dis3pct <- S2_MuniRevOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
S2_MuniRevOB_dis3pct_NPV <- S2_MuniRevOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S2_MuniRevOB_dis3pct_NPV<- S2_MuniRevOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis3pct= 
           #OB_FedShare_DCA_LGS_LMI_adj_dis3pct +
           #OB_FedShare_DCA_LGS_UN_adj_dis3pct +
           #OB_FedShare_DCA_LPS_adj_dis3pct +
           #OB_FedShare_DCA_ZCEG_UN_adj_dis3pct +
           #OB_FedShare_HMGP_adj_dis3pct +
           OB_FedShare_PA_Debris_adj_dis3pct +
           OB_FedShare_PA_Protect_adj_dis3pct)
           #OB_FedShare_PA_PublicBuildings_adj_dis3pct +
           #OB_FedShare_PA_PublicUtilities_adj_dis3pct +
           #OB_FedShare_PA_Recreational_adj_dis3pct +
           #OB_FedShare_CDL_Muni_adj_dis3pct +
           #OB_FedShare_CDL_School_adj_dis3pct +
           #OB_FedShare_CDL_Services_adj_dis3pct)
                                                          
#Summing Federal Flows for Transportation
S2_MuniRevOB_dis3pct_NPV<- S2_MuniRevOB_dis3pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis3pct= 
           #OB_FedShare_PA_RoadsBridges_adj_dis3pct +
           #OB_FedShare_DOT_Rt35_adj_dis3pct +
           OB_FedShare_DOT_Rt37_adj_dis3pct)
           #OB_FedShare_DOT_LocalAid_adj_dis3pct)

#selecting variables
S2_MuniRevOB_dis3pct_NPV <- S2_MuniRevOB_dis3pct_NPV %>%
  select(OB_PropTaxRevTo_TR_Muni_adj_dis3pct,
         OB_PropTaxRevTo_TR_School_adj_dis3pct,
         OB_FedShare_ResponseRecovery_dis3pct,
         OB_FedShare_RoadsBridges_dis3pct)
         #OB_FedShare_BeachNourish_adj_dis3pct)

#pivoting tables longer
S2_MuniRevOB_dis3pct_NPV <- S2_MuniRevOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```

#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
S2_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniRevOB), value = TRUE)

# Apply the discount function
S2_MuniRevOB_dis6pct <- dis6pct(S2_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniRevOB_dis6pct <- S2_MuniRevOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniRevOB_dis6pct_NPV <- S2_MuniRevOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S2_MuniRevOB_dis6pct_NPV<- S2_MuniRevOB_dis6pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis6pct= 
           #OB_FedShare_DCA_LGS_LMI_adj_dis6pct +
           #OB_FedShare_DCA_LGS_UN_adj_dis6pct +
           #OB_FedShare_DCA_LPS_adj_dis6pct +
           #OB_FedShare_DCA_ZCEG_UN_adj_dis6pct +
           #OB_FedShare_HMGP_adj_dis6pct +
           OB_FedShare_PA_Debris_adj_dis6pct +
           OB_FedShare_PA_Protect_adj_dis6pct)
           #OB_FedShare_PA_PublicBuildings_adj_dis6pct +
           #OB_FedShare_PA_PublicUtilities_adj_dis6pct +
           #OB_FedShare_PA_Recreational_adj_dis6pct +
           #OB_FedShare_CDL_Muni_adj_dis6pct +
           #OB_FedShare_CDL_School_adj_dis6pct +
           #OB_FedShare_CDL_Services_adj_dis6pct)
                                                          
#Summing Federal Flows for Transportation
S2_MuniRevOB_dis6pct_NPV<- S2_MuniRevOB_dis6pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis6pct= 
           #OB_FedShare_PA_RoadsBridges_adj_dis6pct +
           #OB_FedShare_DOT_Rt35_adj_dis6pct +
           OB_FedShare_DOT_Rt37_adj_dis6pct)
           #OB_FedShare_DOT_LocalAid_adj_dis6pct)

#selecting variables
S2_MuniRevOB_dis6pct_NPV <- S2_MuniRevOB_dis6pct_NPV %>%
  select(OB_PropTaxRevTo_TR_Muni_adj_dis6pct,
         OB_PropTaxRevTo_TR_School_adj_dis6pct,
         OB_FedShare_ResponseRecovery_dis6pct,
         OB_FedShare_RoadsBridges_dis6pct)
         #OB_FedShare_BeachNourish_adj_dis6pct)

#pivoting tables longer
S2_MuniRevOB_dis6pct_NPV <- S2_MuniRevOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S2_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniRevOB), value = TRUE)

# Apply the discount function
S2_MuniRevOB_dis7pct <- dis7pct(S2_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniRevOB_dis7pct <- S2_MuniRevOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniRevOB_dis7pct_NPV <- S2_MuniRevOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S2_MuniRevOB_dis7pct_NPV<- S2_MuniRevOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis7pct= 
           #OB_FedShare_DCA_LGS_LMI_adj_dis7pct +
           #OB_FedShare_DCA_LGS_UN_adj_dis7pct +
           #OB_FedShare_DCA_LPS_adj_dis7pct +
           #OB_FedShare_DCA_ZCEG_UN_adj_dis7pct +
           #OB_FedShare_HMGP_adj_dis7pct +
           OB_FedShare_PA_Debris_adj_dis7pct +
           OB_FedShare_PA_Protect_adj_dis7pct)
           #OB_FedShare_PA_PublicBuildings_adj_dis7pct +
           #OB_FedShare_PA_PublicUtilities_adj_dis7pct +
           #OB_FedShare_PA_Recreational_adj_dis7pct +
           #OB_FedShare_CDL_Muni_adj_dis7pct +
           #OB_FedShare_CDL_School_adj_dis7pct +
           #OB_FedShare_CDL_Services_adj_dis7pct)
                                                          
#Summing Federal Flows for Transportation
S2_MuniRevOB_dis7pct_NPV<- S2_MuniRevOB_dis7pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis7pct= 
           #OB_FedShare_PA_RoadsBridges_adj_dis7pct +
           #OB_FedShare_DOT_Rt35_adj_dis7pct +
           OB_FedShare_DOT_Rt37_adj_dis7pct)
           #OB_FedShare_DOT_LocalAid_adj_dis7pct)

#selecting variables
S2_MuniRevOB_dis7pct_NPV <- S2_MuniRevOB_dis7pct_NPV %>%
  select(OB_PropTaxRevTo_TR_Muni_adj_dis7pct,
         OB_PropTaxRevTo_TR_School_adj_dis7pct,
         OB_FedShare_ResponseRecovery_dis7pct,
         OB_FedShare_RoadsBridges_dis7pct)
         #OB_FedShare_BeachNourish_adj_dis7pct)

#pivoting tables longer
S2_MuniRevOB_dis7pct_NPV <- S2_MuniRevOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```




#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S2_MuniRevOB_NPV <- S2_MuniRevOB_dis2pct_NPV %>%
  left_join(S2_MuniRevOB_dis3pct_NPV, by = "variable") %>%
  left_join(S2_MuniRevOB_dis6pct_NPV, by = "variable") %>%
  left_join(S2_MuniRevOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S2_OB_Inputs'

write.csv(S2_MuniRevOB_NPV, file.path(path, "S2_MuniRevOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S2_MuniRevOB_dis2pct_NPV <- sum(S2_MuniRevOB_NPV$dis2pct)
S2_MuniRevOB_dis3pct_NPV <- sum(S2_MuniRevOB_NPV$dis3pct)
S2_MuniRevOB_dis6pct_NPV <- sum(S2_MuniRevOB_NPV$dis6pct)
S2_MuniRevOB_dis7pct_NPV <- sum(S2_MuniRevOB_NPV$dis7pct)

S2_MuniRevOB_dis2pct_NPV
S2_MuniRevOB_dis3pct_NPV
S2_MuniRevOB_dis6pct_NPV
S2_MuniRevOB_dis7pct_NPV
```