---
title: "S4_Base_MuniRevFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for benefits to Toms River for Scenario 4. We discount the benefits and group them into buckets for analysis.

Assumptions for S4:
  - Property Tax revenue (Muni): After 2013, only for remaining properties
  - Property Tax revenue (school): After 2013, only for remaining properties
  - Federal cost share for response & recovery: Assume Sticky expenditures (same as S1) 
  - Federal cost share for roads and bridges: Assume Sticky expenditures (same as S1) 
  - Federal cost share for beach nourishment: Assume doubles (doubling beach profile) 


We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniRevOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Revenues_ToTR_fromOB/MuniRevOB.csv')
MuniRevOB <- MuniRevOB[,c(-1)]
OB_S4_M4_TS <- read.csv('./data/BaselineData/OB_MOD4_Produced/TimeSeries/S4/OB_S4_MOD4_TimeSeries.csv')

InflationAdjusters <- read.csv('./data/BaselineData/InflationAdjusters.csv')
TR_TaxData <- read.csv('./data/BaselineData/TomsRiver_TaxData.csv')

```



#Adjusting inputs for property tax revenue####

```{r include=FALSE}
#Obtaining Estimated Property Tax data lost due to buyout cluster ####
S4_MuniRevOB <- MuniRevOB

#clearing out property value tax data post 2012 to make way for new inputs based on cluster
S4_MuniRevOB$OB_PropTaxRevTo_TR_Muni_adj[S4_MuniRevOB$Year > 2012] <- NA
S4_MuniRevOB$OB_PropTaxRevTo_TR_School_adj[S4_MuniRevOB$Year > 2012] <- NA

#Obtaining improvement value for remaining properties in cluster 1 for years 2013-2022
OB_S4 <- OB_S4_M4_TS %>%
  select(ends_with(c("_improvement_value")))

#imputing missing data based on means
colSums(is.na(OB_S4))
OB_S4_imp <- OB_S4 %>%
  mutate(across(ends_with("_improvement_value"), 
                ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

#creating totals by year
OB_S4_y13_TotImpVal <- sum(OB_S4_imp$y13_improvement_value)
OB_S4_y14_TotImpVal <- sum(OB_S4_imp$y14_improvement_value)
OB_S4_y15_TotImpVal <- sum(OB_S4_imp$y15_improvement_value)
OB_S4_y16_TotImpVal <- sum(OB_S4_imp$y16_improvement_value)
OB_S4_y17_TotImpVal <- sum(OB_S4_imp$y17_improvement_value)
OB_S4_y18_TotImpVal <- sum(OB_S4_imp$y18_improvement_value)
OB_S4_y19_TotImpVal <- sum(OB_S4_imp$y19_improvement_value)
OB_S4_y20_TotImpVal <- sum(OB_S4_imp$y20_improvement_value)
OB_S4_y21_TotImpVal <- sum(OB_S4_imp$y21_improvement_value)
OB_S4_y22_TotImpVal <- sum(OB_S4_imp$y22_improvement_value)

TotImpVal <- c(OB_S4_y13_TotImpVal, OB_S4_y14_TotImpVal, OB_S4_y15_TotImpVal, OB_S4_y16_TotImpVal, OB_S4_y17_TotImpVal, OB_S4_y18_TotImpVal, OB_S4_y19_TotImpVal, OB_S4_y20_TotImpVal, OB_S4_y21_TotImpVal, OB_S4_y22_TotImpVal)

OB_S4_A <- data.frame(
  Year = 2013:2022,
  OB_S4_TotalImprovementValue = TotImpVal)


#adjusting improvement value based on equalization ratio and SLIPD
## reducing tables for join
TR_TaxData <- TR_TaxData %>%
  select(Year, TR_Eq_Ratio, TR_TotalMuniTaxRate, TR_SchoolTaxRate)
SLIPD <- InflationAdjusters %>%
  select(Year, SLIPD_V3_Multiplier)

## joining tables
OB_S4_B <- left_join(OB_S4_A, TR_TaxData, by="Year")
OB_S4_B <- left_join(OB_S4_B, SLIPD, by="Year")

## adjusting improvement values by equalization ratio and SLIPD
OB_S4_B$OB_S4_TotalImpVal_adj <- OB_S4_B$OB_S4_TotalImprovementValue * OB_S4_B$TR_Eq_Ratio * OB_S4_B$SLIPD_V3_Multiplier

## Obtaining Net adjusted municipal benefits from property taxes LOST from buyout cluster
OB_S4_B$OB_S4_PropTaxRevLOST_toTR_Muni_adj <- OB_S4_B$OB_S4_TotalImpVal_adj * OB_S4_B$TR_TotalMuniTaxRate

## Obtaining Net adjusted school benefits from property taxes LOST from buyout cluster
OB_S4_B$OB_S4_PropTaxRevLOST_toTR_School_adj <- OB_S4_B$OB_S4_TotalImpVal_adj * OB_S4_B$TR_SchoolTaxRate

# Obtaining Estimated property taxes from remaining properties ####
OB_S4_C <- OB_S4_B %>% select(
  Year,OB_S4_PropTaxRevLOST_toTR_Muni_adj,OB_S4_PropTaxRevLOST_toTR_School_adj)

#joining cluster data to original data
OB_S4_D <- left_join(MuniRevOB, OB_S4_C, by="Year")

#subtracting cluster revenues from total
OB_S4_D$OB_S4_PropTaxRev_toTR_Muni_adj <- OB_S4_D$OB_PropTaxRevTo_TR_Muni_adj - OB_S4_D$OB_S4_PropTaxRevLOST_toTR_Muni_adj

OB_S4_D$OB_S4_PropTaxRev_toTR_School_adj <- OB_S4_D$OB_PropTaxRevTo_TR_School_adj - OB_S4_D$OB_S4_PropTaxRevLOST_toTR_School_adj

#keeping original 2012 data
OB_S4_D$OB_S4_PropTaxRev_toTR_Muni_adj[OB_S4_D$Year == 2012] <- OB_S4_D$OB_PropTaxRevTo_TR_Muni_adj[OB_S4_D$Year == 2012]

OB_S4_D$OB_S4_PropTaxRev_toTR_School_adj[OB_S4_D$Year == 2012] <- OB_S4_D$OB_PropTaxRevTo_TR_School_adj[OB_S4_D$Year == 2012]

#Dropping unneccessary columns
OB_S4_E <- OB_S4_D %>%
  select(-OB_PropTaxRevTo_TR_Muni_adj,
         -OB_PropTaxRevTo_TR_School_adj,
         -OB_S4_PropTaxRevLOST_toTR_Muni_adj,
         -OB_S4_PropTaxRevLOST_toTR_School_adj)


```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}
S4_MuniRevOB <- OB_S4_E

###discounting flows at 2% ####
#Creating a year index 
S4_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_MuniRevOB), value = TRUE)

# Apply the discount function
S4_MuniRevOB_dis2pct <- dis2pct(S4_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_MuniRevOB_dis2pct <- S4_MuniRevOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S4_MuniRevOB_dis2pct_NPV <- S4_MuniRevOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S4_MuniRevOB_dis2pct_NPV<- S4_MuniRevOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis2pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis2pct +
           OB_FedShare_DCA_LGS_UN_adj_dis2pct +
           OB_FedShare_DCA_LPS_adj_dis2pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis2pct +
           OB_FedShare_HMGP_adj_dis2pct +
           OB_FedShare_PA_Debris_adj_dis2pct +
           OB_FedShare_PA_Protect_adj_dis2pct +
           OB_FedShare_PA_PublicBuildings_adj_dis2pct +
           OB_FedShare_PA_PublicUtilities_adj_dis2pct +
           OB_FedShare_PA_Recreational_adj_dis2pct +
           OB_FedShare_CDL_Muni_adj_dis2pct +
           OB_FedShare_CDL_School_adj_dis2pct +
           OB_FedShare_CDL_Services_adj_dis2pct)
                                                          
#Summing Federal Flows for Transportation
S4_MuniRevOB_dis2pct_NPV<- S4_MuniRevOB_dis2pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis2pct= 
           OB_FedShare_PA_RoadsBridges_adj_dis2pct +
           OB_FedShare_DOT_Rt35_adj_dis2pct +
           OB_FedShare_DOT_Rt37_adj_dis2pct +
           OB_FedShare_DOT_LocalAid_adj_dis2pct)

#selecting variables


S4_MuniRevOB_dis2pct_NPV <- S4_MuniRevOB_dis2pct_NPV %>%
  select(OB_S4_PropTaxRev_toTR_Muni_adj_dis2pct,
         OB_S4_PropTaxRev_toTR_School_adj_dis2pct,
         OB_FedShare_ResponseRecovery_dis2pct,
         OB_FedShare_RoadsBridges_dis2pct,
         OB_FedShare_BeachNourish_adj_dis2pct)

#pivoting tables longer
S4_MuniRevOB_dis2pct_NPV <- S4_MuniRevOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```




#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S4_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
diS4pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_MuniRevOB), value = TRUE)

# Apply the discount function
S4_MuniRevOB_dis3pct <- diS4pct(S4_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_MuniRevOB_dis3pct <- S4_MuniRevOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
S4_MuniRevOB_dis3pct_NPV <- S4_MuniRevOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S4_MuniRevOB_dis3pct_NPV<- S4_MuniRevOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis3pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis3pct +
           OB_FedShare_DCA_LGS_UN_adj_dis3pct +
           OB_FedShare_DCA_LPS_adj_dis3pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis3pct +
           OB_FedShare_HMGP_adj_dis3pct +
           OB_FedShare_PA_Debris_adj_dis3pct +
           OB_FedShare_PA_Protect_adj_dis3pct +
           OB_FedShare_PA_PublicBuildings_adj_dis3pct +
           OB_FedShare_PA_PublicUtilities_adj_dis3pct +
           OB_FedShare_PA_Recreational_adj_dis3pct +
           OB_FedShare_CDL_Muni_adj_dis3pct +
           OB_FedShare_CDL_School_adj_dis3pct +
           OB_FedShare_CDL_Services_adj_dis3pct)
                                                          
#Summing Federal Flows for Transportation
S4_MuniRevOB_dis3pct_NPV<- S4_MuniRevOB_dis3pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis3pct= 
           OB_FedShare_PA_RoadsBridges_adj_dis3pct +
           OB_FedShare_DOT_Rt35_adj_dis3pct +
           OB_FedShare_DOT_Rt37_adj_dis3pct +
           OB_FedShare_DOT_LocalAid_adj_dis3pct)

#selecting variables
S4_MuniRevOB_dis3pct_NPV <- S4_MuniRevOB_dis3pct_NPV %>%
  select(OB_S4_PropTaxRev_toTR_Muni_adj_dis3pct,
         OB_S4_PropTaxRev_toTR_School_adj_dis3pct,
         OB_FedShare_ResponseRecovery_dis3pct,
         OB_FedShare_RoadsBridges_dis3pct,
         OB_FedShare_BeachNourish_adj_dis3pct)

#pivoting tables longer
S4_MuniRevOB_dis3pct_NPV <- S4_MuniRevOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
S4_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_MuniRevOB), value = TRUE)

# Apply the discount function
S4_MuniRevOB_dis6pct <- dis6pct(S4_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_MuniRevOB_dis6pct <- S4_MuniRevOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (6060USD)
S4_MuniRevOB_dis6pct_NPV <- S4_MuniRevOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S4_MuniRevOB_dis6pct_NPV<- S4_MuniRevOB_dis6pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis6pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis6pct +
           OB_FedShare_DCA_LGS_UN_adj_dis6pct +
           OB_FedShare_DCA_LPS_adj_dis6pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis6pct +
           OB_FedShare_HMGP_adj_dis6pct +
           OB_FedShare_PA_Debris_adj_dis6pct +
           OB_FedShare_PA_Protect_adj_dis6pct +
           OB_FedShare_PA_PublicBuildings_adj_dis6pct +
           OB_FedShare_PA_PublicUtilities_adj_dis6pct +
           OB_FedShare_PA_Recreational_adj_dis6pct +
           OB_FedShare_CDL_Muni_adj_dis6pct +
           OB_FedShare_CDL_School_adj_dis6pct +
           OB_FedShare_CDL_Services_adj_dis6pct)
                                                          
#Summing Federal Flows for Transportation
S4_MuniRevOB_dis6pct_NPV<- S4_MuniRevOB_dis6pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis6pct= 
           OB_FedShare_PA_RoadsBridges_adj_dis6pct +
           OB_FedShare_DOT_Rt35_adj_dis6pct +
           OB_FedShare_DOT_Rt37_adj_dis6pct +
           OB_FedShare_DOT_LocalAid_adj_dis6pct)

#selecting variables
S4_MuniRevOB_dis6pct_NPV <- S4_MuniRevOB_dis6pct_NPV %>%
  select(OB_S4_PropTaxRev_toTR_Muni_adj_dis6pct,
         OB_S4_PropTaxRev_toTR_School_adj_dis6pct,
         OB_FedShare_ResponseRecovery_dis6pct,
         OB_FedShare_RoadsBridges_dis6pct,
         OB_FedShare_BeachNourish_adj_dis6pct)


#pivoting tables longer
S4_MuniRevOB_dis6pct_NPV <- S4_MuniRevOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S4_MuniRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_MuniRevOB), value = TRUE)

# Apply the discount function
S4_MuniRevOB_dis7pct <- dis7pct(S4_MuniRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_MuniRevOB_dis7pct <- S4_MuniRevOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
S4_MuniRevOB_dis7pct_NPV <- S4_MuniRevOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Federal Flows for Response and Recovery
S4_MuniRevOB_dis7pct_NPV<- S4_MuniRevOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ResponseRecovery_dis7pct= 
           OB_FedShare_DCA_LGS_LMI_adj_dis7pct +
           OB_FedShare_DCA_LGS_UN_adj_dis7pct +
           OB_FedShare_DCA_LPS_adj_dis7pct +
           OB_FedShare_DCA_ZCEG_UN_adj_dis7pct +
           OB_FedShare_HMGP_adj_dis7pct +
           OB_FedShare_PA_Debris_adj_dis7pct +
           OB_FedShare_PA_Protect_adj_dis7pct +
           OB_FedShare_PA_PublicBuildings_adj_dis7pct +
           OB_FedShare_PA_PublicUtilities_adj_dis7pct +
           OB_FedShare_PA_Recreational_adj_dis7pct +
           OB_FedShare_CDL_Muni_adj_dis7pct +
           OB_FedShare_CDL_School_adj_dis7pct +
           OB_FedShare_CDL_Services_adj_dis7pct)
                                                          
#Summing Federal Flows for Transportation
S4_MuniRevOB_dis7pct_NPV<- S4_MuniRevOB_dis7pct_NPV %>%
  mutate(OB_FedShare_RoadsBridges_dis7pct= 
           OB_FedShare_PA_RoadsBridges_adj_dis7pct +
           OB_FedShare_DOT_Rt35_adj_dis7pct +
           OB_FedShare_DOT_Rt37_adj_dis7pct +
           OB_FedShare_DOT_LocalAid_adj_dis7pct)

#selecting variables
S4_MuniRevOB_dis7pct_NPV <- S4_MuniRevOB_dis7pct_NPV %>%
  select(OB_S4_PropTaxRev_toTR_Muni_adj_dis7pct,
         OB_S4_PropTaxRev_toTR_School_adj_dis7pct,
         OB_FedShare_ResponseRecovery_dis7pct,
         OB_FedShare_RoadsBridges_dis7pct,
         OB_FedShare_BeachNourish_adj_dis7pct)

#pivoting tables longer
S4_MuniRevOB_dis7pct_NPV <- S4_MuniRevOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```




#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S4_MuniRevOB_NPV <- S4_MuniRevOB_dis2pct_NPV %>%
  left_join(S4_MuniRevOB_dis3pct_NPV, by = "variable") %>%
  left_join(S4_MuniRevOB_dis6pct_NPV, by = "variable") %>%
  left_join(S4_MuniRevOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S4_OB_Inputs'

write.csv(S4_MuniRevOB_NPV, file.path(path, "S4_MuniRevOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S4_MuniRevOB_dis2pct_NPV <- sum(S4_MuniRevOB_NPV$dis2pct)
S4_MuniRevOB_dis3pct_NPV <- sum(S4_MuniRevOB_NPV$dis3pct)
S4_MuniRevOB_dis6pct_NPV <- sum(S4_MuniRevOB_NPV$dis6pct)
S4_MuniRevOB_dis7pct_NPV <- sum(S4_MuniRevOB_NPV$dis7pct)

S4_MuniRevOB_dis2pct_NPV
S4_MuniRevOB_dis3pct_NPV
S4_MuniRevOB_dis6pct_NPV
S4_MuniRevOB_dis7pct_NPV
```