---
title: "NJST: FEMA CDL Program Overview"
author: "Laura Geronimo"
date: '2024-08-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)

options(scipen=999)

##Importing Data
CDL <- read.csv('./data/DataDownloads/NJ_SandyTransparencyData/Federal_Agency_Datasets/Sandy_CDL_Master_Data.csv')

## Explore####
colSums(is.na(CDL))
table(CDL$Project)

#Workflow:
#########PART 1: Data subset and processing: CPI Adjustments
#Subset data to Ocean County
#Extract Year from a suitable category using library stringr (see example code below)
#create dataframe for CPI Adjustments (year and CPI - see sample code)
#join to your Ocean County dataframe by year (this will integrate CPI multipliers into Ocean county dataframe)
#Create a new variable that reflects adjusted federal cost share (choose suitable variable like federal.distributed.amount, multiply the by the CPI value, and append 'Adj' to variable name --see sample code). Keep track of what values you use by using similar naming convention. (e.g. federal.distributed.amount.adj)
#Create a new variable that reflects adjusted local cost share (choose suitable variable like local.share, multiply the by the CPI value, and append 'Adj' to variable name --see sample code). 
#Spot check your dataframe to make sure that calculations completed

#########PART 2: Redo links with CPI adjusted values. PA cumulative cost share by municipality (local and federal) by project type in Ocean County
    #A. how much CDL assistance (Federal Share Adj) went to each municipality by project type
    #B. How much CDL assistance (Local Share Adj)  went to each municipality by project type

#########PART 3: Obtain generalized costs by unit using assumptions- WORKING on Method
#Note that I am still working on the methodology for select projects
#Subset to Countywide and Toms River to start
#for Toms River, if the project is related to buildings, divide by 9651 (estimate of residential properties affected by Sandy storm surge)
#for Toms River, if the project is related to roads, divide by 134 (estimate of number of miles of roadway affected by Sandy storm surge)
```

### Part 1: Data subset and processing: CPI Adjustments
```{r, include=FALSE}
#subset to Ocean County or Toms River (depending on complexity of your data and ease of processing)
table(is.na(CDL$County))
table(is.na(CDL$Municipality))
CDL_Ocean <- subset(CDL, County=="OCEAN")
CDL_TR <- subset(CDL, Municipality=="TOMS RIVER TOWNSHIP")

#examine missing data
colSums(is.na(CDL_Ocean)) 
colSums(is.na(CDL_TR))

#extracting year
CDL_Ocean$As.Of.Date <- sub(".*(.{4})$", "\\1", CDL_Ocean$As.Of.Date) #note: disbursed data not available
table(CDL_Ocean$As.Of.Date) #missing year on 73/637 projects. Ignore for now.
class(CDL_Ocean$As.Of.Date)
CDL_Ocean$As.Of.Date <- as.numeric(CDL_Ocean$As.Of.Date) #needs to be compatible class for join

#creating dataframe for CPI Adjustments
##adjusting for Inflation####
As.Of.Date<- c(2011,2012,2013,2014,2015,2016,2017,2018,2019,2020,2021,2022,2023)
CPI <-c(1.17,1.14,1.12,1.10,1.10,1.09,1.06,1.04,1.02,1.00,0.99,0.92,0.86)
CPI  <- data.frame(As.Of.Date, CPI)
class(CPI$As.Of.Date)

#join CPI multipliers to Ocean dataframe by year var
CDL_Ocean<-left_join(CDL_Ocean, CPI, by="As.Of.Date", copy=F)
colSums(is.na(CDL_Ocean))

##multiply the $ amounts you are using in links by CPI and create new var, append 'adj' to var
CDL_Ocean$Federal.Disbursed.Amount.Adj <- CDL_Ocean$Disbursed.Amount * CDL_Ocean$CPI

#Spot Check:
QC1<-sum(CDL_Ocean$Disbursed.Amount, na.rm=T)
QC2<-sum(CDL_Ocean$Federal.Disbursed.Amount.Adj, na.rm=T) #should be greater than QC1
QC3<-QC2-QC1 #looks like it processed


```


## Part 2: Redo links with CPI adjusted values.
## Part 2:A: PA cumulative cost share ADJUSTED by municipality (local and federal) by project type in Ocean County
### Part 2.A: PA Assistance (Federal Cost Share Adjusted) by Municipality by Project Type in Ocean County
```{r, include=FALSE}
#Q2.A: CDL_OCEAN adjusted Federal cost share by Project by Municipality

# 2.A.1. nodes --------------------------------------------------------------------
names(CDL_Ocean)

sankey_nodes2A <- list(
  CDL_Ocean %>% select(6) %>% unique,          #Node 1 #Project
  CDL_Ocean %>% select(9) %>% unique) %>%      #Node 2 #Municipality
  unlist(use.names = F) %>% 
  unique()
# 
sankey_nodes2A <- sankey_nodes2A %>% data.frame(names = .)

sankey_nodes2A_df <- sankey_nodes2A %>% 
  mutate(node_id = 0:(nrow(sankey_nodes2A)-1))

# 2.A.2.  Links ADJUSTED-------------------------------------------------------------
names(CDL_Ocean)

link_2A <- CDL_Ocean %>% 
  select(`Project`, Municipality, `Federal.Disbursed.Amount.Adj`) %>% #Use Adjusted version
  group_by(`Project`, Municipality) %>% 
  summarise(value=sum(`Federal.Disbursed.Amount.Adj`,na.rm = T))%>% 
  ungroup()
names(link_2A) <- c('names_src','names_tgt','Federal.Disbursed.Amount.Adj') #keep naming convention for tracking

# 2.A.3. Merging nodes in links
#getting source and target numbers in 
link_2A <- left_join(link_2A, sankey_nodes2A_df, by=c("names_src"= "names"), copy=F)
link_2A <- left_join(link_2A, sankey_nodes2A_df, by=c("names_tgt"= "names"), copy=F)

names(link_2A)
names(link_2A) <- c('names_src','names_tgt','Federal.Disbursed.Amount.Adj', 'source', 'target')
```

```{r echo=FALSE,  message=FALSE, warning=FALSE}
# 2.A.4. Visualization --------------------------------------------------------

p2A <- sankeyNetwork(Links = link_2A, 
                   Nodes = sankey_nodes2A_df, 
                   Source = "source", Target = "target", Value = "Federal.Disbursed.Amount.Adj", 
                   NodeID = "names",
                   units = "$", fontSize = 12, nodeWidth = 30,
                   iterations= 0)
p2A
```

#Creating a subset for Toms River by Year 
##Federal Cost Share by year
```{r, include=FALSE}
CDL_TR <- subset(CDL_Ocean, Municipality=="TOMS RIVER TOWNSHIP")
CDL_TR_FederalShare_ByYear <- CDL_TR %>% 
  select(`Project`, As.Of.Date, `Federal.Disbursed.Amount.Adj`) %>% #Use Adjusted version
  group_by(`Project`, As.Of.Date) %>% 
  summarise(value=sum(`Federal.Disbursed.Amount.Adj`,na.rm = T))%>% 
  ungroup()

names(CDL_TR_FederalShare_ByYear) <- c('Project','As.Of.Date','Federal.Disbursed.Amount.Adj')
CDL_TR_FederalShare_ByYear <- na.omit(CDL_TR_FederalShare_ByYear)

path <- 'C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/BaselineData/NJ_SandyTransparencey_produced'
write.csv(CDL_TR_FederalShare_ByYear, file.path(path, "CDL_TR_FederalShare_ByYear.csv"), row.names = TRUE)
```

##Local Cost Share - NA
