---
title: "S1_FedCostsFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---

The purpose of this worksheet is to estimate the NPV for costs to the Federal Government for Scenario 0. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, and 7% based on justifications provided in the SI.

Simplifying Assumptions:
-Federal Cost Share for Municipal Response and Recovery = NA (no Sandy)
-Federal Cost Share for Individual Response and Recovery = NA (no Sandy)
-Federal Cost Share for Repairs to Roads & Bridges = NA (no Sandy)
-Federal Cost Share for Beach Nourishment = (2021 event only)
-Federal NFIP Claims = Post- Sandy events only





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToFed_fromOB/FedCostOB.csv')
FedCostOB  <- FedCostOB [,c(-1)]

```


#Adjusting baseline data per S0 simplifying assumptions
-Federal Cost Share for Municipal Response and Recovery = NA (no Sandy)
-Federal Cost Share for Individual Response and Recovery = NA (no Sandy)
-Federal Cost Share for Repairs to Roads & Bridges = NA (no Sandy)
-Federal Cost Share for Beach Nourishment = (2021 event only)
-Federal NFIP Claims = Post- Sandy events only

```{r include=FALSE}
#dropping columns
S0_FedCostOB <- FedCostOB %>%
  select(Year,
         OB_FedShare_BeachNourish_adj,
         OB_NFIP_BuildingsClaims_adj,
         OB_NFIP_ContentsClaims_adj,
         OB_NFIP_ICCClaims_adj)

#adjusting Beach nourishment event to 2021 only
S0_FedCostOB$OB_FedShare_BeachNourish_adj[S0_FedCostOB$Year < 2021] <- NA

#adjusting federal claims to post-sandy events only
S0_FedCostOB$OB_NFIP_BuildingsClaims_adj[S0_FedCostOB$Year < 2013] <- NA
S0_FedCostOB$OB_NFIP_ContentsClaims_adj[S0_FedCostOB$Year < 2013] <- NA
S0_FedCostOB$OB_NFIP_ICCClaims_adj[S0_FedCostOB$Year < 2013] <- NA


```



#Discounting values to 2%, summing to NPV & grouping into buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S0_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_FedCostOB), value = TRUE)

# Apply the discount function
S0_FedCostOB_dis2pct <- dis2pct(S0_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_FedCostOB_dis2pct <- S0_FedCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S0_FedCostOB_dis2pct_NPV <- S0_FedCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for NFIP claims
S0_FedCostOB_dis2pct_NPV  <- S0_FedCostOB_dis2pct_NPV  %>%
  mutate(OB_FedShare_NFIP_dis2pct=
           OB_NFIP_BuildingsClaims_adj_dis2pct +
           OB_NFIP_ContentsClaims_adj_dis2pct +
           OB_NFIP_ICCClaims_adj_dis2pct)
  
#selecting variables
S0_FedCostOB_dis2pct_NPV  <- S0_FedCostOB_dis2pct_NPV %>%
  select(OB_FedShare_BeachNourish_adj_dis2pct,
         OB_FedShare_NFIP_dis2pct)

#pivoting tables longer
S0_FedCostOB_dis2pct_NPV <- S0_FedCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV & grouping into buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S0_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_FedCostOB), value = TRUE)

# Apply the discount function
S0_FedCostOB_dis3pct <- dis3pct(S0_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_FedCostOB_dis3pct <- S0_FedCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S0_FedCostOB_dis3pct_NPV <- S0_FedCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for NFIP claims
S0_FedCostOB_dis3pct_NPV  <- S0_FedCostOB_dis3pct_NPV  %>%
  mutate(OB_FedShare_NFIP_dis3pct=
           OB_NFIP_BuildingsClaims_adj_dis3pct +
           OB_NFIP_ContentsClaims_adj_dis3pct +
           OB_NFIP_ICCClaims_adj_dis3pct)
  
#selecting variables
S0_FedCostOB_dis3pct_NPV  <- S0_FedCostOB_dis3pct_NPV %>%
  select(OB_FedShare_BeachNourish_adj_dis3pct,
         OB_FedShare_NFIP_dis3pct)

#pivoting tables longer
S0_FedCostOB_dis3pct_NPV <- S0_FedCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```




#Discounting values to 7%, summing to NPV & grouping into buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S0_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_FedCostOB), value = TRUE)

# Apply the discount function
S0_FedCostOB_dis7pct <- dis7pct(S0_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_FedCostOB_dis7pct <- S0_FedCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S0_FedCostOB_dis7pct_NPV <- S0_FedCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for NFIP claims
S0_FedCostOB_dis7pct_NPV  <- S0_FedCostOB_dis7pct_NPV  %>%
  mutate(OB_FedShare_NFIP_dis7pct=
           OB_NFIP_BuildingsClaims_adj_dis7pct +
           OB_NFIP_ContentsClaims_adj_dis7pct +
           OB_NFIP_ICCClaims_adj_dis7pct)
  
#selecting variables
S0_FedCostOB_dis7pct_NPV  <- S0_FedCostOB_dis7pct_NPV %>%
  select(OB_FedShare_BeachNourish_adj_dis7pct,
         OB_FedShare_NFIP_dis7pct)

#pivoting tables longer
S0_FedCostOB_dis7pct_NPV <- S0_FedCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S0_FedCostOB_NPV <- S0_FedCostOB_dis2pct_NPV %>%
  left_join(S0_FedCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S0_FedCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S0_OB_Inputs'

write.csv(S0_FedCostOB_NPV, file.path(path, "S0_FedCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S0_FedCostOB_dis2pct_NPV <- sum(S0_FedCostOB_NPV$dis2pct)
S0_FedCostOB_dis3pct_NPV <- sum(S0_FedCostOB_NPV$dis3pct)
S0_FedCostOB_dis7pct_NPV <- sum(S0_FedCostOB_NPV$dis7pct)

S0_FedCostOB_dis2pct_NPV
S0_FedCostOB_dis3pct_NPV
S0_FedCostOB_dis7pct_NPV
```