---
title: "Federal Costs from Ortley Beach - Preprocessing"
author: "Laura Geronimo"
date: "2024-08-23"
output: html_document
---
The purpose of this worksheet is to estimate the costs to the Federal Government of maintaining properties on Ortley Beach for Scenario 1 (replicating the historical record). Costs are estimated for the 2012-2022 period. Specifically we estimate:

1) Federal Costs direct to the municipal government for response and recovery (preprocessed)
2) Federal Costs direct to individual households for response and recovery
3) Federal cost share for repairs to roads and bridges (Route 35 & 37) (preprocessed)
4) NFIP Claims from damages to homes in Ortley Beach
5) Federal cost share for beach nourishment (preprocessed)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)

##Importing Data ####
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')


#Data for downscaling
OB_DemoData <- read_xlsx('./data/BaselineData/OrtleyBeach_DemographicData.xlsx', sheet=1)
ZCTA_DemoData <- read_xlsx('./data/BaselineData/ZCTA_08751_DemographicData.xlsx', sheet=1)
TR_DemoData <- read_xlsx('./data/BaselineData/TomsRiver_DemographicData.xlsx', sheet=1)

TR_Sandy <- read_xlsx('./data/BaselineData/TomsRiver_SandyImpacts.xlsx', sheet=1)


#Federal cost share for Municipal response and recovery to OB (already process as municipal revenues and downscaled to Ortley Beach (OB))
FedCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Revenues_ToTR_fromOB/MuniRevOB.csv')

#Federal cost share for Individual response and recovery to OB (needs downscaling)
TR_FedShare_DCA <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/DCA_TR_FederalShare_ByYear.csv')
TR_FedShare_IA <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/IA_TR_FedShare_ByYear.csv')

#Federal NFIP claims paid to OB ZCTA (must be downscaled by number of policies in force)
ZOB_NFIP_Claims <- read.csv('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/DataDownloads/NFIP/Claims/produced/ZCTA_OB_NFIP_Claims.csv')

ZOB_NFIP_PolicyData <- read.csv('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/DataDownloads/NFIP/Policies/produced/ZCTA_08571/NFIP_PolicyData_08571.csv')


```

#1) Federal Costs direct to the municipal government for response and recovery (preprocessed in MuniRevOB)
#3) Federal cost share for repairs to roads and bridges (Route 35 & 37) (preprocessed in MuniRevOB)
#5) Federal cost share for beach nourishment (preprocessed)

```{r include= FALSE}
#selecting relevant columns from pre-processed MuniRevOB file
FedCostOB <- FedCostOB %>%
  select(c(-X,
           -OB_PropTaxRevTo_TR_Muni_adj,
           -OB_PropTaxRevTo_TR_School_adj))
```

#2) Federal Costs direct to individual households for response and recovery

##2a) Downscaling HUD / DCA data for individuals in Ortley Beach by project
```{r include=FALSE}

# List of project names to drop (already processed in MuniRevOB, or focused on rental properties which is less than 3% in OB)
projects_to_drop <- c("DCA LGS Support for Governmental Entities LMI", 
                      "DCA LGS Support for Governmental Entities UN", 
                      "DCA Local Planning Services",
                      "DCA LRRP",
                      "DCA LRRP")

# Drop rows where 'project' is in 'projects_to_drop'
TR_FedShare_DCA <- TR_FedShare_DCA[!(TR_FedShare_DCA$Project %in% projects_to_drop), ]

#Processing Projects ####

##DCA Housing Landlord Incentive Program LMI####
###Subset the DCA projects to DCA Housing Landlord Incentive Program LMI####
TR_FedShare_DCA_Housing_LandLordIncentive_LMI <- subset(TR_FedShare_DCA, Project=="DCA Housing Landlord Incentive Program LMI")

###Downscaling DCA DCA Housing Landlord Incentive Program LMI (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA LMI funds disbursed per capita #### 
TR_FedShare_DCA_Housing_LandLordIncentive_LMI$LMI_Disbursed_Per_Capita <- TR_FedShare_DCA_Housing_LandLordIncentive_LMI$Federal.Disbursed.Amount.Adj / TR_lmi_pop_estimate

###Estimate DCA LMI funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_Housing_LandLordIncentive_LMI$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_Housing_LandLordIncentive_LMI$LMI_Disbursed_Per_Capita * OB_lmi_pop_estimate 

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_FedShare_DCA_Housing_LandLordIncentive_LMI <- TR_FedShare_DCA_Housing_LandLordIncentive_LMI %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_Housing_LandLordIncentive_LMI ) <- c("Year", "OB_FedShare_DCA_Housing_LandLordIncentive_LMI")

##DCA Housing NEP LMI####
###Subset the DCA projects to DCA Housing NEP LMI ####
TR_FedShare_DCA_Housing_NEP_LMI <- subset(TR_FedShare_DCA, Project=="DCA Housing NEP LMI")

###Downscaling DCA Housing NEP LMI (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA LMI funds disbursed per capita #### 
TR_FedShare_DCA_Housing_NEP_LMI$DCA_Disbursed_Per_Capita <- TR_FedShare_DCA_Housing_NEP_LMI$Federal.Disbursed.Amount.Adj / 
TR_lmi_pop_estimate 

###Estimate DCA LMI funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_Housing_NEP_LMI$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_Housing_NEP_LMI$DCA_Disbursed_Per_Capita * OB_lmi_pop_estimate   

###Separate downscaled OB estimates for projects as standalone table####
OB_FedShare_DCA_Housing_NEP_LMI <- TR_FedShare_DCA_Housing_NEP_LMI %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_Housing_NEP_LMI) <- c("Year", "OB_FedShare_DCA_Housing_NEP_LMI")


##DCA LMI Homeowner Rebuilding Program ####
###Subset the DCA LMI Homeowner Rebuilding Program ####
TR_DCA_LMI_HomeownerRebuilding <- subset(TR_FedShare_DCA, Project=="DCA LMI Homeowner Rebuilding Program")

###Downscaling DCA LMI Homeowner Rebuilding Program  (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA LMI Homeowner Rebuilding Program  disbursed per LMI Pop #### 
TR_DCA_LMI_HomeownerRebuilding$DCA_LMI_Disbursed_Per_LMI_pop <- TR_DCA_LMI_HomeownerRebuilding $Federal.Disbursed.Amount.Adj / 
TR_lmi_pop_estimate 

###Estimate DCA LMI Homeowner Rebuilding Program  funds disbursed to Ortley Beach #### 
TR_DCA_LMI_HomeownerRebuilding$Federal.Disbursed.Amount.Adj.OB <- TR_DCA_LMI_HomeownerRebuilding$DCA_LMI_Disbursed_Per_LMI_pop * OB_lmi_pop_estimate  

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_LMI_HomeownerRebuilding <- TR_DCA_LMI_HomeownerRebuilding  %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_LMI_HomeownerRebuilding) <- c("Year", "OB_FedShare_DCA_LMI_HomeownerRebuilding")


##DCA Resettlement LMI####
###Subset the DCA Resettlement LMI####
TR_FedShare_DCA_Resettlement_LMI <- subset(TR_FedShare_DCA, Project=="DCA Resettlement LMI")

###Downscaling DCA Resettlement LMI  (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA  DCA Resettlement LMI disbursed per LMI #### 
TR_FedShare_DCA_Resettlement_LMI$DCA_LMI_Disbursed_Per_LMI_pop <- TR_FedShare_DCA_Resettlement_LMI$Federal.Disbursed.Amount.Adj / TR_lmi_pop_estimate 

###Estimate  DCA Resettlement LMI funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_Resettlement_LMI$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_Resettlement_LMI$DCA_LMI_Disbursed_Per_LMI_pop * OB_lmi_pop_estimate

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_Resettlement_LMI <- TR_FedShare_DCA_Resettlement_LMI %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_Resettlement_LMI) <- c("Year", "OB_FedShare_DCA_Resettlement_LMI")


##DCA Resettlement UN####
###Subset the DCA Resettlement UN####
TR_FedShare_DCA_Resettlement_UN <- subset(TR_FedShare_DCA, Project=="DCA Resettlement UN")

###Downscaling DCA Resettlement UN  (downscaled by estimates of OwnerOccupied Housing Units impacted by Sandy) ####
TR_OwnOccSandy_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_HU_OwnOcc_SandyImpact"]
OB_OwnOccSandy_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_HU_OwnOcc_SandyImpact"]

###Estimate DCA Local Planning Services disbursed per OwnOccHU #### 
TR_FedShare_DCA_Resettlement_UN$DCA_UN_Disbursed_Per_OwnOccHU <- TR_FedShare_DCA_Resettlement_UN$Federal.Disbursed.Amount.Adj / TR_OwnOccSandy_estimate 

###Estimate DCA Local Planning Services funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_Resettlement_UN$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_Resettlement_UN$DCA_UN_Disbursed_Per_OwnOccHU * OB_OwnOccSandy_estimate

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_Resettlement_UN <- TR_FedShare_DCA_Resettlement_UN %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_Resettlement_UN) <- c("Year", "OB_FedShare_DCA_Resettlement_UN")


##DCA RREM LMI####
###Subset the RREM LMI####
TR_FedShare_DCA_RREM_LMI <- subset(TR_FedShare_DCA, Project=="DCA RREM LMI")

###Downscaling RREM LMI  (downscaled by LMI pop) ####
TR_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LMI_Pop_SandyImpact"]
OB_lmi_pop_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LMI_Pop_SandyImpact"]

###Estimate DCA Local Planning Services disbursed per capita #### 
TR_FedShare_DCA_RREM_LMI$DCA_UN_Disbursed_Per_Capita <- TR_FedShare_DCA_RREM_LMI$Federal.Disbursed.Amount.Adj / TR_lmi_pop_estimate 

###Estimate DCA Local Planning Services funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_RREM_LMI$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_RREM_LMI$DCA_UN_Disbursed_Per_Capita * OB_lmi_pop_estimate 

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_RREM_LMI <- TR_FedShare_DCA_RREM_LMI %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_RREM_LMI) <- c("Year", "OB_FedShare_DCA_RREM_LMI")


##DCA RREM UN####
###Subset the RREM UN####
TR_FedShare_DCA_RREM_UN <- subset(TR_FedShare_DCA, Project=="DCA RREM UN")

###Downscaling RREM UN  (downscaled by estimates of OwnerOccupied Housing Units impacted by Sandy) ####
TR_OwnOccSandy_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_HU_OwnOcc_SandyImpact"]
OB_HU_OwnOccSandy_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_HU_OwnOcc_SandyImpact"]

###Estimate DCA Local Planning Services disbursed per capita #### 
TR_FedShare_DCA_RREM_UN$DCA_UN_Disbursed_Per_Capita <- TR_FedShare_DCA_RREM_UN$Federal.Disbursed.Amount.Adj / TR_OwnOccSandy_estimate 

###Estimate DCA Local Planning Services funds disbursed to Ortley Beach #### 
TR_FedShare_DCA_RREM_UN$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_DCA_RREM_UN$DCA_UN_Disbursed_Per_Capita * OB_HU_OwnOccSandy_estimate

###Separate downscaled OB estimates for Local Planning Service projects as standalone table####
OB_FedShare_DCA_RREM_UN <- TR_FedShare_DCA_RREM_UN %>% select(
  "Disbursed.Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_DCA_RREM_UN) <- c("Year", "OB_FedShare_DCA_RREM_UN")

```

##2a) Downscaling IA data for individuals in Ortley Beach 
```{R include=FALSE}
#Processing IA####
head(TR_FedShare_IA)

###Downscaling IA (downscaled by estimates of owner occupied HUs impacted by Sandy) ####
TR_HU_OwnOcc_SandyImpact_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_HU_OwnOcc_SandyImpact"]
OB_HU_OwnOcc_SandyImpact_estimate <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_HU_OwnOcc_SandyImpact"]

###Estimate IA funds disbursed per HU #### 
TR_FedShare_IA$IA_Disbursed_Per_HU <- TR_FedShare_IA$Disbursed.Amount_adj / TR_HU_OwnOcc_SandyImpact_estimate

###Estimate DCA LMI funds disbursed to Ortley Beach #### 
TR_FedShare_IA$Federal.Disbursed.Amount.Adj.OB <- TR_FedShare_IA$IA_Disbursed_Per_HU * OB_HU_OwnOcc_SandyImpact_estimate

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_FedShare_IA <- TR_FedShare_IA %>% select(
  "Year",
  "Federal.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_FedShare_IA ) <- c("Year", "OB_FedShare_IA")

```


# 4) NFIP Claims from damages to homes in Ortley Beach 
### Downscaling Claims from ZCTA to OB by estimated number of homes with policies in force at the time of storm ####

```{r include=False}
#number of housing units in 2012 in ZCTA
ZCTA_NumberHU2012 <- as.numeric(ZCTA_DemoData$Z_HU[ZCTA_DemoData$Year==2012])

#number of housing units impacted by Sandy in 2012 in Ortley Beach
OB_NumberHU2012 <- as.numeric(TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"])

#number of policies in force in 2012 in ZCTA
ZCTA_NumberPoliciesInForce2012 <-  as.numeric(ZOB_NFIP_PolicyData$numberOfPolicies[ZOB_NFIP_PolicyData$policyEffectiveYear == 2012])


#proportion of Housing Units with policies in force in 2012 in ZCTA
ZCTA_PctPoliciesInForce2012 <- ZCTA_NumberPoliciesInForce2012/ ZCTA_NumberHU2012 

#estimate of number of housing units with policies in force in Ortley Beach in 2012
OB_NumberPoliciesInForce2012_E <- OB_NumberHU2012 * ZCTA_PctPoliciesInForce2012 


#downscaling Claims payouts by estimate of housing units with policies in force
ZOB_NFIP_Claims$OB_NFIP_BuildingsClaims_adj <- ZOB_NFIP_Claims$NFIP_BuildingsClaims_adj / ZCTA_NumberPoliciesInForce2012 * OB_NumberPoliciesInForce2012_E

ZOB_NFIP_Claims$OB_NFIP_ContentsClaims_adj <- ZOB_NFIP_Claims$NFIP_ContentsClaims_adj / ZCTA_NumberPoliciesInForce2012 * OB_NumberPoliciesInForce2012_E

ZOB_NFIP_Claims$OB_NFIP_ICCClaims_adj <- ZOB_NFIP_Claims$NFIP_ICCClaims_adj / ZCTA_NumberPoliciesInForce2012 * OB_NumberPoliciesInForce2012_E

#retaining OB variables
OB_NFIP_Claims <- ZOB_NFIP_Claims %>%
  select(Year,
         OB_NFIP_BuildingsClaims_adj,
         OB_NFIP_ContentsClaims_adj,
         OB_NFIP_ICCClaims_adj)


```


#Knitting tables together to obtain Federal costs from Ortley Beach by Year
```{r include=FALSE}
library(purrr)
list_of_dfs <- list(FedCostOB,
                    OB_FedShare_DCA_Housing_LandLordIncentive_LMI,
                    OB_FedShare_DCA_Housing_NEP_LMI,
                    OB_FedShare_DCA_LMI_HomeownerRebuilding,
                    OB_FedShare_DCA_Resettlement_LMI,
                    OB_FedShare_DCA_Resettlement_UN,
                    OB_FedShare_DCA_RREM_LMI,
                    OB_FedShare_DCA_RREM_UN,
                    OB_FedShare_IA,
                    OB_NFIP_Claims)

FedCostOB <- reduce(list_of_dfs, function(x, y) left_join(x, y, by = "Year"))


```

#Exporting S1 OB input tables created so far:
```{r include=FALSE}
path <- './data/BaselineData/S1_OB_Inputs/Costs_ToFed_fromOB'

write.csv(FedCostOB, file.path(path, "FedCostOB.csv"), row.names = TRUE)


```
