---
title: "S2_FedCostsFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for costs to the Federal Government for Scenario 2. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, and 7% based on justifications provided in the SI.

Assumptions:
-Federal Cost Share for Municipal Response and Recovery (essential only for 2012 and 2013)
    -PA Debris Removal (2012, 2013)
    -PA Protective measures (2012, 2013)
    -Repairs to Route 37 (2012, 2013)
-Federal cost share for response and recovery direct to household (NA: redirected to buyouts)
-Federal IA / NFIP claims (NA: redirected to buyouts)
-Federal cost share for beach nourishment (NA)

-Federal cost share for buyouts = 75% of market price for all homes



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToFed_fromOB/FedCostOB.csv')
FedCostOB <- FedCostOB[,c(-1)]

#for buyout cost estimates
OB_2011_SP <- read.csv('./data/BaselineData/OB_MOD4_Produced/TimeSeries/S1/OB_2011_SalesPrice.csv')
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')

```

#Adjusting basline data per simplifying assumptions for S2
# Retaining essential response and recovery costs from federal government, and reducing to 2012 / 2013
```{r include = FALSE}
#retaining essential lines
S2_FedCostOB <- FedCostOB %>%
  select(Year,
         OB_FedShare_PA_Debris_adj,
         OB_FedShare_PA_Protect_adj,
         OB_FedShare_DOT_Rt37_adj)

#reducing to 2012 & 2013 only
S2_FedCostOB$OB_FedShare_PA_Debris_adj[S2_FedCostOB$Year > 2013] <- NA
S2_FedCostOB$OB_FedShare_PA_Protect_adj[S2_FedCostOB$Year > 2013] <- NA
S2_FedCostOB$OB_FedShare_DOT_Rt37_adj[S2_FedCostOB$Year > 2013] <- NA

```


# Calculating Federal cost share for Buyouts (75 % of market price)
```{r include= FALSE}
#subset estimated sales price to the winzorized estimates
OB_2011_SP <- OB_2011_SP[,c("SalesPrice_LG_W")]

#obtain total sales price value in Ortley Beach
OB_2011_SP_Tot <- sum(OB_2011_SP)


#adjusting by CPI to 2020 USD
y11_CPI_multiplier <- Adjust$CPI_Multiplier_USD2020[Adjust$Year==2011]

OB_2011_SP_Tot_adj <- OB_2011_SP_Tot * y11_CPI_multiplier

#obtaining Federal share of buyout cost (75% cost share)
OB_FedShare_BuyoutCost_adj <- OB_2011_SP_Tot_adj * 0.75

#Adding buyout costs to data frame 
S2_FedCostOB$OB_FedShare_BuyoutCost_adj[S2_FedCostOB$Year==2012] <- OB_FedShare_BuyoutCost_adj
```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S2_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_FedCostOB), value = TRUE)

# Apply the discount function
S2_FedCostOB_dis2pct <- dis2pct(S2_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_FedCostOB_dis2pct <- S2_FedCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_FedCostOB_dis2pct_NPV <- S2_FedCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government
S2_FedCostOB_dis2pct_NPV <- S2_FedCostOB_dis2pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis2pct= 
           OB_FedShare_PA_Debris_adj_dis2pct+
           OB_FedShare_PA_Protect_adj_dis2pct)


#selecting variables
S2_FedCostOB_dis2pct_NPV <- S2_FedCostOB_dis2pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis2pct,
         OB_FedShare_DOT_Rt37_adj_dis2pct,
         OB_FedShare_BuyoutCost_adj_dis2pct)

#pivoting tables longer
S2_FedCostOB_dis2pct_NPV <- S2_FedCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```


#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S2_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_FedCostOB), value = TRUE)

# Apply the discount function
S2_FedCostOB_dis3pct <- dis3pct(S2_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_FedCostOB_dis3pct <- S2_FedCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_FedCostOB_dis3pct_NPV <- S2_FedCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government
S2_FedCostOB_dis3pct_NPV <- S2_FedCostOB_dis3pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis3pct= 
           OB_FedShare_PA_Debris_adj_dis3pct+
           OB_FedShare_PA_Protect_adj_dis3pct)


#selecting variables
S2_FedCostOB_dis3pct_NPV <- S2_FedCostOB_dis3pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis3pct,
         OB_FedShare_DOT_Rt37_adj_dis3pct,
         OB_FedShare_BuyoutCost_adj_dis3pct)

#pivoting tables longer
S2_FedCostOB_dis3pct_NPV <- S2_FedCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S2_FedCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_FedCostOB), value = TRUE)

# Apply the discount function
S2_FedCostOB_dis7pct <- dis7pct(S2_FedCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_FedCostOB_dis7pct <- S2_FedCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_FedCostOB_dis7pct_NPV <- S2_FedCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Fed Share for Response and Recovery Direct to Municipal Government
S2_FedCostOB_dis7pct_NPV <- S2_FedCostOB_dis7pct_NPV %>%
  mutate(OB_FedShare_ToTR_ResponseRecovery_dis7pct= 
           OB_FedShare_PA_Debris_adj_dis7pct+
           OB_FedShare_PA_Protect_adj_dis7pct)


#selecting variables
S2_FedCostOB_dis7pct_NPV <- S2_FedCostOB_dis7pct_NPV %>%
  select(OB_FedShare_ToTR_ResponseRecovery_dis7pct,
         OB_FedShare_DOT_Rt37_adj_dis7pct,
         OB_FedShare_BuyoutCost_adj_dis7pct)

#pivoting tables longer
S2_FedCostOB_dis7pct_NPV <- S2_FedCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S2_FedCostOB_NPV <- S2_FedCostOB_dis2pct_NPV %>%
  left_join(S2_FedCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S2_FedCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S2_OB_Inputs'

write.csv(S2_FedCostOB_NPV, file.path(path, "S2_FedCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S2_FedCostOB_dis2pct_NPV <- sum(S2_FedCostOB_NPV$dis2pct)
S2_FedCostOB_dis3pct_NPV <- sum(S2_FedCostOB_NPV$dis3pct)
S2_FedCostOB_dis7pct_NPV <- sum(S2_FedCostOB_NPV$dis7pct)

S2_FedCostOB_dis2pct_NPV
S2_FedCostOB_dis3pct_NPV
S2_FedCostOB_dis7pct_NPV
```