---
title: "S0_MuniCostFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for costs to Toms River for Scenario 0. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.

Simplifying Assumptions:
-Public works remains constant at pre-storm rate
-Public safety remains constant at pre-storm rate
-School district costs remain constant at pre-storm rate
-Municipal cost share for beach nourishment: retain cost for 2022 event only
-Drop all response and recovery costs


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToTR_fromOB/MuniCostOB.csv')
MuniCostOB <- MuniCostOB[,c(-1)]

```

# Keeping public works, public safety, and school district costs constant at pre-storm rates
```{r include = FALSE}

S0_MuniCostOB <- MuniCostOB

S0_MuniCostOB$OB_Public_Works_adj[S0_MuniCostOB$Year > 2012] <- 
  S0_MuniCostOB$OB_Public_Works_adj[S0_MuniCostOB$Year == 2012]

S0_MuniCostOB$OB_Public_Safety_adj[S0_MuniCostOB$Year > 2012] <- 
  S0_MuniCostOB$OB_Public_Safety_adj[S0_MuniCostOB$Year == 2012]

S0_MuniCostOB$OB_SchoolExpend_adj[S0_MuniCostOB$Year > 2012] <- 
  S0_MuniCostOB$OB_SchoolExpend_adj[S0_MuniCostOB$Year == 2012]

```

#Reducing Municipal cost share for beach nourishment to 2022 event only
```{r include = FALSE}

S0_MuniCostOB$OB_LocalShare_BeachNourish_adj[S0_MuniCostOB$Year < 2022] <- NA

```

#dropping all response and recovery costs
```{r include = FALSE}

S0_MuniCostOB <- S0_MuniCostOB %>% 
  select(Year,
         OB_Public_Works_adj,
         OB_Public_Safety_adj,
         OB_SchoolExpend_adj,
         OB_LocalShare_BeachNourish_adj)

```



#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S0_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_MuniCostOB), value = TRUE)

# Apply the discount function
S0_MuniCostOB_dis2pct <- dis2pct(S0_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_MuniCostOB_dis2pct <- S0_MuniCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S0_MuniCostOB_dis2pct_NPV <- S0_MuniCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

#pivoting tables longer
S0_MuniCostOB_dis2pct_NPV <- S0_MuniCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S0_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_MuniCostOB), value = TRUE)

# Apply the discount function
S0_MuniCostOB_dis3pct <- dis3pct(S0_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_MuniCostOB_dis3pct <- S0_MuniCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
S0_MuniCostOB_dis3pct_NPV <- S0_MuniCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

#pivoting tables longer
S0_MuniCostOB_dis3pct_NPV <- S0_MuniCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
S0_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_MuniCostOB), value = TRUE)

# Apply the discount function
S0_MuniCostOB_dis6pct <- dis6pct(S0_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_MuniCostOB_dis6pct <- S0_MuniCostOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (6060USD)
S0_MuniCostOB_dis6pct_NPV <- S0_MuniCostOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

#pivoting tables longer
S0_MuniCostOB_dis6pct_NPV <- S0_MuniCostOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 

```


#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S0_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S0_MuniCostOB), value = TRUE)

# Apply the discount function
S0_MuniCostOB_dis7pct <- dis7pct(S0_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S0_MuniCostOB_dis7pct <- S0_MuniCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
S0_MuniCostOB_dis7pct_NPV <- S0_MuniCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

#pivoting tables longer
S0_MuniCostOB_dis7pct_NPV <- S0_MuniCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 


```




#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S0_MuniCostOB_NPV <- S0_MuniCostOB_dis2pct_NPV %>%
  left_join(S0_MuniCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S0_MuniCostOB_dis6pct_NPV, by = "variable") %>%
  left_join(S0_MuniCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S0_OB_Inputs'

write.csv(S0_MuniCostOB_NPV, file.path(path, "S0_MuniCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S0_MuniCostOB_dis2pct_NPV <- sum(S0_MuniCostOB_NPV$dis2pct)
S0_MuniCostOB_dis3pct_NPV <- sum(S0_MuniCostOB_NPV$dis3pct)
S0_MuniCostOB_dis6pct_NPV <- sum(S0_MuniCostOB_NPV$dis6pct)
S0_MuniCostOB_dis7pct_NPV <- sum(S0_MuniCostOB_NPV$dis7pct)

S0_MuniCostOB_dis2pct_NPV
S0_MuniCostOB_dis3pct_NPV
S0_MuniCostOB_dis6pct_NPV
S0_MuniCostOB_dis7pct_NPV
```