---
title: "S1_Base_MuniCostFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for costs to Toms River for Scenario 1. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToTR_fromOB/MuniCostOB.csv')
MuniCostOB <- MuniCostOB[,c(-1)]

```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(MuniCostOB), value = TRUE)

# Apply the discount function
MuniCostOB_dis2pct <- dis2pct(MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
MuniCostOB_dis2pct <- MuniCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
MuniCostOB_dis2pct_NPV <- MuniCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
MuniCostOB_dis2pct_NPV<- MuniCostOB_dis2pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_dis2pct= 
           OB_LocalShare_HMGP_adj_dis2pct +
           OB_LocalShare_PA_Debris_adj_dis2pct +
           OB_LocalShare_PA_Protect_adj_dis2pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis2pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis2pct +
           OB_LocalShare_PA_Recreational_adj_dis2pct)
  
#selecting variables
MuniCostOB_dis2pct_NPV <- MuniCostOB_dis2pct_NPV %>%
  select(OB_Public_Safety_adj_dis2pct,
         OB_Public_Works_adj_dis2pct,
         OB_SchoolExpend_adj_dis2pct,
         OB_LocalShare_ResponseRecovery_dis2pct,
         OB_LocalShare_BeachNourish_adj_dis2pct)

#pivoting tables longer
MuniCostOB_dis2pct_NPV <- MuniCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```


#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(MuniCostOB), value = TRUE)

# Apply the discount function
MuniCostOB_dis3pct <- dis3pct(MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
MuniCostOB_dis3pct <- MuniCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
MuniCostOB_dis3pct_NPV <- MuniCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
MuniCostOB_dis3pct_NPV<- MuniCostOB_dis3pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_dis3pct= 
           OB_LocalShare_HMGP_adj_dis3pct +
           OB_LocalShare_PA_Debris_adj_dis3pct +
           OB_LocalShare_PA_Protect_adj_dis3pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis3pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis3pct +
           OB_LocalShare_PA_Recreational_adj_dis3pct)
  
#selecting variables
MuniCostOB_dis3pct_NPV <- MuniCostOB_dis3pct_NPV %>%
  select(OB_Public_Safety_adj_dis3pct,
         OB_Public_Works_adj_dis3pct,
         OB_SchoolExpend_adj_dis3pct,
         OB_LocalShare_ResponseRecovery_dis3pct,
         OB_LocalShare_BeachNourish_adj_dis3pct)

#pivoting tables longer
MuniCostOB_dis3pct_NPV <- MuniCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(MuniCostOB), value = TRUE)

# Apply the discount function
MuniCostOB_dis6pct <- dis6pct(MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
MuniCostOB_dis6pct <- MuniCostOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (6060USD)
MuniCostOB_dis6pct_NPV <- MuniCostOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
MuniCostOB_dis6pct_NPV<- MuniCostOB_dis6pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_dis6pct= 
           OB_LocalShare_HMGP_adj_dis6pct +
           OB_LocalShare_PA_Debris_adj_dis6pct +
           OB_LocalShare_PA_Protect_adj_dis6pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis6pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis6pct +
           OB_LocalShare_PA_Recreational_adj_dis6pct)
  
#selecting variables
MuniCostOB_dis6pct_NPV <- MuniCostOB_dis6pct_NPV %>%
  select(OB_Public_Safety_adj_dis6pct,
         OB_Public_Works_adj_dis6pct,
         OB_SchoolExpend_adj_dis6pct,
         OB_LocalShare_ResponseRecovery_dis6pct,
         OB_LocalShare_BeachNourish_adj_dis6pct)

#pivoting tables longer
MuniCostOB_dis6pct_NPV <- MuniCostOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 
```


#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(MuniCostOB), value = TRUE)

# Apply the discount function
MuniCostOB_dis7pct <- dis7pct(MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
MuniCostOB_dis7pct <- MuniCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
MuniCostOB_dis7pct_NPV <- MuniCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
MuniCostOB_dis7pct_NPV<- MuniCostOB_dis7pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_dis7pct= 
           OB_LocalShare_HMGP_adj_dis7pct +
           OB_LocalShare_PA_Debris_adj_dis7pct +
           OB_LocalShare_PA_Protect_adj_dis7pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis7pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis7pct +
           OB_LocalShare_PA_Recreational_adj_dis7pct)
  
#selecting variables
MuniCostOB_dis7pct_NPV <- MuniCostOB_dis7pct_NPV %>%
  select(OB_Public_Safety_adj_dis7pct,
         OB_Public_Works_adj_dis7pct,
         OB_SchoolExpend_adj_dis7pct,
         OB_LocalShare_ResponseRecovery_dis7pct,
         OB_LocalShare_BeachNourish_adj_dis7pct)

#pivoting tables longer
MuniCostOB_dis7pct_NPV <- MuniCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S1_MuniCostOB_NPV <- MuniCostOB_dis2pct_NPV %>%
  left_join(MuniCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(MuniCostOB_dis6pct_NPV, by = "variable") %>%
  left_join(MuniCostOB_dis7pct_NPV, by = "variable")

summary(S1_MuniCostOB_NPV )
```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S1_OB_Inputs'

write.csv(S1_MuniCostOB_NPV, file.path(path, "S1_MuniCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S1_MuniCostOB_dis2pct_NPV <- sum(S1_MuniCostOB_NPV$dis2pct)
S1_MuniCostOB_dis3pct_NPV <- sum(S1_MuniCostOB_NPV$dis3pct)
S1_MuniCostOB_dis6pct_NPV <- sum(S1_MuniCostOB_NPV$dis6pct)
S1_MuniCostOB_dis7pct_NPV <- sum(S1_MuniCostOB_NPV$dis7pct)

S1_MuniCostOB_dis2pct_NPV
S1_MuniCostOB_dis3pct_NPV
S1_MuniCostOB_dis6pct_NPV
S1_MuniCostOB_dis7pct_NPV
```