---
title: "S2_MuniCostFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for costs to Toms River for Scenario 2. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.

Assumptions:
-Public works for OB line drops by 100% , proportional to local road network removed
-Public safety line for OB drops by 100% , proportional to people relocated
-School district line for OB drops by 100% , proportional to school children relocated
-Municipal cost share for federal response and recovery (essential only for 2012 and 2013):
    -From PA, retaining only debris removal and protective measures for 2012 and 2013
    -NA on HMPG for Ortley Beach
-Municipal cost share for beach nourishment drops by 100%

-Municipal cost share for buyouts = 25% of market price for all homes


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToTR_fromOB/MuniCostOB.csv')
MuniCostOB <- MuniCostOB[,c(-1)]

#for buyout cost estimates
OB_2011_SP <- read.csv('./data/BaselineData/OB_MOD4_Produced/TimeSeries/S1/OB_2011_SalesPrice.csv')
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')


```


# Retaining only essential response and recovery costs
###Limited local response and recovery (essentials only). Retaining only:
    - PA Debris Removal (2012, 2013)
    - PA Protective measures (2012, 2013)
```{r include = FALSE}
S2_MuniCostOB <- MuniCostOB

S2_MuniCostOB$OB_LocalShare_PA_Debris_adj[S2_MuniCostOB$Year > 2013] <- NA
S2_MuniCostOB$OB_LocalShare_PA_Protect_adj[S2_MuniCostOB$Year > 2013] <- NA

#removing other columns
S2_MuniCostOB <- S2_MuniCostOB %>%
  select(Year,
         OB_Public_Safety_adj,
         OB_Public_Works_adj,
         OB_SchoolExpend_adj,
         OB_LocalShare_PA_Debris_adj,
         OB_LocalShare_PA_Protect_adj)


```

# Calculating municipal cost share for Buyouts (25 % of market price)
```{r include= FALSE}
#subset estimated sales price to the winzorized estimates
OB_2011_SP <- OB_2011_SP[,c("SalesPrice_LG_W")]

#obtain total sales price value in Ortley Beach
OB_2011_SP_Tot <- sum(OB_2011_SP)


#adjusting by CPI to 2020 USD
y11_CPI_multiplier <- Adjust$CPI_Multiplier_USD2020[Adjust$Year==2011]

OB_2011_SP_Tot_adj <- OB_2011_SP_Tot * y11_CPI_multiplier

#obtaining local share of buyout cost (25% cost share)
OB_LocalShare_BuyoutCost_adj <- OB_2011_SP_Tot_adj * 0.25

#Adding buyout costs to data frame 
S2_MuniCostOB$OB_LocalShare_BuyoutCost_adj[S2_MuniCostOB$Year==2012] <- OB_LocalShare_BuyoutCost_adj
```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S2_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniCostOB), value = TRUE)

# Apply the discount function
S2_MuniCostOB_dis2pct <- dis2pct(S2_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniCostOB_dis2pct <- S2_MuniCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniCostOB_dis2pct_NPV <- S2_MuniCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S2_MuniCostOB_dis2pct_NPV<- S2_MuniCostOB_dis2pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis2pct= 
           OB_LocalShare_PA_Debris_adj_dis2pct +
           OB_LocalShare_PA_Protect_adj_dis2pct)
  
#selecting variables
S2_MuniCostOB_dis2pct_NPV <- S2_MuniCostOB_dis2pct_NPV %>%
  select(OB_LocalShare_ResponseRecovery_adj_dis2pct,
         OB_LocalShare_BuyoutCost_adj_dis2pct)

#pivoting tables longer
S2_MuniCostOB_dis2pct_NPV <- S2_MuniCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S2_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniCostOB), value = TRUE)

# Apply the discount function
S2_MuniCostOB_dis3pct <- dis3pct(S2_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniCostOB_dis3pct <- S2_MuniCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniCostOB_dis3pct_NPV <- S2_MuniCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S2_MuniCostOB_dis3pct_NPV<- S2_MuniCostOB_dis3pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis3pct= 
           OB_LocalShare_PA_Debris_adj_dis3pct +
           OB_LocalShare_PA_Protect_adj_dis3pct)
  
#selecting variables
S2_MuniCostOB_dis3pct_NPV <- S2_MuniCostOB_dis3pct_NPV %>%
  select(OB_LocalShare_ResponseRecovery_adj_dis3pct,
         OB_LocalShare_BuyoutCost_adj_dis3pct)

#pivoting tables longer
S2_MuniCostOB_dis3pct_NPV <- S2_MuniCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```


#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
S2_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniCostOB), value = TRUE)

# Apply the discount function
S2_MuniCostOB_dis6pct <- dis6pct(S2_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniCostOB_dis6pct <- S2_MuniCostOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniCostOB_dis6pct_NPV <- S2_MuniCostOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S2_MuniCostOB_dis6pct_NPV<- S2_MuniCostOB_dis6pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis6pct= 
           OB_LocalShare_PA_Debris_adj_dis6pct +
           OB_LocalShare_PA_Protect_adj_dis6pct)
  
#selecting variables
S2_MuniCostOB_dis6pct_NPV <- S2_MuniCostOB_dis6pct_NPV %>%
  select(OB_LocalShare_ResponseRecovery_adj_dis6pct,
         OB_LocalShare_BuyoutCost_adj_dis6pct)

#pivoting tables longer
S2_MuniCostOB_dis6pct_NPV <- S2_MuniCostOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 
```


#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S2_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S2_MuniCostOB), value = TRUE)

# Apply the discount function
S2_MuniCostOB_dis7pct <- dis7pct(S2_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S2_MuniCostOB_dis7pct <- S2_MuniCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S2_MuniCostOB_dis7pct_NPV <- S2_MuniCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S2_MuniCostOB_dis7pct_NPV<- S2_MuniCostOB_dis7pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis7pct= 
           OB_LocalShare_PA_Debris_adj_dis7pct +
           OB_LocalShare_PA_Protect_adj_dis7pct)
  
#selecting variables
S2_MuniCostOB_dis7pct_NPV <- S2_MuniCostOB_dis7pct_NPV %>%
  select(OB_LocalShare_ResponseRecovery_adj_dis7pct,
         OB_LocalShare_BuyoutCost_adj_dis7pct)

#pivoting tables longer
S2_MuniCostOB_dis7pct_NPV <- S2_MuniCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S2_MuniCostOB_NPV <- S2_MuniCostOB_dis2pct_NPV %>%
  left_join(S2_MuniCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S2_MuniCostOB_dis6pct_NPV, by = "variable") %>%
  left_join(S2_MuniCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S2_OB_Inputs'

write.csv(S2_MuniCostOB_NPV, file.path(path, "S2_MuniCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S2_MuniCostOB_dis2pct_NPV <- sum(S2_MuniCostOB_NPV$dis2pct)
S2_MuniCostOB_dis3pct_NPV <- sum(S2_MuniCostOB_NPV$dis3pct)
S2_MuniCostOB_dis6pct_NPV <- sum(S2_MuniCostOB_NPV$dis6pct)
S2_MuniCostOB_dis7pct_NPV <- sum(S2_MuniCostOB_NPV$dis7pct)

S2_MuniCostOB_dis2pct_NPV
S2_MuniCostOB_dis3pct_NPV
S2_MuniCostOB_dis6pct_NPV
S2_MuniCostOB_dis7pct_NPV
```