---
title: "Municipal Revenues from Ortley Beach - Preprocessing"
author: "Laura Geronimo"
date: "2024-08-23"
output: html_document
---
The purpose of this worksheet is to estimate the costs to Toms River municipality of maintaining properties on Ortley Beach for Scenario 1 (replicating the historical record). Costs are estimated for the 2012-2022 period. Specifically we estimate:
1) Municipal public works expenditures for Ortley Beach
2) Municipal public safety expenditures for Ortley Beach
3) School district costs for Ortley Beach
4) Municipal cost share for Federal Response and Recovery for Ortley Beach
5) Municipal cost share for beach replenishment for Ortley Beach


Note that no municipal cost share was identified for the following:
- DCA / HUD (CDBG-DR) grants
- DOT grants

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)

##Importing Data ####
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')
TR_MuniExpend <- read.csv('./data/BaselineData/TomsRiver_Budget_ExpendituresSelection.csv')
TR_SchoolExpend <- read.csv('./data/BaselineData/TomsRiver_SchoolDistrictInfo.csv')

#Data for downscaling
OB_DemoData <- read_xlsx('./data/BaselineData/OrtleyBeach_DemographicData.xlsx', sheet=1)
TR_DemoData <- read_xlsx('./data/BaselineData/TomsRiver_DemographicData.xlsx', sheet=1)

TR_Sandy <- read_xlsx('./data/BaselineData/TomsRiver_SandyImpacts.xlsx', sheet=1)

#Municipal cost share for recovery to TR (needs further downscaling to OB)
TR_LocalShare_PA <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/PA_TR_LocalShare_ByYear.csv')
TR_LocalShare_HMGP <- read.csv('./data/BaselineData/NJ_SandyTransparencey_produced/HMGP_TR_LocalShare_ByYear.csv')


#Municipal cost share for recovery to OB (already downscaled to Ortley Beach (OB))
OB_BeachNourish <- read.csv('./data/BaselineData/OrtleyBeach_BeachNourishment.csv')


```

#1: Estimating Municipal Public Works expenditures for Ortley Beach 
```{r include=FALSE}
#selecting multipliers
Adjust <- Adjust %>% select(
  "Year",
  "CPI_Multiplier_USD2020",
  "SLIPD_V3_Multiplier"
)

#reducing to year 2012+
TR_MuniExpend <- subset(TR_MuniExpend, Year > 2011)


#joining inflation adjusters by year
TR_MuniExpend <- left_join(TR_MuniExpend, Adjust, by="Year")

#adjusting public works line using SLIPD
TR_MuniExpend$Public_Works_adj <- TR_MuniExpend$Public_Works * TR_MuniExpend$SLIPD_V3_Multiplier

###Downscaling Public Works (downscaled by total local road miles) ####
TR_LocalRoads_TotalMiles <-  TR_Sandy$Estimate[TR_Sandy$Variable == "TR_LocalRoads_TotalMiles"]
OB_LocalRoads_TotalMiles <-  TR_Sandy$Estimate[TR_Sandy$Variable == "OB_LocalRoads_TotalMiles"]

###Estimate Public Works funds disbursed per capita #### 
TR_MuniExpend$Public_Works_Per_Mile <- TR_MuniExpend$Public_Works_adj / TR_LocalRoads_TotalMiles

###Estimate Public Works funds disbursed to Ortley Beach #### 
TR_MuniExpend$OB_Public_Works_adj <- TR_MuniExpend$Public_Works_Per_Mile * OB_LocalRoads_TotalMiles 

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_Public_Works_adj <- TR_MuniExpend %>% select(
  "Year",
  "OB_Public_Works_adj")

```


#2: Estimating Municipal Public Safety expenditures for Ortley Beach 
```{r include=FALSE}


#adjusting Public Safety line using SLIPD
TR_MuniExpend$Public_Safety_adj <- TR_MuniExpend$Public_Safety * TR_MuniExpend$SLIPD_V3_Multiplier

##Downscaling Public Safety (downscaled by total population) ####
TR_tot_pop_estimate <-  TR_DemoData$TR_TotPopE[TR_DemoData$Year== "2012"]
OB_tot_pop_estimate <-  OB_DemoData$OB_TotPop[OB_DemoData$Year== "2012"]

###Estimate Public Safety funds disbursed per capita #### 
TR_MuniExpend$Public_Safety_Per_Capita <- TR_MuniExpend$Public_Safety_adj / TR_tot_pop_estimate

###Estimate Public Safety funds disbursed to Ortley Beach #### 
TR_MuniExpend$OB_Public_Safety_adj <- TR_MuniExpend$Public_Safety_Per_Capita * OB_tot_pop_estimate 

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_Public_Safety_adj <- TR_MuniExpend %>% select(
  "Year",
  "OB_Public_Safety_adj")

```



#3: Estimating School district costs for Ortley Beach
```{r include=FALSE}
#selecting multipliers
Adjust <- Adjust %>% select(
  "Year",
  "CPI_Multiplier_USD2020",
  "SLIPD_V3_Multiplier"
)

#reducing to year 2012+
TR_SchoolExpend <- subset(TR_SchoolExpend, Year > 2011)


#joining inflation adjusters by year
TR_SchoolExpend <- left_join(TR_SchoolExpend, Adjust, by="Year")


###Downscaling School Expenditure line (using adjusted cost per pupil * total school age children in Ortley Beach) ####
OB_TotSchoolAgeChildren_5_17_E <-  OB_DemoData %>% select(Year, OB_TotSchoolAgeChildren_5_17_E)

TR_SchoolExpend <- left_join(TR_SchoolExpend, OB_TotSchoolAgeChildren_5_17_E, by="Year")


###Estimate School Expend disbursed to Ortley Beach #### 
TR_SchoolExpend$OB_SchoolExpend_adj <- TR_SchoolExpend$TR_CostPerPupilBestEst_SLIPDAdj * TR_SchoolExpend$OB_TotSchoolAgeChildren_5_17_E

###Separate downscaled OB estimates for LMI projects as standalone table####
OB_SchoolExpend_adj <- TR_SchoolExpend %>% select(
  "Year",
  "OB_SchoolExpend_adj")

```


#4) Municipal cost share for Federal Response and Recovery for Ortley Beach


##4a) Downscaling HMGP data for to Ortley Beach by project
```{r include=FALSE}
#note: Based on project information, HMGP funds went to public buildings outside of Ortley Beach
head(TR_LocalShare_HMGP)

###Downscaling Generators by public buildings####
TR_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_PublicBuildings_SandyImpact"]
OB_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_PublicBuildings_SandyImpact"]

###Estimate HMGP disbursed per Sandy impacted public buildings #### 
TR_LocalShare_HMGP$HMGP_Disbursed_Per_PublicBuildings <- TR_LocalShare_HMGP$Local.Disbursed.Amount.Adj / 
TR_PublicBuildings 

###Estimate HMGP disbursed funds disbursed to Ortley Beach #### 
TR_LocalShare_HMGP$Local.Disbursed.Amount.Adj.OB <- TR_LocalShare_HMGP$HMGP_Disbursed_Per_PublicBuildings * OB_PublicBuildings

###Cutting line for Ocean County Health Dept Generator since located outside of OB ####
TR_LocalShare_HMGP <- subset(TR_LocalShare_HMGP, Project !="Ocean County Health Department Generator")

###Separate downscaled OB estimates for HMGP projects as standalone table####
OB_LocalShare_HMGP<- TR_LocalShare_HMGP %>% select(
  "Disbursed.Year",
  "Local.Disbursed.Amount.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_HMGP) <- c("Year", "OB_LocalShare_HMGP_adj")

```

##4c) Downscaling PA data for to Ortley Beach by project

```{r include=FALSE}


##PA Debris Removal by housing units impacted by Sandy####
###Subset PA to Debris Removal####
TR_LocalShare_PA_Debris <- subset(TR_LocalShare_PA, Project=="A - Debris Removal")

###Downscaling PA Debris Removal (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Debris Removal disbursed per housing unit #### 
TR_LocalShare_PA_Debris$PA_Debris_Disbursed_Per_Capita <- TR_LocalShare_PA_Debris$Local.Share.Adj / TR_HU_SandyImpact 

###Estimate PA Debris Removal funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_Debris$Local.Share.Adj.OB <- TR_LocalShare_PA_Debris$PA_Debris_Disbursed_Per_Capita * OB_HU_SandyImpact  

###Separate downscaled OB estimates for Debris Removal projects as standalone table####
OB_LocalShare_PA_Debris <- TR_LocalShare_PA_Debris %>% select(
  "Disbursed.Year",
  "Local.Share.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_PA_Debris) <- c("Year", "OB_LocalShare_PA_Debris_adj")


##PA Protective Measures by housing units impacted by Sandy####
###Subset PA to Protective Measures####
TR_LocalShare_PA_Protect <- subset(TR_LocalShare_PA, Project=="B - Protective Measures")

###Downscaling PA Protective Measures (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Protective Measures disbursed per housing unit #### 
TR_LocalShare_PA_Protect$PA_Protect_LocalShare_Per_HU <- TR_LocalShare_PA_Protect$Local.Share.Adj / TR_HU_SandyImpact 

###Estimate PA Protective Measures funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_Protect$Local.Share.Adj.OB <- TR_LocalShare_PA_Protect$PA_Protect_LocalShare_Per_HU * OB_HU_SandyImpact  

###Separate downscaled OB estimates for PA Protective Measures as standalone table####
OB_LocalShare_PA_Protect <- TR_LocalShare_PA_Protect %>% select(
  "Disbursed.Year",
  "Local.Share.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_PA_Protect) <- c("Year", "OB_LocalShare_PA_Protect_adj")


##PA  Roads & Bridges downscaled by miles of local roadway impacted by Sandy####
###Subset PA to Roads and Bridges####
TR_LocalShare_PA_RoadsBridges <- subset(TR_LocalShare_PA, Project=="C - Roads & Bridges")

###Downscaling PA Roads and Bridges (downscaled by miles of local roadway impacted by Sandy) ####
TR_Roads_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_LocalRoads_TotalMiles_SandyImpact"]
OB_Roads_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_LocalRoads_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_LocalShare_PA_RoadsBridges$PA_Roads_LocalShare_Per_RoadMiles <- TR_LocalShare_PA_RoadsBridges$Local.Share.Adj / TR_Roads_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_RoadsBridges$LocalShare.Adj.OB <- TR_LocalShare_PA_RoadsBridges$PA_Roads_LocalShare_Per_RoadMiles * OB_Roads_SandyImpact  

###Separate downscaled OB estimates for PA Roads and Bridges as standalone table####
OB_LocalShare_PA_RoadsBridges <- TR_LocalShare_PA_RoadsBridges  %>% select(
  "Disbursed.Year",
  "LocalShare.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_PA_RoadsBridges) <- c("Year", "OB_LocalShare_PA_RoadsBridges_adj")

##PA  Public Buildings downscaled by public buildings impacted by Sandy####
###Subset PA to Public Buildings####
TR_LocalShare_PA_PublicBuildings<- subset(TR_LocalShare_PA, Project=="E - Public Buildings")

###Downscaling PA Public Buildings (downscaled by public buildings impacted by Sandy ####
TR_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_PublicBuildings_SandyImpact"]
OB_PublicBuildings <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_PublicBuildings_SandyImpact"]

###Estimate PA Public Buildings disbursed per road miles impacted#### 
TR_LocalShare_PA_PublicBuildings$PA_PublicBuildings_Disbursed_Per_PublicBuildings <- TR_LocalShare_PA_PublicBuildings$Local.Share.Adj / TR_PublicBuildings

###Estimate PA Public Buildings funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_PublicBuildings$Local.Share.Adj.OB <- TR_LocalShare_PA_PublicBuildings$PA_PublicBuildings_Disbursed_Per_PublicBuildings * OB_PublicBuildings 

###Separate downscaled OB estimates for Public Buildings projects as standalone table####
OB_LocalShare_PA_PublicBuildings<- TR_LocalShare_PA_PublicBuildings %>% select(
  "Disbursed.Year",
  "Local.Share.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_PA_PublicBuildings) <- c("Year", "OB_LocalShare_PA_PublicBuildings_adj")


##PA  Public Utilities downscaled by housing units impacted by Sandy####
###Subset PA to Public Utilities####
TR_LocalShare_PA_PublicUtilities<- subset(TR_LocalShare_PA, Project=="F - Public Utilities")

###Downscaling PA Public Utilities (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_LocalShare_PA_PublicUtilities$PA_PublicUtilities_Disbursed_Per_PublicUtilities <- TR_LocalShare_PA_PublicUtilities$Local.Share.Adj / TR_HU_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_PublicUtilities$Local.Share.Adj.OB <- TR_LocalShare_PA_PublicUtilities$PA_PublicUtilities_Disbursed_Per_PublicUtilities * OB_HU_SandyImpact 

###Separate downscaled OB estimates for Public Utilities projects as standalone table####
OB_LocalShare_PA_PublicUtilities<- TR_LocalShare_PA_PublicUtilities %>% select(
  "Disbursed.Year",
  "Local.Share.Adj.OB")

###Renaming columns####
colnames(OB_LocalShare_PA_PublicUtilities) <- c("Year", "OB_LocalShare_PA_PublicUtilities_adj")



##PA  Recreational or Other downscaled by housing units impacted by Sandy####
###Subset PA to Recreational or Other####
TR_LocalShare_PA_Recreational<- subset(TR_LocalShare_PA, Project=="G - Recreational or Other")

###Downscaling PA Recreational or Other (downscaled by housing units impacted by Sandy) ####
TR_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="TR_HU_SandyImpact"]
OB_HU_SandyImpact <- TR_Sandy$Estimate[TR_Sandy$Variable=="OB_HU_SandyImpact"]

###Estimate PA Roads and Bridges disbursed per road miles impacted#### 
TR_LocalShare_PA_Recreational$PA_Recreational_Disbursed_Per_HU <- TR_LocalShare_PA_Recreational$Local.Share.Adj / TR_HU_SandyImpact 

###Estimate PA Roads and Bridges funds disbursed to Ortley Beach #### 
TR_LocalShare_PA_Recreational$Local.Share.Adj.OB <- TR_LocalShare_PA_Recreational$PA_Recreational_Disbursed_Per_HU * OB_HU_SandyImpact 

###Separate downscaled OB estimates for Recreational or Other projects as standalone table####
OB_LocalShare_PA_Recreational<- TR_LocalShare_PA_Recreational %>% select(
  "Disbursed.Year",
  "Local.Share.Adj.OB")

OB_LocalShare_PA_Recreational <- subset(OB_LocalShare_PA_Recreational, Disbursed.Year < 2023)

###Renaming columns####
colnames(OB_LocalShare_PA_Recreational) <- c("Year", "OB_LocalShare_PA_Recreational_adj")

```


#5) Estimating Local cost share for Beach and dune nourishment.
```{r include=FALSE}
#Note: this has already downscaled to Ortley beach via literature review. Reducing to local contribution

OB_LocalShare_BeachNourish <- OB_BeachNourish %>% select(
  "Year",
  "LocalCost_Adj"
)

###Renaming columns####
colnames(OB_LocalShare_BeachNourish) <- c("Year", "OB_LocalShare_BeachNourish_adj")

```


#Knitting tables together to obtain Toms River Municipal Revenues from Ortley Beach by Year
```{r include=FALSE}
library(purrr)
list_of_dfs <- list(OB_Public_Safety_adj,
                    OB_Public_Works_adj,
                    OB_SchoolExpend_adj,
                    OB_LocalShare_HMGP,
                    OB_LocalShare_PA_Debris,
                    OB_LocalShare_PA_Protect,
                    OB_LocalShare_PA_RoadsBridges,
                    OB_LocalShare_PA_PublicBuildings,
                    OB_LocalShare_PA_PublicUtilities,
                    OB_LocalShare_PA_Recreational,
                    OB_LocalShare_BeachNourish)

MuniCostOB <- reduce(list_of_dfs, function(x, y) left_join(x, y, by = "Year"))

#dropping 2011
MuniCostOB <- MuniCostOB %>% 
  filter(Year !=2011)

```

#Exporting S1 OB input tables:
```{r include=FALSE}
path <- 'C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/BaselineData/S1_OB_Inputs/Costs_ToTR_fromOB'

write.csv(MuniCostOB, file.path(path, "MuniCostOB.csv"), row.names = TRUE)


```
