---
title: "S3_MuniCostFomOB_Table"
author: "Laura Geronimo"
date: "2024-08-29"
output: html_document
---
The purpose of this worksheet is to estimate the NPV for costs to Toms River for Scenario 2. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, 6%, and 7% based on justifications provided in the SI.

Assumptions:
-Public works for OB line drops by 23% after 2012, proportional to local road network removed
-Public safety line for OB drops by 23% after 2012, proportional to housing units relocated (proxy for people relocated)
-School district line for OB drops by 23% after 2012, proportional to housing units relocated (proxy for school children relocated)
-Municipal cost share for federal response and recovery (assume sticky expenditure)
-Municipal cost share for beach nourishment (assume sticky expenditure)

-Municipal cost share for buyouts = 25% of market price for 421 bought out homes


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
MuniCostOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Costs_ToTR_fromOB/MuniCostOB.csv')
MuniCostOB <- MuniCostOB[,c(-1)]

OB_S3_2011_SP <- read.csv('./data/BaselineData/OB_MOD4_Produced/TimeSeries/S3/OB_S3_2011_SalesPrice.csv')
Adjust <- read.csv('./data/BaselineData/InflationAdjusters.csv')


```


# Reducing public works, public safety, and school expenditures post 2012 by 23% (proportional to properties bought out and road network removed). Retaining 77%
```{r include = FALSE}

S3_MuniCostOB <- MuniCostOB
S3_MuniCostOB$OB_Public_Works_adj[S3_MuniCostOB$Year > 2012] <- 
  S3_MuniCostOB$OB_Public_Works_adj[S3_MuniCostOB$Year > 2012] * 0.77
S3_MuniCostOB$OB_Public_Safety_adj[S3_MuniCostOB$Year > 2012] <- 
  S3_MuniCostOB$OB_Public_Safety_adj[S3_MuniCostOB$Year > 2012] * 0.77
S3_MuniCostOB$OB_SchoolExpend_adj[S3_MuniCostOB$Year > 2012] <- 
  S3_MuniCostOB$OB_SchoolExpend_adj[S3_MuniCostOB$Year > 2012] * 0.77

```

# Calculating municipal cost share for Buyouts (25 % of market price for 421 homes)
```{r include= FALSE}
#subset estimated sales price to the winzorized estimates
OB_S3_2011_SP <- OB_S3_2011_SP[,c("SalesPrice_LG_W")]

#obtain total sales price value in Ortley Beach
OB_S3_2011_SP_Tot <- sum(OB_S3_2011_SP)


#adjusting by CPI to 2020 USD
y11_CPI_multiplier <- Adjust$CPI_Multiplier_USD2020[Adjust$Year==2011]

OB_S3_2011_SP_Tot_adj <- OB_S3_2011_SP_Tot * y11_CPI_multiplier

#obtaining local share of buyout cost (25% cost share)
OB_LocalShare_BuyoutCost_adj <- OB_S3_2011_SP_Tot_adj * 0.25

#Adding buyout costs to data frame 
S3_MuniCostOB$OB_LocalShare_BuyoutCost_adj[S3_MuniCostOB$Year==2012] <- OB_LocalShare_BuyoutCost_adj
```


#Discounting values to 2%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S3_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S3_MuniCostOB), value = TRUE)

# Apply the discount function
S3_MuniCostOB_dis2pct <- dis2pct(S3_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S3_MuniCostOB_dis2pct <- S3_MuniCostOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S3_MuniCostOB_dis2pct_NPV <- S3_MuniCostOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S3_MuniCostOB_dis2pct_NPV<- S3_MuniCostOB_dis2pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis2pct= 
           OB_LocalShare_HMGP_adj_dis2pct +
           OB_LocalShare_PA_Debris_adj_dis2pct +
           OB_LocalShare_PA_Protect_adj_dis2pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis2pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis2pct +
           OB_LocalShare_PA_Recreational_adj_dis2pct)
  
#selecting variables
S3_MuniCostOB_dis2pct_NPV <- S3_MuniCostOB_dis2pct_NPV %>%
  select(OB_Public_Safety_adj_dis2pct,
         OB_Public_Works_adj_dis2pct,
         OB_SchoolExpend_adj_dis2pct,
         OB_LocalShare_ResponseRecovery_adj_dis2pct,
         OB_LocalShare_BeachNourish_adj_dis2pct,
         OB_LocalShare_BuyoutCost_adj_dis2pct)

#pivoting tables longer
S3_MuniCostOB_dis2pct_NPV <- S3_MuniCostOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```


#Discounting values to 3%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S3_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S3_MuniCostOB), value = TRUE)

# Apply the discount function
S3_MuniCostOB_dis3pct <- dis3pct(S3_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S3_MuniCostOB_dis3pct <- S3_MuniCostOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (3030USD)
S3_MuniCostOB_dis3pct_NPV <- S3_MuniCostOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S3_MuniCostOB_dis3pct_NPV<- S3_MuniCostOB_dis3pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis3pct= 
           OB_LocalShare_HMGP_adj_dis3pct +
           OB_LocalShare_PA_Debris_adj_dis3pct +
           OB_LocalShare_PA_Protect_adj_dis3pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis3pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis3pct +
           OB_LocalShare_PA_Recreational_adj_dis3pct)
  
#selecting variables
S3_MuniCostOB_dis3pct_NPV <- S3_MuniCostOB_dis3pct_NPV %>%
  select(OB_Public_Safety_adj_dis3pct,
         OB_Public_Works_adj_dis3pct,
         OB_SchoolExpend_adj_dis3pct,
         OB_LocalShare_ResponseRecovery_adj_dis3pct,
         OB_LocalShare_BeachNourish_adj_dis3pct,
         OB_LocalShare_BuyoutCost_adj_dis3pct)

#pivoting tables longer
S3_MuniCostOB_dis3pct_NPV <- S3_MuniCostOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```



#Discounting values to 6%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 6% ####
#Creating a year index 
S3_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 6% 
dis6pct <- function(df, amount_cols, year_index_col, discount_rate = 0.06, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis6pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S3_MuniCostOB), value = TRUE)

# Apply the discount function
S3_MuniCostOB_dis6pct <- dis6pct(S3_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S3_MuniCostOB_dis6pct <- S3_MuniCostOB_dis6pct %>%
  select(Year, YearIndex, ends_with("_dis6pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (6060USD)
S3_MuniCostOB_dis6pct_NPV <- S3_MuniCostOB_dis6pct %>%
  summarise(across(ends_with("_dis6pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S3_MuniCostOB_dis6pct_NPV<- S3_MuniCostOB_dis6pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis6pct= 
           OB_LocalShare_HMGP_adj_dis6pct +
           OB_LocalShare_PA_Debris_adj_dis6pct +
           OB_LocalShare_PA_Protect_adj_dis6pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis6pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis6pct +
           OB_LocalShare_PA_Recreational_adj_dis6pct)
  
#selecting variables
S3_MuniCostOB_dis6pct_NPV <- S3_MuniCostOB_dis6pct_NPV %>%
  select(OB_Public_Safety_adj_dis6pct,
         OB_Public_Works_adj_dis6pct,
         OB_SchoolExpend_adj_dis6pct,
         OB_LocalShare_ResponseRecovery_adj_dis6pct,
         OB_LocalShare_BeachNourish_adj_dis6pct,
         OB_LocalShare_BuyoutCost_adj_dis6pct)

#pivoting tables longer
S3_MuniCostOB_dis6pct_NPV <- S3_MuniCostOB_dis6pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis6pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis6pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV, and grouping in buckets
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S3_MuniCostOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S3_MuniCostOB), value = TRUE)

# Apply the discount function
S3_MuniCostOB_dis7pct <- dis7pct(S3_MuniCostOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S3_MuniCostOB_dis7pct <- S3_MuniCostOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
S3_MuniCostOB_dis7pct_NPV <- S3_MuniCostOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

###creating buckets ####
#Summing Muni Share for Response and Recovery
S3_MuniCostOB_dis7pct_NPV<- S3_MuniCostOB_dis7pct_NPV %>%
  mutate(OB_LocalShare_ResponseRecovery_adj_dis7pct= 
           OB_LocalShare_HMGP_adj_dis7pct +
           OB_LocalShare_PA_Debris_adj_dis7pct +
           OB_LocalShare_PA_Protect_adj_dis7pct +
           OB_LocalShare_PA_PublicBuildings_adj_dis7pct +
           OB_LocalShare_PA_PublicUtilities_adj_dis7pct +
           OB_LocalShare_PA_Recreational_adj_dis7pct)
  
#selecting variables
S3_MuniCostOB_dis7pct_NPV <- S3_MuniCostOB_dis7pct_NPV %>%
  select(OB_Public_Safety_adj_dis7pct,
         OB_Public_Works_adj_dis7pct,
         OB_SchoolExpend_adj_dis7pct,
         OB_LocalShare_ResponseRecovery_adj_dis7pct,
         OB_LocalShare_BeachNourish_adj_dis7pct,
         OB_LocalShare_BuyoutCost_adj_dis7pct)

#pivoting tables longer
S3_MuniCostOB_dis7pct_NPV <- S3_MuniCostOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S3_MuniCostOB_NPV <- S3_MuniCostOB_dis2pct_NPV %>%
  left_join(S3_MuniCostOB_dis3pct_NPV, by = "variable") %>%
  left_join(S3_MuniCostOB_dis6pct_NPV, by = "variable") %>%
  left_join(S3_MuniCostOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S3_OB_Inputs'

write.csv(S3_MuniCostOB_NPV, file.path(path, "S3_MuniCostOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S3_MuniCostOB_dis2pct_NPV <- sum(S3_MuniCostOB_NPV$dis2pct)
S3_MuniCostOB_dis3pct_NPV <- sum(S3_MuniCostOB_NPV$dis3pct)
S3_MuniCostOB_dis6pct_NPV <- sum(S3_MuniCostOB_NPV$dis6pct)
S3_MuniCostOB_dis7pct_NPV <- sum(S3_MuniCostOB_NPV$dis7pct)

S3_MuniCostOB_dis2pct_NPV
S3_MuniCostOB_dis3pct_NPV
S3_MuniCostOB_dis6pct_NPV
S3_MuniCostOB_dis7pct_NPV
```