---
title: "OB_AAL_Estimates"
author: "Laura Geronimo"
date: "2024-09-17"
output: html_document
---


The purpose of this worksheet is to estimate Average Annual Losses (AAL) for Scenarios with properties remaining on the island, based on First Street Data. First street provides data for 2021 and 2051. For each scenario, we take the total AAL for the properties in the scenario, and do linear interpolation to estimate the values for years 2012:2062


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include= FALSE}
#setup
getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

library(dplyr)
library(tidyr)

#import data ####
S3_OB <- read.csv('./data/BaselineData/OB_MOD4_Produced/FromGIS/ClusterInfo/S3_Cluster1_gispin.csv')
S4_OB <- read.csv('./data/BaselineData/OB_MOD4_Produced/FromGIS/ClusterInfo/S4_ClusterBay_gispin.csv')

S1_OB <- read.csv('data/BaselineData/FirstStreet/OB_FirstStreetData.csv') #licensed parcel level First street data, clipped to MOD-IV residential parcels in GIS
```

#Examining missing data
```{r include= FALSE}

table(is.na(S1_OB$aal_2021_h)) #1000 #
table(is.na(S1_OB$aal_2021_m)) #13 #
table(is.na(S1_OB$aal_2021_l)) #1000 

table(is.na(S1_OB$aal_2051_h)) #364
table(is.na(S1_OB$aal_2051_m)) #13
table(is.na(S1_OB$aal_2051_l)) #364

summary(S1_OB$aalUSD_2051_mid)

```

#Assigning properties with missing AVM the median AVM
```{r include=FALSE}
table(is.na(S1_OB$avm))
summary(S1_OB$avm) #401347 = median
S1_OB$avm[is.na(S1_OB$avm)] <- 401347

```

#Calculating Scenario 1 (S1) AALs
Calculating 2021 & 2051 values per FS guidance - personal communication and guidance####
First Street Guidance: https://help.firststreet.org/hc/en-us/articles/6016946455831-Average-Annual-Loss-AAL-Data
```{r include= FALSE}

##S1 Low Climate Scenario ####
S1_OB$aalUSD_2021_low <- ((S1_OB$aal_2021_l*S1_OB$avm)*0.70) # we multiply by 0.7 per First street to obtain the true average market value 
summary(S1_OB$aalUSD_2021_low) #median = 6457
range(S1_OB$aalUSD_2021_low, na.rm=T)
hist(S1_OB$aalUSD_2021_low)
Known2021AAL_Low <- sum(S1_OB$aalUSD_2021_low, na.rm=T) #7654218

names(S1_OB)
S1_OB$aalUSD_2051_low <- ((S1_OB$aal_2051_l*S1_OB$avm)*0.70)
range(S1_OB$aalUSD_2051_low)
hist(S1_OB$aalUSD_2051_low)
Known2051AAL_Low <- sum(S1_OB$aalUSD_2051_low, na.rm=T) #11447111


##S1 Mid Climate Scenario ####
S1_OB$aalUSD_2021_mid <- ((S1_OB$aal_2021_m*S1_OB$avm)*0.70)
range(S1_OB$aalUSD_2021_mid)
hist(S1_OB$aalUSD_2021_mid)
Known2021AAL_Mid <- sum(S1_OB$aalUSD_2021_mid, na.rm=T) #7906273


S1_OB$aalUSD_2051_mid <- ((S1_OB$aal_2051_m*S1_OB$avm)*0.70)
range(S1_OB$aalUSD_2051_mid)
hist(S1_OB$aalUSD_2051_mid)
Known2051AAL_Mid <- sum(S1_OB$aalUSD_2051_mid, na.rm=T) #13953576



##S1 High Climate Scenario ####

S1_OB$aalUSD_2021_high <- ((S1_OB$aal_2021_h*S1_OB$avm)*0.70)
range(S1_OB$aalUSD_2021_high)
hist(S1_OB$aalUSD_2021_high)
Known2021AAL_High <- sum(S1_OB$aalUSD_2021_high, na.rm=T) #8206059


S1_OB$aalUSD_2051_high <- ((S1_OB$aal_2051_h*S1_OB$avm)*0.70)
range(S1_OB$aalUSD_2051_high)
hist(S1_OB$aalUSD_2051_high)
Known2051AAL_High <- sum(S1_OB$aalUSD_2051_high, na.rm=T) #16665310



##S1 data.frame  AALs ####
S1_AAL <- data.frame(
  year = c(2021, 2051),
  KnownTotalAAL_Low = c(Known2021AAL_Low, Known2051AAL_Low),
  KnownTotalAAL_Mid = c(Known2021AAL_Mid, Known2051AAL_Mid),
  KnownTotalAAL_High = c(Known2021AAL_High, Known2051AAL_High)
)

#Reducing Columns####
names(S1_OB)

S1_OB <- S1_OB[,c("GIS_PIN","fsid","long","lat",
              "avm",
              "aalUSD_2051_low",
              "aalUSD_2021_low",
              "aalUSD_2051_mid",
              "aalUSD_2021_mid",
              "aalUSD_2051_high",
              "aalUSD_2021_high")]
```

#Scenario 3 (Atlantic Side Cluster) AALS
```{r include= FALSE}

#Scenario 3: Cluster 1####

#joining AAL data to subset of parcels in the S3 cluster
names(S3_OB)[names(S3_OB)=="gis_pin"] <- "GIS_PIN"
S3_OB <- left_join(S3_OB, S1_OB, by="GIS_PIN", copy=F)

#checking
table(is.na(S3_OB$aalUSD_2021_low)) #384
table(is.na(S3_OB$aalUSD_2051_low)) #200
table(is.na(S3_OB$aalUSD_2021_mid)) #3
table(is.na(S3_OB$aalUSD_2051_mid)) #3
table(is.na(S3_OB$aalUSD_2021_high)) #384
table(is.na(S3_OB$aalUSD_2051_high)) #200


sum(S3_OB$avm, na.rm=T) #185,524,157
sum(S3_OB$aalUSD_2021_mid, na.rm = T) #5161.866
sum(S3_OB$aalUSD_2051_mid, na.rm = T) #179224.2


##S3 Low Climate Scenario ####
Known2021AAL_Low <- sum(S3_OB$aalUSD_2021_low, na.rm=T) #2873.974
Known2051AAL_Low <- sum(S3_OB$aalUSD_2051_low, na.rm=T) #53829.2


##S3 Mid Climate Scenario ####
Known2021AAL_Mid <- sum(S3_OB$aalUSD_2021_mid, na.rm=T) #5161.866
Known2051AAL_Mid <- sum(S3_OB$aalUSD_2051_mid, na.rm=T) #179224.2



##S3 High Climate Scenario ####
Known2021AAL_High <- sum(S3_OB$aalUSD_2021_high, na.rm=T) #8304.223
Known2051AAL_High <- sum(S3_OB$aalUSD_2051_high, na.rm=T) #383838.9


## S3  data.frame  ####
S3_AAL <- data.frame(
  year = c(2021, 2051),
  KnownTotalAAL_Low = c(Known2021AAL_Low, Known2051AAL_Low),
  KnownTotalAAL_Mid = c(Known2021AAL_Mid, Known2051AAL_Mid),
  KnownTotalAAL_High = c(Known2021AAL_High, Known2051AAL_High)
)

```

#Scenario 4: Bayside cluster ####
```{r include=FALSE}

#Scenario 4: Cluster Bayside####

#joining AAL data to subset of parcels in the S4 cluster
names(S4_OB)[names(S4_OB)=="gis_pin"] <- "GIS_PIN"
S4_OB <- left_join(S4_OB, S1_OB, by="GIS_PIN", copy=F)

#checking
table(is.na(S4_OB$aalUSD_2021_low)) #384
table(is.na(S4_OB$aalUSD_2051_low)) #200
table(is.na(S4_OB$aalUSD_2021_mid)) #3
table(is.na(S4_OB$aalUSD_2051_mid)) #3
table(is.na(S4_OB$aalUSD_2021_high)) #384
table(is.na(S4_OB$aalUSD_2051_high)) #200


sum(S4_OB$avm, na.rm=T) #185,524,157
sum(S4_OB$aalUSD_2021_mid, na.rm = T) #5161.866
sum(S4_OB$aalUSD_2051_mid, na.rm = T) #179224.2


##S4 Low Climate Scenario ####
Known2021AAL_Low <- sum(S4_OB$aalUSD_2021_low, na.rm=T) #2873.974
Known2051AAL_Low <- sum(S4_OB$aalUSD_2051_low, na.rm=T) #53829.2


##S4 Mid Climate Scenario ####
Known2021AAL_Mid <- sum(S4_OB$aalUSD_2021_mid, na.rm=T) #5161.866
Known2051AAL_Mid <- sum(S4_OB$aalUSD_2051_mid, na.rm=T) #179224.2



##S4 High Climate Scenario ####
Known2021AAL_High <- sum(S4_OB$aalUSD_2021_high, na.rm=T) #8304.223
Known2051AAL_High <- sum(S4_OB$aalUSD_2051_high, na.rm=T) #383838.9



## S4 data.frame AALs  ####
S4_AAL <- data.frame(
  year = c(2021, 2051),
  KnownTotalAAL_Low = c(Known2021AAL_Low, Known2051AAL_Low),
  KnownTotalAAL_Mid = c(Known2021AAL_Mid, Known2051AAL_Mid),
  KnownTotalAAL_High = c(Known2021AAL_High, Known2051AAL_High)
)

```

#Writing out files
```{r include=FALSE}
path <- './data/BaselineData/FirstStreet'

write.csv(S1_AAL, file.path(path, "S1_AAL.csv"), row.names = FALSE)
write.csv(S3_AAL, file.path(path, "S3_AAL.csv"), row.names = FALSE)
write.csv(S4_AAL, file.path(path, "S4_AAL.csv"), row.names = FALSE)


```

