---
title: "OB_AAL_Estimates"
author: "Laura Geronimo"
date: "2024-09-17"
output: html_document
---


The purpose of this worksheet is to estimate AAL for Scenarios with properties remaining on the island, based on First Street Data. First street provides data for 2021 and 2051. For each scenario, we take the total AAL for the properties in the scenario, and do linear interpolation to estimate the values for years 2012:2062


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include= FALSE}
#setup
getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

library(dplyr)
library(tidyr)

#import data ####
S1_AAL <- read.csv('data/BaselineData/FirstStreet/S1_AAL.csv', check.names = FALSE) 
S3_AAL <- read.csv('data/BaselineData/FirstStreet/S3_AAL.csv') 
S4_AAL <- read.csv('data/BaselineData/FirstStreet/S4_AAL.csv') 

```


#Defining function to linearly interpolate AAL data for missing years based on known data
```{r include = FALSE}
interpolate_AAL <- function(df, start_year = 2012, end_year = 2062) {
  # Define the years for interpolation/extrapolation
  interpolation_years <- start_year:end_year
  
  # Get the slope between 2021 and 2051 for each AAL level
  slope_low <- (df$KnownTotalAAL_Low[2] - df$KnownTotalAAL_Low[1]) / (df$year[2] - df$year[1])
  slope_mid <- (df$KnownTotalAAL_Mid[2] - df$KnownTotalAAL_Mid[1]) / (df$year[2] - df$year[1])
  slope_high <- (df$KnownTotalAAL_High[2] - df$KnownTotalAAL_High[1]) / (df$year[2] - df$year[1])
  
  # Perform linear interpolation with extrapolation for each AAL level
  interpolated_low <- approx(df$year, df$KnownTotalAAL_Low, xout = interpolation_years, rule = 1)$y
  interpolated_mid <- approx(df$year, df$KnownTotalAAL_Mid, xout = interpolation_years, rule = 1)$y
  interpolated_high <- approx(df$year, df$KnownTotalAAL_High, xout = interpolation_years, rule = 1)$y
  
  # Extrapolate values before 2021 and after 2051
  for (i in 1:length(interpolation_years)) {
    if (interpolation_years[i] < 2021) {
      interpolated_low[i] <- df$KnownTotalAAL_Low[1] + slope_low * (interpolation_years[i] - 2021)
      interpolated_mid[i] <- df$KnownTotalAAL_Mid[1] + slope_mid * (interpolation_years[i] - 2021)
      interpolated_high[i] <- df$KnownTotalAAL_High[1] + slope_high * (interpolation_years[i] - 2021)
    }
    
    if (interpolation_years[i] > 2051) {
      interpolated_low[i] <- df$KnownTotalAAL_Low[2] + slope_low * (interpolation_years[i] - 2051)
      interpolated_mid[i] <- df$KnownTotalAAL_Mid[2] + slope_mid * (interpolation_years[i] - 2051)
      interpolated_high[i] <- df$KnownTotalAAL_High[2] + slope_high * (interpolation_years[i] - 2051)
    }
  }
  
  # Replace negative values with 0
  interpolated_low[interpolated_low < 0] <- 0
  interpolated_mid[interpolated_mid < 0] <- 0
  interpolated_high[interpolated_high < 0] <- 0
  
  # Create a new dataframe with interpolated/extrapolated values
  interpolated_df <- data.frame(
    year = interpolation_years,
    InterpolatedAAL_Low = interpolated_low,
    InterpolatedAAL_Mid = interpolated_mid,
    InterpolatedAAL_High = interpolated_high
  )
  
  # Return the interpolated dataframe
  return(interpolated_df)
}

```


### Interpolating for S1
```{r include=FALSE}
S1_AAL_interpolated <- interpolate_AAL(S1_AAL)


#QC
plot(S1_AAL_interpolated$year, S1_AAL_interpolated$InterpolatedAAL_Mid)
```

### Interpolating for S3
```{r include=FALSE}
S3_AAL_interpolated <- interpolate_AAL(S3_AAL)

#QC
plot(S3_AAL_interpolated$year, S3_AAL_interpolated$InterpolatedAAL_Mid)
```


### Interpolating for S4
```{r include=FALSE}
S4_AAL_interpolated <- interpolate_AAL(S4_AAL)

#QC
plot(S4_AAL_interpolated$year, S4_AAL_interpolated$InterpolatedAAL_Mid)
```


# Discounting 
## Defining function to discount values at 2 percent
```{r include = FALSE}
discount_AAL_2pct <- function(df, discount_rate = 0.02, base_year = 2012) {
  # Ensure discount factors are calculated for every year in the dataframe
  discount_factors <- 1 / ((1 + discount_rate) ^ (df$year - base_year))
  
  # Apply the discount factor to the interpolated AAL values
  df$AAL_Low_dis2pct <- df$InterpolatedAAL_Low * discount_factors
  df$AAL_Mid_dis2pct <- df$InterpolatedAAL_Mid * discount_factors
  df$AAL_High_dis2pct <- df$InterpolatedAAL_High * discount_factors
  
  # Return the dataframe with discounted values
  return(df)
}

```

### Discounting AALs at 2% by scenario
```{r include = FALSE}

S1_AAL_dis2pct <- discount_AAL_2pct(S1_AAL_interpolated)
S3_AAL_dis2pct <- discount_AAL_2pct(S3_AAL_interpolated)
S4_AAL_dis2pct <- discount_AAL_2pct(S4_AAL_interpolated)

```

## Defining function to discount values at 3 percent
```{r include = FALSE}
discount_AAL_3pct <- function(df, discount_rate = 0.03, base_year = 2012) {
  # Ensure discount factors are calculated for every year in the dataframe
  discount_factors <- 1 / ((1 + discount_rate) ^ (df$year - base_year))
  
  # Apply the discount factor to the interpolated AAL values
  df$AAL_Low_dis3pct <- df$InterpolatedAAL_Low * discount_factors
  df$AAL_Mid_dis3pct <- df$InterpolatedAAL_Mid * discount_factors
  df$AAL_High_dis3pct <- df$InterpolatedAAL_High * discount_factors
  
  # Return the dataframe with discounted values
  return(df)
}

```

### Discounting AALs at 3% by scenario
```{r include = FALSE}

S1_AAL_dis3pct <- discount_AAL_3pct(S1_AAL_interpolated)
S3_AAL_dis3pct <- discount_AAL_3pct(S3_AAL_interpolated)
S4_AAL_dis3pct <- discount_AAL_3pct(S4_AAL_interpolated)

```


## Defining function to discount values at 7 percent
```{r include = FALSE}
discount_AAL_7pct <- function(df, discount_rate = 0.07, base_year = 2012) {
  # Ensure discount factors are calculated for every year in the dataframe
  discount_factors <- 1 / ((1 + discount_rate) ^ (df$year - base_year))
  
  # Apply the discount factor to the interpolated AAL values
  df$AAL_Low_dis7pct <- df$InterpolatedAAL_Low * discount_factors
  df$AAL_Mid_dis7pct <- df$InterpolatedAAL_Mid * discount_factors
  df$AAL_High_dis7pct <- df$InterpolatedAAL_High * discount_factors
  
  # Return the dataframe with discounted values
  return(df)
}

```

### Discounting AALs at 7% by scenario
```{r include = FALSE}

S1_AAL_dis7pct <- discount_AAL_7pct(S1_AAL_interpolated)
S3_AAL_dis7pct <- discount_AAL_7pct(S3_AAL_interpolated)
S4_AAL_dis7pct <- discount_AAL_7pct(S4_AAL_interpolated)

```

#Obtaining NPV of residual risk (2023-2062) by buyout scenario, discount rate, and climate scenario

##S1 NPV 
```{r include = FALSE}
S1_NPV_ResRisk_Low_dis0pct <- sum(S1_AAL_interpolated$InterpolatedAAL_Low[S1_AAL_interpolated$year %in% 2023:2062])

S1_NPV_ResRisk_Low_dis2pct <- sum(S1_AAL_dis2pct$AAL_Low_dis2pct[S1_AAL_dis2pct$year %in% 2023:2062])
S1_NPV_ResRisk_Low_dis3pct <- sum(S1_AAL_dis3pct$AAL_Low_dis3pct[S1_AAL_dis3pct$year %in% 2023:2062])
S1_NPV_ResRisk_Low_dis7pct <- sum(S1_AAL_dis7pct$AAL_Low_dis7pct[S1_AAL_dis7pct$year %in% 2023:2062])

S1_NPV_ResRisk_Mid_dis2pct <- sum(S1_AAL_dis2pct$AAL_Mid_dis2pct[S1_AAL_dis2pct$year %in% 2023:2062])
S1_NPV_ResRisk_Mid_dis3pct <- sum(S1_AAL_dis3pct$AAL_Mid_dis3pct[S1_AAL_dis3pct$year %in% 2023:2062])
S1_NPV_ResRisk_Mid_dis7pct <- sum(S1_AAL_dis7pct$AAL_Mid_dis7pct[S1_AAL_dis7pct$year %in% 2023:2062])

S1_NPV_ResRisk_High_dis2pct <- sum(S1_AAL_dis2pct$AAL_High_dis2pct[S1_AAL_dis2pct$year %in% 2023:2062])
S1_NPV_ResRisk_High_dis3pct <- sum(S1_AAL_dis3pct$AAL_High_dis3pct[S1_AAL_dis3pct$year %in% 2023:2062])
S1_NPV_ResRisk_High_dis7pct <- sum(S1_AAL_dis7pct$AAL_High_dis7pct[S1_AAL_dis7pct$year %in% 2023:2062])


# Create a new dataframe with the desired structure
S1_NPV_ResRisk <- data.frame(
  Scenario = c("Low", "Medium", "High"),
  dis2pct = c(S1_NPV_ResRisk_Low_dis2pct, S1_NPV_ResRisk_Mid_dis2pct, S1_NPV_ResRisk_High_dis2pct),
  dis3pct = c(S1_NPV_ResRisk_Low_dis3pct, S1_NPV_ResRisk_Mid_dis3pct, S1_NPV_ResRisk_High_dis3pct),
  dis7pct = c(S1_NPV_ResRisk_Low_dis7pct, S1_NPV_ResRisk_Mid_dis7pct, S1_NPV_ResRisk_High_dis7pct)
)

# View the dataframe
print(S1_NPV_ResRisk)
```



##S3 NPV 
#Note that once we calculate the residual risk for the cluster, we subtract it from the S1 results to obtain the residual risk to remaining homes after S3 cluster is bought out
```{r include = FALSE}

#low
S3_NPV_ResRisk_Low_dis2pct <- S1_NPV_ResRisk_Low_dis2pct - sum(S3_AAL_dis2pct$AAL_Low_dis2pct[S3_AAL_dis2pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_Low_dis3pct <- S1_NPV_ResRisk_Low_dis3pct - sum(S3_AAL_dis3pct$AAL_Low_dis3pct[S3_AAL_dis3pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_Low_dis7pct <- S1_NPV_ResRisk_Low_dis7pct - sum(S3_AAL_dis7pct$AAL_Low_dis7pct[S3_AAL_dis7pct$year %in% 2023:2062]) 

#mid
S3_NPV_ResRisk_Mid_dis2pct <- S1_NPV_ResRisk_Mid_dis2pct - sum(S3_AAL_dis2pct$AAL_Mid_dis2pct[S3_AAL_dis2pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_Mid_dis3pct <- S1_NPV_ResRisk_Mid_dis3pct - sum(S3_AAL_dis3pct$AAL_Mid_dis3pct[S3_AAL_dis3pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_Mid_dis7pct <- S1_NPV_ResRisk_Mid_dis7pct - sum(S3_AAL_dis7pct$AAL_Mid_dis7pct[S3_AAL_dis7pct$year %in% 2023:2062]) 

#high
S3_NPV_ResRisk_High_dis2pct <- S1_NPV_ResRisk_High_dis2pct - sum(S3_AAL_dis2pct$AAL_High_dis2pct[S3_AAL_dis2pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_High_dis3pct <- S1_NPV_ResRisk_High_dis3pct - sum(S3_AAL_dis3pct$AAL_High_dis3pct[S3_AAL_dis3pct$year %in% 2023:2062]) 

S3_NPV_ResRisk_High_dis7pct <- S1_NPV_ResRisk_High_dis7pct - sum(S3_AAL_dis7pct$AAL_High_dis7pct[S3_AAL_dis7pct$year %in% 2023:2062]) 



# Create a new dataframe 
S3_NPV_ResRisk <- data.frame(
  Scenario = c("Low", "Medium", "High"),
  dis2pct = c(S3_NPV_ResRisk_Low_dis2pct, S3_NPV_ResRisk_Mid_dis2pct, S3_NPV_ResRisk_High_dis2pct),
  dis3pct = c(S3_NPV_ResRisk_Low_dis3pct, S3_NPV_ResRisk_Mid_dis3pct, S3_NPV_ResRisk_High_dis3pct),
  dis7pct = c(S3_NPV_ResRisk_Low_dis7pct, S3_NPV_ResRisk_Mid_dis7pct, S3_NPV_ResRisk_High_dis7pct)
)

# View the dataframe
print(S3_NPV_ResRisk)
```




##S4 NPV 
#Note that once we calculate the residual risk for the cluster, we subtract it from the S1 results to obtain the residual risk to remaining homes after S3 cluster is bought out
```{r include = FALSE}

#low
S4_NPV_ResRisk_Low_dis2pct <- S1_NPV_ResRisk_Low_dis2pct - sum(S4_AAL_dis2pct$AAL_Low_dis2pct[S4_AAL_dis2pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_Low_dis3pct <- S1_NPV_ResRisk_Low_dis3pct - sum(S4_AAL_dis3pct$AAL_Low_dis3pct[S4_AAL_dis3pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_Low_dis7pct <- S1_NPV_ResRisk_Low_dis7pct - sum(S4_AAL_dis7pct$AAL_Low_dis7pct[S4_AAL_dis7pct$year %in% 2023:2062]) 

#mid
S4_NPV_ResRisk_Mid_dis2pct <- S1_NPV_ResRisk_Mid_dis2pct - sum(S4_AAL_dis2pct$AAL_Mid_dis2pct[S4_AAL_dis2pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_Mid_dis3pct <- S1_NPV_ResRisk_Mid_dis3pct - sum(S4_AAL_dis3pct$AAL_Mid_dis3pct[S4_AAL_dis3pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_Mid_dis7pct <- S1_NPV_ResRisk_Mid_dis7pct - sum(S4_AAL_dis7pct$AAL_Mid_dis7pct[S4_AAL_dis7pct$year %in% 2023:2062]) 

#high
S4_NPV_ResRisk_High_dis2pct <- S1_NPV_ResRisk_High_dis2pct - sum(S4_AAL_dis2pct$AAL_High_dis2pct[S4_AAL_dis2pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_High_dis3pct <- S1_NPV_ResRisk_High_dis3pct - sum(S4_AAL_dis3pct$AAL_High_dis3pct[S4_AAL_dis3pct$year %in% 2023:2062]) 

S4_NPV_ResRisk_High_dis7pct <- S1_NPV_ResRisk_High_dis7pct - sum(S4_AAL_dis7pct$AAL_High_dis7pct[S4_AAL_dis7pct$year %in% 2023:2062]) 



# Create a new dataframe 
S4_NPV_ResRisk <- data.frame(
  Scenario = c("Low", "Medium", "High"),
  dis2pct = c(S4_NPV_ResRisk_Low_dis2pct, S4_NPV_ResRisk_Mid_dis2pct, S4_NPV_ResRisk_High_dis2pct),
  dis3pct = c(S4_NPV_ResRisk_Low_dis3pct, S4_NPV_ResRisk_Mid_dis3pct, S4_NPV_ResRisk_High_dis3pct),
  dis7pct = c(S4_NPV_ResRisk_Low_dis7pct, S4_NPV_ResRisk_Mid_dis7pct, S4_NPV_ResRisk_High_dis7pct)
)

# View the dataframe
print(S4_NPV_ResRisk)
```

#S0 NPV
#using the simplyfing assumption that homes are not elevated. Through damages are likely much higher, for illustrative purposes we multiply by a factor of 10 from S0 baseline (See SI for details.
```{r include}
#multiplying AALs by a factor of 10
S0_AAL_interpolated <- S1_AAL_interpolated
S0_AAL_interpolated$InterpolatedAAL_Low <- S0_AAL_interpolated$InterpolatedAAL_Low *10
S0_AAL_interpolated$InterpolatedAAL_Mid <- S0_AAL_interpolated$InterpolatedAAL_Mid *10
S0_AAL_interpolated$InterpolatedAAL_High <- S0_AAL_interpolated$InterpolatedAAL_High *10


#discounting
S0_AAL_dis2pct <- discount_AAL_2pct(S0_AAL_interpolated)
S0_AAL_dis3pct <- discount_AAL_3pct(S0_AAL_interpolated)
S0_AAL_dis7pct <- discount_AAL_7pct(S0_AAL_interpolated)

```


##S0 NPV 
```{r include = FALSE}


S0_NPV_ResRisk_Low_dis2pct <- sum(S0_AAL_dis2pct$AAL_Low_dis2pct[S0_AAL_dis2pct$year %in% 2023:2062])
S0_NPV_ResRisk_Low_dis3pct <- sum(S0_AAL_dis3pct$AAL_Low_dis3pct[S0_AAL_dis3pct$year %in% 2023:2062])
S0_NPV_ResRisk_Low_dis7pct <- sum(S0_AAL_dis7pct$AAL_Low_dis7pct[S0_AAL_dis7pct$year %in% 2023:2062])

S0_NPV_ResRisk_Mid_dis2pct <- sum(S0_AAL_dis2pct$AAL_Mid_dis2pct[S0_AAL_dis2pct$year %in% 2023:2062])
S0_NPV_ResRisk_Mid_dis3pct <- sum(S0_AAL_dis3pct$AAL_Mid_dis3pct[S0_AAL_dis3pct$year %in% 2023:2062])
S0_NPV_ResRisk_Mid_dis7pct <- sum(S0_AAL_dis7pct$AAL_Mid_dis7pct[S0_AAL_dis7pct$year %in% 2023:2062])

S0_NPV_ResRisk_High_dis2pct <- sum(S0_AAL_dis2pct$AAL_High_dis2pct[S0_AAL_dis2pct$year %in% 2023:2062])
S0_NPV_ResRisk_High_dis3pct <- sum(S0_AAL_dis3pct$AAL_High_dis3pct[S0_AAL_dis3pct$year %in% 2023:2062])
S0_NPV_ResRisk_High_dis7pct <- sum(S0_AAL_dis7pct$AAL_High_dis7pct[S0_AAL_dis7pct$year %in% 2023:2062])


# Create a new dataframe with the desired structure
S0_NPV_ResRisk <- data.frame(
  Scenario = c("Low", "Medium", "High"),
  dis2pct = c(S0_NPV_ResRisk_Low_dis2pct, S0_NPV_ResRisk_Mid_dis2pct, S0_NPV_ResRisk_High_dis2pct),
  dis3pct = c(S0_NPV_ResRisk_Low_dis3pct, S0_NPV_ResRisk_Mid_dis3pct, S0_NPV_ResRisk_High_dis3pct),
  dis7pct = c(S0_NPV_ResRisk_Low_dis7pct, S0_NPV_ResRisk_Mid_dis7pct, S0_NPV_ResRisk_High_dis7pct)
)

# View the dataframe
print(S0_NPV_ResRisk)
```

#writing out files
```{r include= FALSE}
path <- './data/BaselineData/FirstStreet'

write.csv(S1_NPV_ResRisk, file.path(path, "S1_NPV_ResRisk.csv"), row.names = FALSE)
write.csv(S3_NPV_ResRisk, file.path(path, "S3_NPV_ResRisk.csv"), row.names = FALSE)
write.csv(S4_NPV_ResRisk, file.path(path, "S4_NPV_ResRisk.csv"), row.names = FALSE)
write.csv(S0_NPV_ResRisk, file.path(path, "S0_NPV_ResRisk.csv"), row.names = FALSE)

path <- './data/BaselineData/S0_OB_Inputs'
write.csv(S0_NPV_ResRisk, file.path(path, "S0_NPV_ResRisk.csv"), row.names = FALSE)

path <- './data/BaselineData/S1_OB_Inputs'
write.csv(S1_NPV_ResRisk, file.path(path, "S1_NPV_ResRisk.csv"), row.names = FALSE)

path <- './data/BaselineData/S3_OB_Inputs'
write.csv(S3_NPV_ResRisk, file.path(path, "S3_NPV_ResRisk.csv"), row.names = FALSE)

path <- './data/BaselineData/S4_OB_Inputs'
write.csv(S4_NPV_ResRisk, file.path(path, "S4_NPV_ResRisk.csv"), row.names = FALSE)

```