---
title: "Federal Revenues from Ortley Beach"
author: "Laura Geronimo"
date: "2024-08-23"
output: html_document
---
The purpose of this worksheet is to estimate the benefits to the Federal Government originating from properties on Ortley Beach for Scenario 1 (replicating the historical record). Benefits are estimated for the 2012-2022 period. 

We only estimate benefits from NFIP premiums.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)

##Importing Data ####
NFIP_PolicyCost_08571 <- read.csv('./data/DataDownloads/NFIP/Policies/produced/ZCTA_08571/NFIP_PolicyData_08571.csv')

#Data for downscaling
OB_DemoData <- read_xlsx('./data/BaselineData/OrtleyBeach_DemographicData.xlsx', sheet=1)
TR_DemoData <- read_xlsx('./data/BaselineData/TomsRiver_DemographicData.xlsx', sheet=1)
ZCTA_DemoData <- read_xlsx('./data/BaselineData/ZCTA_08751_DemographicData.xlsx', sheet=1)

TR_Sandy <- read_xlsx('./data/BaselineData/TomsRiver_SandyImpacts.xlsx', sheet=1)



```



##1) Downscaling NFIP policies data based on estimated proportion of housing units with policies in force
```{r include=FALSE}
#selecting colums for analysis
NFIP_PolicyCost_08571 <- NFIP_PolicyCost_08571 %>% 
  select("policyEffectiveYear",
         "numberOfPolicies",
         "AvePolicyCostAdj")

#joining NFIP data to ZCTA demographic data
ZCTA_NFIP <- left_join(ZCTA_DemoData, NFIP_PolicyCost_08571, by = c("Year" = "policyEffectiveYear"))

#imputing 2021 and 2022 values based on averages
ZCTA_NFIP$numberOfPolicies[ZCTA_NFIP$Year==2021] <- mean(ZCTA_NFIP$numberOfPolicies[ZCTA_NFIP$Year < 2021])
ZCTA_NFIP$numberOfPolicies[ZCTA_NFIP$Year==2022] <- mean(ZCTA_NFIP$numberOfPolicies[ZCTA_NFIP$Year < 2021])

ZCTA_NFIP$AvePolicyCostAdj[ZCTA_NFIP$Year==2021] <- mean(ZCTA_NFIP$AvePolicyCostAdj[ZCTA_NFIP$Year < 2021])
ZCTA_NFIP$AvePolicyCostAdj[ZCTA_NFIP$Year==2022] <- mean(ZCTA_NFIP$AvePolicyCostAdj[ZCTA_NFIP$Year < 2021])

#estimating proportion of housing units in force
ZCTA_NFIP$PoliciesInForce_pct <- ZCTA_NFIP$numberOfPolicies / ZCTA_NFIP$Z_HU

#joining Ortley Beach housing unit estimates
OB_HU <- OB_DemoData %>%
  select(Year,
         OB_TotHU_E,
         OB_TotHU_OwnOcc_E)

OB_NFIP <- left_join(OB_HU, ZCTA_NFIP, by="Year")
OB_NFIP <- OB_NFIP[OB_NFIP$Year > 2011,]

#Downscaling amount of policy cost in Ortley Beach by proportion of homes with policies in force in ZCTA * total housing units in Ortley Beach
OB_NFIP$OB_NFIP_PolicyCost_adj <- OB_NFIP$OB_TotHU_E * OB_NFIP$PoliciesInForce_pct * OB_NFIP$AvePolicyCostAdj

#selecting columns
OB_NFIP_PolicyCost_adj <- OB_NFIP %>%
  select(Year,
         OB_NFIP_PolicyCost_adj)

```


#Exporting S1 OB input tables :
```{r include=FALSE}
path <- 'C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4/data/BaselineData/S1_OB_Inputs/Revenues_ToFed_fromOB'

write.csv(OB_NFIP_PolicyCost_adj, file.path(path, "S1_OB_FedRev_NFIP_PolicyCost_adj.csv"), row.names = TRUE)


```
