---
title: "S1_FedRevFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---

The purpose of this worksheet is to estimate the NPV for benefits to the Federal Government for Scenario 1. We discount the costs and group them into buckets for analysis.

We discount at 2%, 3%, and 7% based on justifications provided in the SI.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedRevOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Revenues_ToFed_fromOB/S1_OB_FedRev_NFIP_PolicyCost_adj.csv')
FedRevOB <- FedRevOB[,c(-1)]

```


#Discounting values to 2%, summing to NPV
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedRevOB), value = TRUE)

# Apply the discount function
FedRevOB_dis2pct <- dis2pct(FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedRevOB_dis2pct <- FedRevOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
FedRevOB_dis2pct_NPV <- FedRevOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

#pivoting tables longer
S1_FedRevOB_dis2pct_NPV <- FedRevOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis3pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedRevOB), value = TRUE)

# Apply the discount function
FedRevOB_dis3pct <- dis3pct(FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedRevOB_dis3pct <- FedRevOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
FedRevOB_dis3pct_NPV <- FedRevOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

#pivoting tables longer
S1_FedRevOB_dis3pct_NPV <- FedRevOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(FedRevOB), value = TRUE)

# Apply the discount function
FedRevOB_dis7pct <- dis7pct(FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
FedRevOB_dis7pct <- FedRevOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
FedRevOB_dis7pct_NPV <- FedRevOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

#pivoting tables longer
S1_FedRevOB_dis7pct_NPV <- FedRevOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```




#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S1_FedRevOB_NPV <- S1_FedRevOB_dis2pct_NPV %>%
  left_join(S1_FedRevOB_dis3pct_NPV, by = "variable") %>%
  left_join(S1_FedRevOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S1_OB_Inputs'

write.csv(S1_FedRevOB_NPV, file.path(path, "S1_FedRevOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S1_FedRevOB_dis2pct_NPV <- sum(S1_FedRevOB_NPV$dis2pct)
S1_FedRevOB_dis3pct_NPV <- sum(S1_FedRevOB_NPV$dis3pct)
S1_FedRevOB_dis7pct_NPV <- sum(S1_FedRevOB_NPV$dis7pct)

S1_FedRevOB_dis2pct_NPV
S1_FedRevOB_dis3pct_NPV
S1_FedRevOB_dis7pct_NPV
```