---
title: "S4_FedRevFrom_OB"
author: "Laura Geronimo"
date: "2024-09-13"
output: html_document
---

The purpose of this worksheet is to estimate the NPV for benefits to the Federal Government for Scenario 4. 

Simplifying Assumptions:
-Assume 50% less than S1 after 2012  due to homes relocating

We discount at 2%, 3%, and 7% based on justifications provided in the SI.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data
```{r include=FALSE}

getwd()
setwd('C:/Users/lgero/Box/Research/NJSG/Tradeoff_Analysis/V4')

#libraries ####
library(tidyverse)
library(networkD3)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(purrr)

options(scipen=999)


##Importing Data ####
FedRevOB <- read.csv('./data/BaselineData/S1_OB_Inputs/Revenues_ToFed_fromOB/S1_OB_FedRev_NFIP_PolicyCost_adj.csv')
FedRevOB <- FedRevOB[,c(-1)]

```

#Adjusting per simplyfying assumptions
```{r include=FALSE}
S4_FedRevOB <- FedRevOB 

S4_FedRevOB$OB_NFIP_PolicyCost_adj[S4_FedRevOB$Year > 2012] <- S4_FedRevOB$OB_NFIP_PolicyCost_adj[S4_FedRevOB$Year > 2012] * 0.50

```


#Discounting values to 2%, summing to NPV
```{r include=FALSE}

###discounting flows at 2% ####
#Creating a year index 
S4_FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 2% 
dis2pct <- function(df, amount_cols, year_index_col, discount_rate = 0.02, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis2pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedRevOB), value = TRUE)

# Apply the discount function
S4_FedRevOB_dis2pct <- dis2pct(S4_FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedRevOB_dis2pct <- S4_FedRevOB_dis2pct %>%
  select(Year, YearIndex, ends_with("_dis2pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S4_FedRevOB_dis2pct_NPV <- S4_FedRevOB_dis2pct %>%
  summarise(across(ends_with("_dis2pct"), sum, na.rm = TRUE))

#pivoting tables longer
S4_FedRevOB_dis2pct_NPV <- S4_FedRevOB_dis2pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis2pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis2pct", "", variable)) 
```



#Discounting values to 3%, summing to NPV
```{r include=FALSE}

###discounting flows at 3% ####
#Creating a year index 
S4_FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 3% 
dis4pct <- function(df, amount_cols, year_index_col, discount_rate = 0.03, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis3pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedRevOB), value = TRUE)

# Apply the discount function
S4_FedRevOB_dis3pct <- dis4pct(S4_FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedRevOB_dis3pct <- S4_FedRevOB_dis3pct %>%
  select(Year, YearIndex, ends_with("_dis3pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (2020USD)
S4_FedRevOB_dis3pct_NPV <- S4_FedRevOB_dis3pct %>%
  summarise(across(ends_with("_dis3pct"), sum, na.rm = TRUE))

#pivoting tables longer
S4_FedRevOB_dis3pct_NPV <- S4_FedRevOB_dis3pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis3pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis3pct", "", variable)) 
```



#Discounting values to 7%, summing to NPV
```{r include=FALSE}

###discounting flows at 7% ####
#Creating a year index 
S4_FedRevOB$YearIndex <- 0:10

#Define the function for discounting flows by 7% 
dis7pct <- function(df, amount_cols, year_index_col, discount_rate = 0.07, base_year = 0) {
  for (col in amount_cols) {
    discounted_col_name <- paste0(col, "_dis7pct")
    df[[discounted_col_name]] <- df[[col]] / ((1 + discount_rate) ^ (df[[year_index_col]] - base_year))
  }
  return(df)
}

# List of columns to discount - select colums starting with OB 
amount_cols <-  grep("^OB_", colnames(S4_FedRevOB), value = TRUE)

# Apply the discount function
S4_FedRevOB_dis7pct <- dis7pct(S4_FedRevOB, amount_cols, year_index_col = "YearIndex")

# Remove undiscounted columns
S4_FedRevOB_dis7pct <- S4_FedRevOB_dis7pct %>%
  select(Year, YearIndex, ends_with("_dis7pct"))

###Summing to NPV ####
#Create a new data frame that sums the discounted columns for NPV (7070USD)
S4_FedRevOB_dis7pct_NPV <- S4_FedRevOB_dis7pct %>%
  summarise(across(ends_with("_dis7pct"), sum, na.rm = TRUE))

#pivoting tables longer
S4_FedRevOB_dis7pct_NPV <- S4_FedRevOB_dis7pct_NPV %>%
  pivot_longer(
    cols = everything(),      # Pivot all columns
    names_to = "variable",     # New column with variable names
    values_to = "dis7pct"      # New column with the values
  ) %>%
  mutate(variable = sub("_dis7pct", "", variable)) 
```


#Joining data for export
```{r include = FALSE}
#pivoting tables longer
S4_FedRevOB_NPV <- S4_FedRevOB_dis2pct_NPV %>%
  left_join(S4_FedRevOB_dis3pct_NPV, by = "variable") %>%
  left_join(S4_FedRevOB_dis7pct_NPV, by = "variable")

```

#Writing out discounted NPV files
```{r include= FALSE}
path <- './data/BaselineData/S4_OB_Inputs'

write.csv(S4_FedRevOB_NPV, file.path(path, "S4_FedRevOB_NPV.csv"), row.names = TRUE)

```

#QC
```{r include= FALSE}
S4_FedRevOB_dis2pct_NPV <- sum(S4_FedRevOB_NPV$dis2pct)
S4_FedRevOB_dis3pct_NPV <- sum(S4_FedRevOB_NPV$dis3pct)
S4_FedRevOB_dis7pct_NPV <- sum(S4_FedRevOB_NPV$dis7pct)

S4_FedRevOB_dis2pct_NPV
S4_FedRevOB_dis3pct_NPV
S4_FedRevOB_dis7pct_NPV
```